### 2020-07-02 presentation/presentation_patronage_network.Rmd
first commit

---
title: "Government through Patronage"
subtitle: "Bargaining for Education in Decentralized Brazil"
author: "Galileu Kim"
institute: "Princeton University"
output:
  beamer_presentation:
    includes:
      in_header: "presentation_template.txt"
    incremental: false
    keep_tex: true
    slide_level: 2
---

# Introduction

- Why should we care about patronage in specific?
- What can I bring into the table that scholars have not thought about?
- What are the gaps in the literature and the theoretical and empirical contribution I bring to the table?
- Can I dream of being ambitious? Yes, I am much better now than I used to be before.
- Bring both my theoretical treatment and empirical estimation to the fore.

# Different types of patronage
- Politicians value different types of positions for patronage.
- For the mayor, who relies less on particularistic transfers (more on that later), control over the higher echelons of the bureaucracy is more important.
- For city councilors, ties to the local community are more important: neighbors rely on public sector employment.

# Betting on horses
- Voters are strategic.
- They should contribute to candidates who they believe will make them likely to obtain employment.
- If I can link the people who donated tho candidates with the likelihood of them entering a bureaucracy.
- Something very interesting can be done.

# Differentiating types of donors
- For some voters, campaign contributino is symbolic.
- In Brazil, it is more likely that donors are thinking strategically about their candidates.
- Who do they contribute to? What is there to win?
- Who are the public sector workers more likely to donate?
- How does this affect their likelihood of being in office?

# Campaign contribution and employment
- It is clear that this study dialogues with another set of studies.
- Colonnelli et al. come to mind. The problem is that they have already done it.
- Well, at least for mayors. Look at how campaign contributions increase the likelihood of obtaining employment.
- So what else is there to be done?
- Emphasizing the city councilor link is not enough. I need to do more.

# Trade-offs
- For voters t, there is a clear trade-off between the immediate gains of employment and the uncertain gains from public goods provision.
- It is essentially the fact of hedging and how citizens are willing to take the public sector employment for themselves, which can help them provide for their family needs, rather than having to rely on public schools or well-staffed hospitals.
- It is a tragedy of the commons, but more subtle.
- Let's spell this out.

# Context and data:
- Brazil has multiple city councilors, who rely on patronage networks to be reelected.
- The question then is what is the additional

### 2020-07-02 presentation/presentation_patronage_network.tex
first commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Government through Patronage},
  pdfauthor={Galileu Kim},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usetheme{metropolis}
\setbeamertemplate{footline}[page number]{}

% packages
\usepackage{xcolor}
\usepackage{ragged2e}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{blindtext}
\captionsetup[figure]{font=scriptsize} % set size of caption
\AtBeginEnvironment{knitrout}{\singlespacing}
\useoutertheme{infolines} % Alternatively: miniframes, infolines, split
\useinnertheme{circles}
\date[]{\today}

% colors
\definecolor{princeton}{rgb}{1.0, 0.6, 0.2} % princeton (primary)
\definecolor{arsenic}{rgb}{0.23, 0.27, 0.29}

\setbeamercolor{title}{fg=black}
\setbeamercolor{frametitle}{fg=white,bg=arsenic}
\setbeamercolor{section in head/foot}{fg=arsenic,bg=princeton}
\setbeamerfont{frametitle}{size=\Large}
\setbeamercolor{palette primary}{fg=white,bg=arsenic}
\setbeamercolor{palette secondary}{fg=arsenic,bg=princeton}
\setbeamercolor{palette tertiary}{fg=princeton,bg=arsenic}
\setbeamercolor{palette quaternary}{fg=princeton,bg=arsenic}
\setbeamercolor{structure}{fg=princeton} % itemize, enumerate, etc
\setbeamercolor{section in toc}{fg=arsenic} % TOC sections
\setbeamerfont{subtitle}{size=\small}

% override palette coloring with secondary
\setbeamercolor{subsection in head/foot}{bg=arsenic,fg=white}

% defining font
\usepackage[T1]{fontenc}

% Cabin
% \usepackage[sfdefault,condensed]{cabin}

% Helvetica
\usepackage[scaled]{helvet}
\renewcommand\familydefault{\sfdefault}

% Inria Sans
%\usepackage[lining]{InriaSans}
%\renewcommand*\oldstylenums[1]{{\fontfamily{InriaSans-OsF}\selectfont #1}}
%% The font package uses mweights.sty which has som issues with the
%% \normalfont command. The following two lines fixes this issue.
%\let\oldnormalfont\normalfont
%\def\normalfont{\oldnormalfont\mdseries}

% FiraSans
%\usepackage[sfdefault]{FiraSans}
%\renewcommand*\oldstylenums[1]{{\firaoldstyle #1}}
%\usepackage{ragged2e}

% aesthetics
\setbeamersize{text margin left=10mm,text margin right=10mm}

% source command
\newcommand{\source}{\footnotesize\textcolor{black!25}}

% suppress section slides
\AtBeginPart{}
\AtBeginSubsection{}
\AtBeginSection{}
\setlength{\emergencystretch}{0em}
\setlength{\parskip}{0pt}

\title{Government through Patronage}
\subtitle{Bargaining for Education in Decentralized Brazil}
\author{Galileu Kim}
\date{}
\institute{Princeton University}

\begin{document}
\frame{\titlepage}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{frame}{Introduction}
\begin{itemize}
\tightlist
\item
  Why should we care about patronage in specific?
\item
  What can I bring into the table that scholars have not thought about?
\item
  What are the gaps in the literature and the theoretical and empirical
  contribution I bring to the table?
\item
  Can I dream of being ambitious? Yes, I am much better now than I used
  to be before.
\item
  Bring both my theoretical treatment and empirical estimation to the
  fore.
\end{itemize}
\end{frame}

\hypertarget{different-types-of-patronage}{%
\section{Different types of
patronage}\label{different-types-of-patronage}}

\begin{frame}{Different types of patronage}
\begin{itemize}
\tightlist
\item
  Politicians value different types of positions for patronage.
\item
  For the mayor, who relies less on particularistic transfers (more on
  that later), control over the higher echelons of the bureaucracy is
  more important.
\item
  For city councilors, ties to the local community are more important:
  neighbors rely on public sector employment.
\end{itemize}
\end{frame}

\hypertarget{betting-on-horses}{%
\section{Betting on horses}\label{betting-on-horses}}

\begin{frame}{Betting on horses}
\begin{itemize}
\tightlist
\item
  Voters are strategic.
\item
  They should contribute to candidates who they believe will make them
  likely to obtain employment.
\item
  If I can link the people who donated tho candidates with the
  likelihood of them entering a bureaucracy.
\item
  Something very interesting can be done.
\end{itemize}
\end{frame}

\hypertarget{differentiating-types-of-donors}{%
\section{Differentiating types of
donors}\label{differentiating-types-of-donors}}

\begin{frame}{Differentiating types of donors}
\begin{itemize}
\tightlist
\item
  For some voters, campaign contributino is symbolic.
\item
  In Brazil, it is more likely that donors are thinking strategically
  about their candidates.
\item
  Who do they contribute to? What is there to win?
\item
  Who are the public sector workers more likely to donate?
\item
  How does this affect their likelihood of being in office?
\end{itemize}
\end{frame}

\hypertarget{campaign-contribution-and-employment}{%
\section{Campaign contribution and
employment}\label{campaign-contribution-and-employment}}

\begin{frame}{Campaign contribution and employment}
\begin{itemize}
\tightlist
\item
  It is clear that this study dialogues with another set of studies.
\item
  Colonnelli et al.~come to mind. The problem is that they have already
  done it.
\item
  Well, at least for mayors. Look at how campaign contributions increase
  the likelihood of obtaining employment.
\item
  So what else is there to be done?
\item
  Emphasizing the city councilor link is not enough. I need to do more.
\end{itemize}
\end{frame}

\hypertarget{trade-offs}{%
\section{Trade-offs}\label{trade-offs}}

\begin{frame}{Trade-offs}
\begin{itemize}
\tightlist
\item
  For voters t, there is a clear trade-off between the immediate gains
  of employment and the uncertain gains from public goods provision.
\item
  It is essentially the fact of hedging and how citizens are willing to
  take the public sector employment for themselves, which can help them
  provide for their family needs, rather than having to rely on public
  schools or well-staffed hospitals.
\item
  It is a tragedy of the commons, but more subtle.
\item
  Let's spell this out.
\end{itemize}
\end{frame}

\hypertarget{context-and-data}{%
\section{Context and data:}\label{context-and-data}}

\begin{frame}{Context and data:}
\begin{itemize}
\tightlist
\item
  Brazil has multiple city councilors, who rely on patronage networks to
  be reelected.
\item
  The question then is what is the additional
\end{itemize}
\end{frame}

\end{document}

### 2020-07-02 build/build.R

@@ -1,0 +1,4 @@
+# data cleaning
+# replication (estimation, visualization)
+# publication (relatorio)
+# presentation

### 2020-07-02 build/data/build.R

@@ -1,0 +1,36 @@
+# ==============================================================================
+# set-up
+# ==============================================================================
+pacman::p_load(
+  "tidyverse",
+  "purrr",
+  "data.table",
+  "R.utils",
+  "here",
+  "parallel",
+  "magrittr"
+)
+
+set.seed(1789)
+
+source(
+  here("build", "data", "functions.R")
+)
+
+# ==============================================================================
+# build data repos
+# ==============================================================================
+repos <- c(
+  "censo_br",
+  "censo_escolar",
+  "finbra",
+  "rais",
+  "saeb",
+  "tse",
+  "spaece"
+)
+
+walk(
+  repos,
+  ~ run_module(.)
+)

### 2020-07-02 build/data/functions.R

@@ -1,0 +1,389 @@
+# functions ---------------------------------------------------------------
+`%<>%` <- magrittr::`%<>%`
+ivreg <- AER::ivreg
+between <- data.table::between
+
+# io ----------------------------------------------------------------------
+here_data <- function(type, dir, file) {
+  path <- here("data", type, dir, file)
+  
+  return(path)
+}
+
+read_data <- function(type, dir, file) {
+  file_path <- here_data(type, dir, file)
+
+  if (str_detect(file, ".rds$")) {
+    data <- read_rds(file_path)
+  } else {
+    data <- fread(file_path)
+  }
+  
+  return(data)
+}
+
+write_data <- function(object, dir, file, type = "clean", compress = "gz") {
+  file_path <- here_data(type, dir, file)
+
+  if (str_detect(file, ".rds$")) {
+    write_rds(object, file_path, compress = compress)
+  } else {
+    fwrite(object, file_path, compress = "gz")
+  }
+}
+
+reset_env <- function(init_env){
+  final_env <- ls(.GlobalEnv)
+
+  rm(
+    envir = .GlobalEnv,
+    list = setdiff(final_env, init_env)
+  )
+
+  gc()
+}
+
+build_repo <- function(module){
+  repo <- here("data", "clean", module)
+
+  if(!dir.exists(repo)){
+    dir.create(repo)
+  }else{
+    unlink(repo, recursive = T)
+    dir.create(repo)
+    }
+  }
+
+run_module <- function(module) {
+  print(
+    paste0("running module ", module, "...")
+  )
+
+  build_repo(module)
+
+  source_file <- paste0("datagen_", module, ".R")
+  path <- here("build", "data", "modules", source_file)
+
+  init_env <- ls(.GlobalEnv)
+  source(path)
+  reset_env(init_env)
+
+  print(
+    paste("module", module, "complete!")
+  )
+}
+
+fread <- partial(
+  data.table::fread,
+  nThread = parallel::detectCores(),
+  integer64 = c("character")
+)
+
+list.files <- partial(
+  base::list.files,
+  full.names = T
+)
+
+list_files <- function(path, pattern) {
+  files <- map2(
+    path,
+    pattern,
+    list.files
+  ) %>%
+    flatten_chr()
+
+  return(files)
+}
+
+import_data <- function(path, pattern, names = F) {
[…]

### 2020-07-02 build/data/modules/datagen_censo_escolar.R

@@ -1,0 +1,279 @@
+# use school census data to produce set of tables
+# 1) all municipal schools
+# 2) all teachers
+# 3) municipal classrooms.
+# 4) all classrooms.
+# 5) teacher turnover (municipal and all)
+init_env <- ls()
+
+censo_school <- read_data(
+  "raw",
+  "censo_escolar",
+  "censo_school.csv.gz"
+)
+
+# keep only active schools and generate municipal table
+censo_school_mun <- censo_school %>%
+  filter(
+    str_detect(active, "(1|2|ativ)")
+  ) %>%
+  group_by(
+    cod_ibge_6,
+    year,
+    dep
+  ) %>%
+  summarise(
+    n_school = n(),
+    n_room = mean(rooms_existing),
+    n_lower_sch = sum(
+      elementary_sch,
+      elementary_sch_8_9,
+      elementary_sch_9,
+      na.rm = T
+    ),
+    n_high_sch = sum(
+      high_sch,
+      high_sch_normal,
+      na.rm = T
+    ),
+    prop_meal = mean(
+      meal,
+      na.rm = T
+    ),
+    prop_electricity = 1 - mean(
+      electricity_inexistent,
+      na.rm = T
+    ),
+    prop_internet = mean(
+      internet,
+      na.rm = T
+    ),
+    prop_kitchen = mean(
+      kitchen,
+      na.rm = T
+    ),
+    prop_sewer = 1 - mean(
+      sewer_inexistent,
+      na.rm = T
+    ),
+    prop_rural = mean(
+      location == "rural",
+      na.rm = T
+    ),
+    n_lab_info = sum(
+      lab_info,
+      na.rm = T
+    ),
+    n_library = sum(
+      library,
+      na.rm = T
+    ),
+    n_staff = sum(
+      staff,
+      na.rm = T
+    ),
+    n_teachers = sum(
+      num_tchr,
+      na.rm = T
+    ),
+    n_pta = sum(
+      association_parents,
+      na.rm = T
+    ),
+    n_tchr_unions = sum(
+      association_tchrs,
+      na.rm = T
+    )
+  ) %>%
+  ungroup()
+
+# fix entries
+censo_school_mun <- censo_school_mun %>%
+  mutate(
+    n_pta = if_else(year == 2004, n_pta, NA_integer_),
+    n_tchr_unions = if_else(year == 2004, n_tchr_unions, NA_integer_)
+  )
+
[…]

### 2020-07-02 build/data/modules/datagen_saeb.R

@@ -1,0 +1,83 @@
+# generate municipal table
+saeb_exam_mun <- read_data(
+  "raw",
+  "saeb",
+  "saeb_exam_mun.csv"
+) %>%
+  add_election_year()
+
+saeb_exam_mun %>%
+  write_data(
+    "saeb",
+    "saeb_exam_mun.rds"
+  )
+
+# saeb_sample
+saeb_student <- read_data(
+  "raw",
+  "saeb",
+  "saeb_student.csv.gz"
+)
+
+# breakdown by administrative department
+saeb_school <- saeb_student %>%
+  filter(year >= 2007) %>%
+  mutate(
+    brown_black_student = if_else(race_student %in% c("brown", "black"), 1, 0)
+  ) %>%
+  group_by(
+    cod_ibge_6,
+    year,
+    dep,
+    cod_school
+  ) %>%
+  summarise(
+    across(
+      c(
+        failed_school_year_student,
+        fridge_student,
+        housekeeper_student,
+        father_home_student,
+        mother_home_student,
+        pc_student,
+        age_student,
+        brown_black_student,
+        grade_exam
+      ),
+      ~ mean(as.numeric(.x), na.rm = T)
+    ),
+    .groups = "drop"
+  )
+
+saeb_student %>%
+  sample_frac(0.25) %>%
+  write_data(
+    "saeb",
+    "saeb_student_sample.rds"
+  )
+
+saeb_school %>%
+  write_data(
+    "saeb",
+    "saeb_school.rds"
+  )
+
+
+# hierarchical model
+saeb_hierarchical <- read_data(
+  "raw",
+  "saeb",
+  "saeb_hierarchical.csv.gz"
+)
+
+saeb_hierarchical <- saeb_hierarchical %>%
+  rename(
+    school_id = cod_school,
+    grade_level = grade
+  )
+  
+saeb_hierarchical %>%
+  write_data(
+    "saeb",
+    "saeb_hierarchical.rds"
+  )

### 2020-07-02 build/replication/build.R

@@ -1,0 +1,744 @@
+# ==============================================================================
+# run modules
+# ==============================================================================
+source(
+  here("build", domain, "setup.R")
+)
+
+# run tasks
+modules <- c(
+  "visual_map"
+  "visual_global_edu",
+  "visual_budget",
+  "visual_bureaucracy",
+  "model_hierarchical",
+  "model_saeb"
+  )
+
+walk(
+  modules,
+  ~run_module(.)
+)
+
+# turnover spaece ---------------------------------------------------------
+plot_spaece <- censo_turnover_score %>% 
+  filter(
+    n >= 5
+  ) %>% 
+  mutate(
+    grade_level = as.factor(grade_level)
+  ) %>% 
+  ggplot(
+    aes(
+      turnover_index,
+      spaece_mean,
+      group = as.factor(grade_level),
+      col = as.factor(grade_level)
+    )
+  ) +
+  geom_smooth(
+    method = lm, formula = y ~ splines::bs(x, 3),
+    se = F
+  ) +
+  annotate(
+    "text",
+    label = "2nd grade",
+    size = 2,
+    col = "grey65",
+    0.45, 165
+  ) +
+  annotate(
+    "text",
+    label = "5th grade",
+    size = 2,
+    col = "grey65",
+    0.6, 200
+  ) +
+  annotate(
+    "text",
+    label = "9th grade",
+    size = 2,
+    col = "grey65",
+    0.7, 225
+  ) +
+  labs(
+    x = "Teacher turnover (index)",
+    y = "Test scores (SPAECE)"
+  ) +
+  theme(
+    legend.position = "none"
+  )
+
+ggsave(
+  plot_spaece,
+  filename = p_file_here('figs', "turnover_spaece.pdf")
+)
+
+# explaining turnover -----------------------------------------------------
+# show that irrespective of level of economic development, the same dynamic is present
+# turnover index on coalition
+saeb_turnover <- censo_school_turnover %>% 
+  join_covariate() %>% 
+  left_join(
+    censo_school %>% 
+      mutate(
+        school_id = as.integer(school_id)
+      ) %>% 
+      select(-state, -location),
[…]

### 2020-07-02 build/replication/functions.R

@@ -1,0 +1,747 @@
+# ==============================================================================
+# workflow functions
+# ==============================================================================
+`%<>%` <- magrittr::`%<>%`
+
+between <- data.table::between
+
+build_repo <- function(module){
+  repo <- here("data", "clean", module)
+
+  if(!dir.exists(repo)){
+    dir.create(repo)
+  }else{
+    unlink(repo, recursive = T)
+    dir.create(repo)
+  }
+}
+
+list_data <- function(dir){
+  path <- here("data", "clean", dir)
+  files <- list.files(path)
+
+  return(files)
+}
+
+here_data <- function(type, dir, file) {
+  path <- here("data", type, dir, file)
+  
+  return(path)
+}
+
+read_data <- function(type, dir, file) {
+  file_path <- here_data(type, dir, file)
+
+  if (str_detect(file, ".rds$")) {
+    data <- read_rds(file_path)
+  } else {
+    data <- fread(file_path)
+  }
+  
+  print_obj_size(data)
+
+  return(data)
+}
+
+print_obj_size <- function(obj) {
+  obj_size <- object.size(obj)
+  print(obj_size, units = "MB")
+}
+
+ls_data <- function(env = .GlobalEnv) {
+  obj <- ls(env)
+
+  is_data <- map_lgl(
+    obj,
+    ~is.data.frame(get(.))
+  )
+
+  data_obj <- obj[is_data]
+
+  return(data_obj)
+}
+
+read_model <- function(file){
+  fit <- read_rds(
+    here("replication", "models", file)
+  )
+
+  print_obj_size(fit)
+
+  return(fit)
+}
+
+write_model <- function(model, file) {
+  print_obj_size(model)
+  
+  write_rds(
+    model,
+    here("replication", "models", file)
+  )
+}
+
+load_p <- function(packages){
+  walk(
+    packages,
+    ~library(., character.only = T)
+  )
+}
+
+detach_p <- function(packages){
+  packages %>%
+    purrr::walk(
+      ~detach(
+      paste0("package:", .),
+      character.only = T,
+      unload = T
+    )
+  )
+}
+
+run_module <- function(domain, module) {
+  print(
+    paste0("running module ", module, "...")
+  )
+
+  source_setup(domain)
+
+  if(domain == "data"){
+    build_repo(module)
+  }
+  
+  module_script <- paste0(module, ".R")
+
[…]

### 2020-07-02 build/replication/modules/model_turnover_saeb.R

@@ -1,0 +1,165 @@
+# ==============================================================================
+# hierarchical estimation
+# ==============================================================================
+f <- stats::as.formula
+
+saeb <- read_data(
+  "saeb",
+  "saeb_hierarchical.rds"
+)
+
+finbra <- read_data(
+  "finbra",
+  "finbra.rds"
+)
+
+censo_school <- read_data(
+  "censo_escolar",
+  "censo_school.rds"
+) %>%
+  mutate(
+    school_id = as.integer(school_id)
+  )
+
+censo_school_turnover <- read_data(
+  "censo_escolar",
+  "censo_school_turnover.rds"
+) %>%
+  select(
+    cod_ibge_6,
+    year,
+    school_id,
+    grade_level,
+    turnover_index,
+    starts_with("percent"),
+    n_teacher = n
+  ) %>%
+  filter(
+    n_teacher >= 5
+  )
+
+# prepare data for estimation
+saeb_hierarchical <- lst(
+  saeb,
+  finbra,
+  censo_school,
+  censo_school_turnover
+) %>%
+  reduce(
+    left_join
+  ) %>%
+  mutate(
+    saeb_principal_experience = fct_relevel(saeb_principal_experience, "2 to 10"),
+    saeb_teacher_work_school = fct_relevel(saeb_teacher_work_school, "2 to 10"),
+    censo_log_pop = log(censo_pop),
+    budget_education_capita = budget_education / censo_pop
+  )
+
+# fix blank strings
+saeb_hierarchical <- saeb_hierarchical %>%
+  mutate_all(
+    ~ na_if(., "")
+  ) %>%
+  mutate_if(
+    is.double,
+    scale
+  )
+
+# estimation of hierarchical linear models to assess effect of teacher turnover on student learning
+fe <- c(
+  "(1 | year)",
+  "(1 | state)"
+)
+
+teacher_cov <- c(
+  "saeb_wage_teacher",
+  "education_teacher",
+  "gender_teacher"
+  # "year_as_teacher"
+)
+
+principal_cov <- c(
+  "saeb_principal_female",
+  "saeb_principal_education",
+  "saeb_principal_appointment"
+)
+
+student_cov <- c(
+  "parent_attend_college_student",
+  "failed_school_year_student"
+)
+
+school_cov <- c(
+  "access_water",
+  "access_electricity",
+  "library",
+  "meal"
+)
+
+mun_cov <- c(
[…]

### 2020-07-02 build/replication/modules/model_turnover_spaece.R

@@ -1,0 +1,134 @@
+# ==============================================================================
+# spaece: annual standardized test data for the state of ceara
+# ==============================================================================
+source(
+  here::here("build", "replication", "setup.R")
+)
+
+spaece <- read_data(
+  "spaece",
+  "spaece.rds"
+)
+
+finbra <- read_data(
+  "finbra",
+  "finbra.rds"
+)
+
+censo_school <- read_data(
+  "censo_escolar",
+  "censo_school.rds"
+) %>%
+  mutate(
+    school_id = as.integer(school_id)
+  )
+
+censo_school_turnover <- read_data(
+  "censo_escolar",
+  "censo_school_turnover.rds"
+) %>%
+  select(
+    cod_ibge_6,
+    year,
+    school_id,
+    grade_level,
+    turnover_index,
+    starts_with("percent"),
+    n_teacher = n
+  ) %>%
+  filter(
+    n_teacher >= 5
+  )
+
+censo_school_ceara <- censo_school_turnover %>% 
+  filter(
+    str_sub(cod_ibge_6, 1, 2) == "23",
+    grade_level %in% c(2, 5, 9)
+  )
+
+spaece_turnover <- censo_school_ceara %>% 
+  left_join(
+    censo_school %>% 
+      filter(dep == 'municipal', year >= 2007),
+    by = c('cod_ibge_6', 'school_id', 'year')
+  ) %>% 
+  left_join(
+    spaece,
+    by = c("cod_ibge_6", "school_id", "year", "grade_level")
+  ) %>% 
+  join_covariate() %>% 
+  mutate(
+    mandate_year = if_else(
+      year %in% seq(2005, 2013, 4), 1, 0
+    )
+  )
+
+spaece_turnover <- spaece_turnover %>% 
+  mutate(
+    state = str_sub(cod_ibge_6, 1, 2),
+    budget_education_capita = budget_education / censo_pop
+  ) %>% 
+  mutate_at(
+    vars(grade_level, year, state),
+    as.factor
+  ) %>% 
+  mutate_if(
+    is.double,
+    scale
+  )
+
+school_cov <- c(
+  "access_water",
+  "access_electricity",
+  "library",
+  "meal"
+)
+
+mun_cov <- c(
+  "censo_median_wage",
+  "censo_log_pop",
+  "censo_rural",
+  "censo_lit_rate",
+  "budget_education_capita"
+)
+
[…]

### 2020-07-02 build/replication/modules/visual_budget.R

@@ -1,0 +1,43 @@
+finbra <- read_data(
+    "finbra",
+    "finbra.rds"
+)
+
+plot_budget <- finbra %>%
+    pivot_longer(
+        -c(cod_ibge_6, year),
+        names_to = "category",
+        values_to = "value"
+    ) %>%
+    group_by(year, category) %>%
+    summarise(
+        value = sum(value, na.rm = T) / 1e9
+    ) %>%
+    filter(
+        !str_detect(category, "total|legislative|industry")
+    ) %>%
+    ggplot(
+        aes(
+            year,
+            value,
+        )
+    ) +
+    geom_area(
+        aes(
+            fill = category
+        )
+    ) +
+    scale_fill_discrete(
+        name = "HELLO",
+        labels = c("A", "B", "C")
+    ) +
+    labs(
+        # fill = "Category",
+        x = "Year",
+        y = "Total (billions)"
+    )
+
+save_fig(
+    plot_budget,
+    "descriptive_budget.pdf"
+)

### 2020-07-02 build/replication/modules/visual_electoral.R

@@ -1,0 +1,102 @@
+election <- read_data(
+    "tse",
+    "election.rds"
+)
+
+censo <- read_data(
+    "censo_br",
+    "censo_2000.rds"
+)
+
+rais <- read_data(
+    "rais",
+    "rais_mun_edu.rds"
+)
+
+# 1) visualize underlying distribution of coalition shares
+hist_coalition <- election %>%
+    gg_histogram(
+        aes(coalition_share * 100),
+        bins = 20
+    ) +
+    labs(
+        x = "Share of seats (percentage)",
+        y = "Count"
+    )
+
+save_fig(
+    hist_coalition,
+    "hist_coalition.pdf"
+)
+
+# 2) add additional context on how these are structured (size, composition)
+# address the problem of confounding even if partially
+covs <- quo(
+    c(
+        coalition_share,
+        chamber_size,
+        censo_pop,
+        censo_rural,
+        censo_median_wage,
+        rais_size,
+        rais_higher_edu,
+        rais_permanent,
+        rais_wage
+    )
+)
+
+rais_mun <- rais %>%
+    group_summarise(
+        c(cod_ibge_6, election_year),
+        c(rais_size, rais_higher_edu, rais_permanent, rais_wage),
+        ~ mean(., na.rm = T)
+    )
+
+election_covariate <- election %>%
+    left_join(
+        censo,
+        by = "cod_ibge_6"
+    ) %>%
+    left_join(
+        rais_mun,
+        by = c("cod_ibge_6", "election_year")
+    ) %>%
+    select(
+        {{covs}}
+    ) %>%
+    na.omit()
+
+cov_balance <- cobalt::bal.tab(
+    treat = election_covariate %>% pull(coalition_share),
+    covs = election_covariate %>% select({{covs}}, -coalition_share),
+    thresholds = c(m = 0.1)
+)
+
+love_plot <- cobalt::love.plot(
+    cov_balance,
+    threshold = 0.1,
+    title = NULL,
+    colors = matte_indigo,
+    position = "none"
+) +
+    scale_y_discrete(
+        labels = c(
+            "chamber_size" = "chamber size",
+            "censo_pop" = "population",
+            "censo_rural" = "rural population",
+            "censo_median_wage" = "median municipal wage",
+            "rais_size" = "educational staff size",
+            "rais_higher_edu" = "prop. of staff with higher education",
+            "rais_permanent" = "prop. of staff with tenured contract",
[…]

### 2020-07-02 build/replication/modules/visual_global_edu.R

@@ -1,0 +1,251 @@
+# create descriptive visualizations for educational data
+edu_global <- read_data(
+  "worldbank",
+  "edu_global.rds"
+)
+
+plot_enroll <- edu_global %>% 
+  filter(!is.na(enrollment_rate_primary)) %>% 
+  ggplot(
+    aes(
+      year,
+      enrollment_rate_primary,
+      col = type
+    )
+  ) +
+  geom_line(
+    linetype = 'solid'
+  ) +
+  geom_point() +
+  annotate(
+    "text",
+    label = "Global average",
+    size = 2,
+    col = "grey65",
+    2010, 85
+  ) +
+  labs(
+    y = "Enrollment (percentage)",
+    x = "Year"
+  ) +
+  theme(
+    legend.position = "none"
+  ) +
+  coord_cartesian(
+    ylim = c(75, 100),
+    xlim = c(2000, 2016)
+  )
+
+plot_pisa <- edu_global %>% 
+  filter(!is.na(pisa)) %>% 
+  ggplot(
+    aes(
+      year,
+      pisa,
+      col = type
+    )
+  ) +
+  geom_line(
+    linetype = 'solid'
+  ) +
+  geom_point() +
+  annotate(
+    "text",
+    label = "Global average",
+    size = 2,
+    col = "grey65",
+    2011, 490
+  ) +
+  labs(
+    y = "Pisa scores (average)",
+    x = "Year"
+  ) +
+  theme(
+    legend.position = "none"
+  ) +
+  coord_cartesian(
+    xlim = c(2000, 2016),
+    ylim = c(325, 500)
+  )
+
+plot_global <- cowplot::plot_grid(
+  plot_enroll,
+  plot_pisa,
+  ncol = 2
+)
+
+ggsave(
+  plot_global,
+  filename = p_file_here('figs', 'plot_global.pdf')
+)
+
+plot_enroll <- edu_global %>%
+  filter(type == "global") %>% 
+  ggplot(
+    aes(
+      year,
+      enrollment_rate_primary,
+      col = "grey"
+    )
+  ) +
+  geom_line(
+    linetype = 'solid'
+  ) +
+  geom_point() +
+  annotate(
+    "text",
+    label = "Global average",
+    size = 2,
+    col = "grey65",
+    2010, 85
+  ) +
+  labs(
+    y = "Enrollment (percentage)",
+    x = "Year"
+  ) +
[…]

### 2020-07-02 build/replication/modules/visual_map.R

@@ -1,0 +1,124 @@
+# plot municipal map
+plot_map <- function(
+                     data,
+                     fill,
+                     breaks = 6,
+                     title = "",
+                     label = "",
+                     palette = "RdYlBu",
+                     legend_position = "bottom",
+                     limits = NULL) {
+    plot <- ggplot() +
+        geom_polygon(
+            data = data,
+            aes(
+                x = long,
+                y = lat,
+                group = group,
+                fill = get(fill)
+            ),
+            color = NA
+        ) +
+        {
+            if (is.factor(data[[fill]])) {
+                scale_fill_manual(
+                    values = rev(
+                        RColorBrewer::brewer.pal(n = length(levels(data[[fill]])), name = palette)
+                    ),
+                    breaks = levels(data[[fill]]),
+                    na.value = "gray50"
+                )
+            } else {
+                scale_fill_distiller(
+                    palette = palette,
+                    breaks = scales::pretty_breaks(breaks),
+                    direction = -1,
+                    na.value = "gray50",
+                    limits = limits
+                )
+            }
+        } +
+        guides(fill = guide_legend(reverse = T)) +
+        labs(fill = label) +
+        theme_void(base_size = 17) +
+        xlim(range(data$long)) +
+        ylim(range(data$lat)) +
+        coord_quickmap() +
+        theme(
+            legend.position = legend_position,
+            legend.text = element_text(size = 18),
+            text = element_text(size = 18),
+            legend.title = element_blank(),
+            plot.caption = element_text(size = 8)
+        ) +
+        ggtitle(title)
+
+    return(plot)
+}
+
+# build map of average test scores by municipality
+saeb_exam_mun <- read_data(
+    "saeb",
+    "saeb_exam_mun.rds"
+)
+
+map_br <- rgdal::readOGR(
+    here("data/clean/maps/"),
+    "municipio"
+)
+
+map_br <- sp::SpatialPolygonsDataFrame(
+    rgeos::gSimplify(map_br, tol = 0.01, topologyPreserve = T),
+    data = map_br@data
+)
+
+map_br@data <- map_br@data %>%
+    mutate(id = row.names(map_br))
+
+map_br <- map_br %>%
+    fortify() %>%
+    left_join(., map_br@data, by = "id") %>%
+    transmute(
[…]

### 2020-07-02 build/replication/modules/visual_model_turnover.R

@@ -1,0 +1,130 @@
+# ==============================================================================
+# visualize combined results from saeb hierarchical and spaece models
+# ==============================================================================
+source(
+  here::here("build", "replication", "setup.R")
+)
+
+fit_lmer <- read_model("fit_saeb_hierarchical.rds")
+fit_spaece <- read_model("fit_spaece.rds")
+
+sink(here("replication", "results", "student_learning.tex"))
+mstar(
+  c(fit_lmer, fit_spaece),
+  keep = c("turnover_index", "saeb_principal_experience", "saeb_teacher_work_school"),
+  add.lines = list(c("Controls", rep(c("\\_", "\\checkmark"), 3))),
+  dep.var.labels.include = F,
+  covariate.labels = c(
+    "Turnover index", "Turnover index $\\times$ Grade 9",
+    "Teacher experience (2 years)", "Teacher experience (10 years)",
+    "School principal experience (2 years)", "School principal experience (10 years)"
+  ),
+  dep.var.caption = "Student learning",
+  column.labels = c("SAEB test score", "SPAECE test score"),
+  column.separate = c(4, 2),
+  model.names = FALSE,
+  keep.stat = "n"
+)
+sink()
+
+# visualization
+mods <- c(
+  fit_lmer[c("turnover_controls", "principal_teacher_controls")],
+  fit_spaece["turnover_controls"]
+)
+
+names(mods) <- c("model turnover saeb", "model principal_teacher", "model turnover_spaece")
+
+estimate_hlm <- map2_dfr(
+  mods,
+  names(mods),
+  ~ tidyfit(
+    .x
+  ) %>%
+    mutate(
+      type = .y
+    )
+)
+
+estimate_hlm <- map2_dfr(
+  mods,
+  names(mods),
+  ~ tidyfit(
+    .x
+  ) %>%
+    mutate(
+      type = .y
+    )
+)
+
+plot_model_turnover <- estimate_hlm %>%
+  ggplot(
+    aes(
+      y = term,
+      x = estimate,
+      color = type,
+      group = type
+    )
+  ) +
+  geom_point(
+    position = ggstance::position_dodgev(height = 0.3),
+    size = 2
+  ) +
+  geom_errorbar_tidy +
+  theme(
+    legend.position = "bottom",
+    legend.title = element_blank()
+  ) +
+  geom_vline(xintercept = 0) +
+  scale_y_discrete(
+    limits = c(
+      "turnover_index",
[…]

### 2020-07-02 build/replication/modules/visual_test_scores.R

@@ -1,0 +1,75 @@
+# test scores -------------------------------------------------------------
+saeb_dep <- read_data(
+    "saeb",
+    "saeb_school.rds"
+)
+
+saeb_public <- saeb_school %>%
+  filter(
+    dep != "",
+    !is.na(grade),
+    dep != "particular",
+    year >= 2003
+  ) %>% 
+  mutate(
+    year = as.factor(year)
+  )
+
+saeb_public %>%
+    gg_summary(
+      year,
+      grade_mean
+    )
+saeb_public %>% 
+  ggplot(
+    aes(
+      year, mean_grade_exam, 
+      group = dep, col = dep
+    )
+  ) + 
+  geom_line() +
+  geom_point() +
+  labs(
+    caption = "Average test scores, by administrative level. Includes math and Portuguese test scores."
+  ) + 
+  facet_wrap(
+    grade ~ subject
+  )
+
+saeb_descriptive <- saeb_exam_mun %>% 
+  join_covariate
+
+saeb_descriptive %>% 
+  gg_summary(
+    censo_pop, 
+    mean_grade_exam
+  ) +
+  scale_x_log10()
+
+saeb_descriptive %>% 
+  gg_summary(
+    I(budget_education/censo_pop),
+    mean_grade_exam
+  )
+
+saeb_exam_school %>% 
+  inner_join(
+    censo_school_turnover,
+    by = c('cod_ibge_6', 'year', 'school_id')
+  ) %>% 
+  mutate(
+    state = str_sub(cod_ibge_6, 1, 2)
+  ) %>% 
+  group_by(
+    state,
+    year
+  ) %>% 
+  mutate(
+    turnover_index = turnover_index - mean(turnover_index, na.rm = T),
+    mean_grade_exam = mean_grade_exam - mean(mean_grade_exam, na.rm = T)
+  ) %>% 
+  ungroup() %>% 
+  gg_summary(
+    turnover_index,
+    mean_grade_exam
+  )

### 2020-07-02 build/replication/scripts/modules/import_data.R

@@ -1,0 +1,245 @@
+# data --------------------------------------------------------------------
+# global education indicators
+
+import_data(
+  here("data/saeb"),
+  "exam_mun|hierarchical|exam_school"
+) %>%
+  assign_data
+
+saeb_exam_mun <- read_data(
+  "saeb",
+  "saeb_exam_mun.rds"
+)
+
+# rais
+rais_edu <- fread(
+  here("data/rais/rais_edu.csv.gz")
+) %>% 
+  filter(
+    year >= 2003
+  )
+
+rais_edu_mun <- fread(
+  here('data/rais/rais_edu_mun.csv')
+) %>% 
+  join_covariate() %>% 
+  mutate(
+    state = str_sub(cod_ibge_6, 1, 2)
+  ) %>% 
+  rescale
+
+rais_mun <- fread(
+  here("data/rais/rais_mun.csv.gz")
+)
+
+rais_fit <- rais_edu %>%
+  sample_group(
+    500,
+    cod_ibge_6
+  ) %>% 
+  join_covariate() %>%
+  mutate(
+    state = str_sub(cod_ibge_6, 1, 2),
+    coalition_share = rescale(coalition_share),
+    rais_category = fct_relevel(rais_category, 'teacher')
+  ) %>% 
+  mutate_at(
+    vars(state, election_year, rais_category),
+    as.factor
+  ) %>% 
+  standardize()
+
+# censo escolar
+censo_class <- fread(
+  here("data/censo_escolar/censo_class.csv.gz")
+) %>% 
+  group_by(
+    year,
+    dep,
+    grade_level
+  ) %>%
+  summarise(
+    num_enroll = sum(num_enroll)
+  ) %>% 
+  mutate(
+    freq = num_enroll/sum(num_enroll)
+  )
+
+censo_school <- fread(
+  here('data/censo_escolar/censo_school.csv.gz')
+)
+
+censo_school_turnover <- fread(
+  here("data/censo_escolar/censo_school_turnover.csv.gz")
+)
+
+censo_turnover_ce <- censo_school_turnover %>% 
+  filter(
+    state == "23",
+    grade_level %in% c(2, 5, 9)
+  )
+
+censo_mun_turnover <- fread(
+  here("data/censo_escolar/censo_mun_turnover.csv.gz")
+) %>% 
+  group_by(
+    cod_ibge_6, year
+  ) %>% 
+  summarise_stats(
+    turnover_index
+  ) %>% 
[…]

### 2020-07-02 build/replication/scripts/saeb_hierarchical.R

@@ -1,0 +1,77 @@
+# saeb hierarchical
+saeb_hlm <- fread(
+  here("data/saeb/saeb_hierarchical.csv.gz")
+)
+
+# join school covariates
+censo_school %<>% 
+  filter(
+    dep == "municipal", year >= 2001
+  ) %>%
+  select(
+    state,
+    cod_ibge_6,
+    year,
+    location,
+    school_id,
+    school_name,
+    internet,
+    kitchen,
+    library,
+    lab_info,
+    meal,
+    starts_with('principal'),
+    sewer_grid,
+    toilet,
+    water_grid
+  )
+
+join_cols <- c("state", "cod_ibge_6", "school_id", "year")
+
+saeb_hlm %<>%
+  mutate(state = str_sub(cod_ibge_6, 1, 2)) %>% 
+  rename(school_id = cod_school) %>% 
+  mutate_at(join_cols, as.character) %>% 
+  left_join(
+    censo_school %>%
+      mutate_at(join_cols, as.character) %>% 
+      select(-location),
+    by = join_cols
+  )
+
+# turnover
+censo_school_turnover <- fread(
+  here("data/censo_escolar/censo_school_turnover.csv.gz")
+) %>% 
+  rename(
+    grade = grade_level,
+    n_teacher = n,
+    n_teacher_lag = n_lag
+  )
+
+join_cols <- c(join_cols, "grade")
+
+saeb_hlm %<>%
+  mutate_at(join_cols, as.character) %>%
+  tidylog::left_join(
+    censo_school_turnover %>% mutate_at(join_cols, as.character),
+    by = join_cols
+  )
+
+# create log pop
+saeb_hlm %<>%
+  mutate(
+    censo_log_pop = log(censo_pop)
+  )
+
+# break down teacher wages into quantiles
+saeb_hlm %<>%
+  mutate(
+    saeb_teacher_work_school = if_else(year_working_school_teacher == "", NA_character_, year_working_school_teacher),
+    grade = recode(grade, `4` = "5", `8` = "9")
+  )
+
+saeb_hlm %>% 
+  fwrite_gz(
+    here("data/saeb/saeb_hierarchical.csv")
+  )

### 2021-01-03 source/data/modules/database/generate_career_prior_bureaucracy.R
fix code for pre-carer

@@ -22,8 +22,6 @@ career_prior_to_bureaucracy <- dbGetQuery(
         cbo_02,
         age,
         edu,
-        gender,
-        nat_jur,
         wage,
         work_experience,
         filiado_mun.party,

### 2021-01-04 paper/draft.tex
add paper

\documentclass[12pt,a4paper]{article}
\usepackage{setspace} \onehalfspacing
\usepackage[top = 1in, bottom = 1in, left = 1in, right = 1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{pdflscape}
% \usepackage{hyperref}
\usepackage[hyphens]{url}
\usepackage{enumitem}
\usepackage[table]{xcolor}
\usepackage{authblk}
\usepackage{appendix}
\usepackage{titletoc}
\usepackage{fancyhdr}
\usepackage{hyperref}

\bibliographystyle{apsr}

\newcommand{\note}{\textcolor{blue}}

\title{How to reduce bureaucratic corruption? Unpacking Brazilian anti-corruption audits}
\author{Galileu Kim}
\affil{Princeton University}

\begin{document}
gu

\end{document}

### 2021-01-04 source/descriptive/modules/entry_into_bureaucracy.R
fix debug script

@@ -100,7 +100,7 @@ plot_transition_occupation <- transition_occupation %>%
         aes(
             x = cbo_02,
             y = cbo_02_entry,
-            fill = n
+            fill = round(n/1000, 2)
         )
     ) +
     scale_fill_continuous(
@@ -110,7 +110,7 @@ plot_transition_occupation <- transition_occupation %>%
     labs(
         x = "private sector occupation",
         y = "bureaucracy occupation",
-        caption = "Total"
+        caption = "Total (thousands)"
     )
 
 message("export plot")

### 2021-01-04 source/data/modules/database/globals.R
fix globals

@@ -5,7 +5,7 @@ library(here)
 library(DBI)
 
 debug <- ifelse(
-    Sys.getenv("USER") == "gali",
+    Sys.getenv("USER") == "gali" & dirname(getwd()) != "/media/gali/gali",
     TRUE,
     FALSE
 )

### 2021-01-04 paper/draft.tex
add beginning of paper 3

@@
 
-\title{How to reduce bureaucratic corruption? Unpacking Brazilian anti-corruption audits}
+\title{A partisan affair: Mapping patronage in municipal bureaucracies of Brazil}
 \author{Galileu Kim}
@@
 \begin{document}
-gu
+
+\maketitle
+
+\abstract{How extensive is patronage in Brazil?
+What are the differences in observables between partisan affiliates and non-partisan members?
+Leveraging a novel dataset of partisan affiliation and employment data on every municipal bureaucrat in Brazil, I find that party members are more likely to be overcompensated than their peers, concentrating in areas of executive leadership, while being less educated than their peers.
+In addition, their employment spells tend to be more durable over time, leading to negative consequences.
+These findings provide raise important questions regarding the duration of partisan ties and its consequences for local bureaucracies.}
+
+\newpage
 

### 2021-01-04 source/data/modules/database/generate_partisanship.R
fix generate_partisanship scipt

@@ -12,49 +12,6 @@ source(
 
 message("extract records and join with filiado data")
 
-rais_filiado <- dbGetQuery(
-    "
-    SELECT 
-        rais.cod_ibge_6,
-        rais.year,
-        rais.id_employee,
-        rais.cbo_02,
-        rais.cnae_95,
-        rais.edu,
-        rais.hours,
-        rais.wage,
-        rais.work_experience,
-        rais.contract_type,
-        filiado_mun.filiado_id,
-        filiado_mun.date_start,
-        filiado_mun.date_end,
-        filiado_mun.is_filiado
-    FROM rais
-    LEFT JOIN (SELECT *, 1.0 AS is_filiado FROM filiado_mun) AS filiado_mun
-    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
-    rais.id_employee = filiado_mun.id_employee
-    "
-)
-
-rais_filiado <- tbl(rais_con, "rais") %>%
-    mutate(
-        occupation = str_sub(cbo_02, 1, 1)
-    ) %>%
-    mutate(
-        cbo_group_detail = case_when(
-            occupation == 0 ~ "army",
-            `1` = "executive",
-            `2` = "liberal arts",
-            `3` = "engineer",
-            `4` = "administration",
-            `5` = "services",
-            `6` = "agriculture",
-            `7` = "industry",
-            `8` = "industry",
-            `9` = "maintenance"
-        )
-    )
-
 group_vars <- c(
     "SUBSTR(cbo_02, 1, 1)", "contract_type", "edu"
 )
@@ -102,30 +59,3 @@ rais_filiado %>%
         dir = "summary_tables",
         file = "partisan_dominance.rds"
     )
-
-# message("generate party dummy and compute partisanship by group")
-# rais_filiado_occupation <- rais_filiado %>%
-#     mutate(
-#         occupation = str_sub(cbo_02, 1, 1)
-#     ) %>%
-#     fix_occupation() %>%
-#     fix_edu() %>%
-#     filter(
-#         !is.na(cbo_group) & !is.na(cbo_group_detail)
-#     )
-
-# # ==============================================================================
-# # compute party dominance across different categories
-# # ==============================================================================
-# message("computing dominance...")
-
-
-# partisan_summary <- map(
-#     group_vars,
-#     ~ compute_summary_stats(
-#         data = rais_filiado_occupation,
-#         .group_vars = all_of(.),
[…]

### 2021-01-04 source/data/modules/database/generate_partisanship.R
fix partisanship dir

@@ -56,6 +56,6 @@ rais_filiado <- rais_filiado %>%
 
 rais_filiado %>%
     write_data(
-        dir = "summary_tables",
+        dir = "summary",
         file = "partisan_dominance.rds"
     )

### 2021-01-04 descriptive/build.R
update source files

@@ -1,0 +1,12 @@
+# ==============================================================================
+# descriptive statistics of turnover of party affiliates
+# ==============================================================================
+build_from_source("descriptive", "requirements.R")
+
+module_selection <- c(
+    # "party_membership",
+    "bureau_partisanship",
+    "prior_to_bureaucy",
+    # "turnover_cycles",
+    "partisan_vs_non_partisan"
+)   

### 2021-01-04 descriptive/modules/bureau_partisanship.R

@@ -1,0 +1,141 @@
+# ==============================================================================
+# input: sql database of party affiliation
+# output: assess extent to which there is party dominance
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here("source/descriptive/modules/requirements.R")
+)
+
+message("extract records and join with filiado data")
+
+rais_filiado <- dbGetQuery(
+    "
+    SELECT 
+        rais.cod_ibge_6,
+        rais.year,
+        rais.id_employee,
+        rais.cbo_02,
+        rais.cnae_95,
+        rais.edu,
+        rais.hours,
+        rais.wage,
+        rais.work_experience,
+        rais.contract_type,
+        filiado_mun.filiado_id,
+        filiado_mun.date_start,
+        filiado_mun.date_end,
+        filiado_mun.is_filiado
+    FROM rais
+    LEFT JOIN (SELECT *, 1.0 AS is_filiado FROM filiado_mun) AS filiado_mun
+    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
+    rais.id_employee = filiado_mun.id_employee
+    "
+)
+
+rais_filiado <- tbl(rais_con, "rais") %>%
+    mutate(
+        occupation = str_sub(cbo_02, 1, 1)
+    ) %>%
+    mutate(
+        cbo_group_detail = case_when(
+        occupation == 0 ~ "army",
+        `1` = "executive",
+        `2` = "liberal arts",
+        `3` = "engineer",
+        `4` = "administration",
+        `5` = "services",
+        `6` = "agriculture",
+        `7` = "industry",
+        `8` = "industry",
+        `9` = "maintenance"
+      )
+    )
+
+dbGetQuery(
+    "
+    SELECT
+    edu,
+    AVG(COALESCE(is_filiado, 0)) AS party_dominance
+    FROM rais
+    LEFT JOIN (SELECT *, 1.0 AS is_filiado FROM filiado_mun) AS filiado_mun
+    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
+    rais.id_employee = filiado_mun.id_employee
+    GROUP BY edu
+    "   
+)
+
+message("generate party dummy and compute partisanship by group")
+rais_filiado_occupation <- rais_filiado %>%
+    mutate(
+        occupation = str_sub(cbo_02, 1, 1)
+    ) %>% 
[…]

### 2021-01-04 descriptive/modules/entry_into_bureaucracy.R

@@ -1,0 +1,126 @@
+# ==============================================================================
+# input: sql database of party affiliation
+# output: compare transition matrices for partisans and non-partisans
+# into the bureaucracy
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here::here("source/descriptive/modules/requirements.R")
+)
+
+rais <- tbl(rais_con, "rais")
+filiado <- tbl(rais_con, "filiado_mun")
+bureaucracy_entry <- tbl(rais_con, "rais_bureaucrat_entry")
+
+# ---------------------------------------------------------------------------- #
+message("generate transition matrix")
+
+# extract entry into bureaucracy
+join_cols <- c("id_employee", "cod_ibge_6", "year")
+outcome_cols <- c("cbo_02", "wage")
+rais_selected <- rais %>%
+    select(
+        all_of(
+            c(join_cols, outcome_cols)
+        )
+    )
+
+# generate entry level job for bureaucracy
+bureaucracy_entry_job <- rais_selected %>%
+    inner_join(
+        bureaucracy_entry,
+        on = c("id_employee", "cod_ibge_6", "year")
+    )
+
+# prepare data for join
+bureaucracy_entry_job <- bureaucracy_entry_job %>%
+    rename(
+        year_entry = year,
+        cbo_02_entry = cbo_02,
+        wage_entry = wage
+    ) %>%
+    collect() %>%
+    filter(
+        !is.na(cbo_02_entry)
+    )
+
+# last private sector job
+private_last_job <- dbGetQuery(
+    "
+    SELECT 
+    rais.id_employee,
+    rais.year,
+    rais.cod_ibge_6,
+    rais.cbo_02,
+    rais.wage
+    FROM rais
+    INNER JOIN rais_bureaucrat_entry
+    ON rais.id_employee = rais_bureaucrat_entry.id_employee
+    AND rais.year <= rais_bureaucrat_entry.year
+    WHERE nat_jur != 1031 AND cbo_02 IS NOT NULL
+    GROUP BY rais.id_employee
+    HAVING rais.year = MAX(rais.year)
+    "
+)
+
+private_last_job <- private_last_job %>%
+    group_by(id_employee) %>%
+    ungroup()
+
+# join at the individual level last private sector job and bureaucracy job
+transition_private_bureaucracy <- private_last_job %>%
+    left_join(
[…]

### 2021-01-04 descriptive/modules/globals.R

@@ -1,0 +1,22 @@
+library(tidyverse)
+library(dbplyr)
+library(data.table)
+library(here)
+library(DBI)
+
+debug <- ifelse(
+    Sys.getenv("USER") == "gali" & dirname(getwd()) != "/media/gali/gali",
+    TRUE,
+    FALSE
+)
+
+data_path <- "data/clean/"
+source_path <- "source/modules/database/"
+rais_sql <- here(
+  sprintf(
+    "data/clean/database/rais%s.sqlite3",
+    ifelse(isTRUE(debug), "_sample", "")
+  )
+)
+
+rais_con <- DBI::dbConnect(RSQLite::SQLite(), rais_sql)

### 2021-01-04 descriptive/modules/partisan_vs_non_partisan.R

@@ -1,0 +1,77 @@
+# ==============================================================================
+# input: rais municipal sql tables
+# output: descriptive statistics on partisan vs. non-partisan
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here("source/descriptive/modules/requirements.R")
+)
+
+# ==============================================================================
+# load-in data
+# ==============================================================================
+rais_mun_tables <- c(
+    partisan = "rais_mun_partisan",
+    non_partisan = "rais_mun_non_partisan"
+)
+
+rais_filiado <- rais_mun_tables %>%
+map_dfr(
+        ~ tbl(rais_con, .) %>%
+            mutate(year = as.integer(year)) %>%
+            collect(),
+        .id = "is_partisan"
+    )
+
+outcome_vars <- c(
+    Age = "age", 
+    Education = "edu",
+    Dismissals = "fired",
+    Hires = "hired",
+    Experience = "work_experience",
+    Wage = "wage"
+)
+
+# # compute statistics per year
+# rais_filiado_summary <- rais_filiado %>%
+#     group_by(is_partisan, year) %>%
+#     summarise(
+#         across(
+#             c("age", "edu", "fired", "hired", "work_experience"),
+#             mean, na.rm = TRUE
+#             ),
+#             wage = median(wage, na.rm = TRUE)
+#     )
+
+plot_descriptive <- map2(
+        outcome_vars,
+        names(outcome_vars),
+        ~ ggplot(
+            aes_string("year", .x, color = "is_partisan"),
+            data = rais_filiado
+        ) +
+        geom_smooth() +
+        guides(color = guide_legend("Partisanship")) +
+        labs(x = "Year", y = .y)
+    )
+
+message("exporting plots")
+plot_filenames <- generate_plot_filenames(
+    "paper/figures/active_employee/",
+    outcome_vars,
+    debug
+)
+
+pmap(
+    list(
+        plot = plot_descriptive,
+        filename = plot_filenames
+    ),
+    ggsave
+)
+
+message("module completed: generate descriptive statistics for active employees.")
+

### 2021-01-04 descriptive/modules/party_membership.R

@@ -1,0 +1,80 @@
+# ==============================================================================
+# input: sql database of party affiliation
+# output: compare membership cycles for workers and non-workers
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here("source/descriptive/modules/requirements.R")
+)
+
+# ==============================================================================
+# generate descriptive statistics of party members
+# ==============================================================================
+rais_sql <- tbl(rais_con, "rais") %>%
+    filter_municipal()
+    
+filiado_sql <- tbl(rais_con, "filiado_mun")
+
+rais_tables <- c(
+    partisan = "rais_mun_partisan",
+    non_partisan = "rais_mun_non_partisan"
+)
+
+rais_filiado <- map_dfr(
+    rais_tables,
+    ~ dbGetQuery(sprintf("SELECT * FROM %s", .)),
+    .id = "is_filiado"
+)
+
+# create entries and exits per year
+# segmented by employment
+filiado_unemployed <- filiado_sql %>%
+    anti_join(
+        rais_sql %>%
+            select(cod_ibge_6, id_employee),
+        by = c("cod_ibge_6", "id_employee")
+    ) %>%
+    collect()
+
+filiado_employed <- filiado_sql %>%
+    inner_join(
+        rais_sql %>%
+            select(cod_ibge_6, id_employee),
+        by = c("cod_ibge_6", "id_employee")
+    ) %>%
+    collect() %>%
+    distinct(filiado_id, .keep_all = TRUE)
+
+# compare duration of party membership between employed vs. unemployed
+filiado_turnover <- list(
+    filiado_employed,
+    filiado_unemployed
+) %>%
+    set_names(c("employed", "unemployed")) %>%
+    map_dfr(
+        . %>%
+            convert_to_date() %>%
+            pivot_longer(
+                c(date_start, date_end),
+                names_prefix = "date_",
+                names_to = "date"
+            ) %>%
+            count(date, value),
+        .id = "is_employed"
+    )
+
+# verify when data for membership termination begins
+filiado_turnover %>%
+    gg_point(
+        aes(x = value, y = n),
+        size = 2
+    ) +
+    plot_mandate_year() +
+    facet_wrap(. ~ date + is_employed) +
+    coord_cartesian(ylim = c(0, 50)) +
+    scale_x_date(limit = as.Date(c("2003-01-01", "2019-01-01")))
+
+# compare membership spells

### 2021-01-04 descriptive/modules/prior_to_bureaucracy.R

@@ -1,0 +1,104 @@
+# ==============================================================================
+# input: sql database of party affiliation
+# output: compare differential observables for partisans and non-partisans
+# before they enter the bureaucracy
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here("source/descriptive/modules/requirements.R")
+)
+
+filiado <- tbl(rais_con, "filiado_mun")
+career_filiado <- tbl(rais_con, "career_pre_bureaucracy_mun") %>%
+    collect()
+
+# # ---------------------------------------------------------------------------- #
+# # compute summary statistics for partisans and non-partisans
+# message("computing summary statistics...")
+
+# # note that the majority of partisans become party members after
+# # joining the bureau
+# plot_breakdown <- career_filiado %>%
+#     group_by(year, is_partisan) %>%
+#     summarise(n = sum(n)) %>%
+#     gg_point_line(
+#         aes(year, n/1e6, color = is_partisan)
+#     ) +
+#     scale_color_discrete(
+#         name = "Partisanship",
+#         labels = c(
+#             "pre_partisan" = "pre-partisan",
+#             "post_partisan" = "post-partisan",
+#             "non_partisan" = "non-partisan"
+#         )
+#     ) +
+#     labs(x = "Year", y = "Total (millions)")
+
+# group_vars <- c("year", "is_partisan")
+
+# career_filiado_mean <- career_filiado %>%
+#     compute_mean(
+#         all_of(group_vars),
+#         c(age, work_experience, edu)
+#     )
+
+# career_filiado_median <- career_filiado %>%
+#     group_by(
+#         across(all_of(group_vars))
+#     ) %>%
+#     compute_median(wage)
+
+# career_filiado_summary <- career_filiado_mean %>%
+#     left_join(
+#         career_filiado_median,
+#         by = group_vars
+#     ) %>%
+#     filter(!is.na(is_partisan))
+
+# message("complete!")
+
+# ---------------------------------------------------------------------------- #
+message("generating descriptive plots")
+
+vars_to_plot <- c("age", "work_experience", "edu") %>%
+    sprintf(fmt = "%s_mean", .) %>%
+    c("median_wage", "n")
+
+label_vars <- c("Age", "Work Experience", "Education", "Wage", "Total (million)")
+
+plot_summary <- map2(
+    vars_to_plot,
+    label_vars,
[…]

### 2021-01-04 descriptive/modules/requirements.R

@@ -1,0 +1,689 @@
+here_data <- partial(
+  here,
+  data_path
+)
+
+here_source <- partial(
+  here,
+  source_path
+)
+
+fread <- purrr::partial(
+  data.table::fread,
+  nThread = parallel::detectCores(),
+  nrows = nrows
+)
+
+dbGetQuery <- purrr::partial(
+  DBI::dbGetQuery,
+  conn = rais_con
+)
+
+nrows <- if (isTRUE(debug)) 1e4 else Inf
+
+# ==============================================================================
+# data io
+# ==============================================================================
+list_files <- function(path, pattern) {
+  files <- map2(
+    path,
+    pattern,
+    list.files
+  ) %>%
+    flatten_chr()
+
+  return(files)
+}
+
+write_data <- function(dir, filename, data){
+  file_extension <- str_extract(filename, "(?<=\\.)[a-z]+")
+
+  switch(
+    file_extension,
+    csv = fwrite(
+      data,
+      here(dir, filename)
+    ),
+    rds = write_rds(
+      data,
+      here(dir, filename)
+    )
+  )
+}
+
+# ==============================================================================
+# rais wrangling
+# ==============================================================================
+is_municipal <- rlang::quo(nat_jur == 1031)
+is_bureau <- rlang::quo(nat_jur == 1031 & occupation == 1)
+is_partisan <- rlang::quo(
+  year >= year_date_start & (year <= year_date_end | is.na(year_date_end))
+)
+
+year_to_char <- function(data) {
+  data <- data %>%
+    mutate_at(
+      vars(year),
+      as.character
+    )
+
+  return(data)
+}
+
+convert_to_date <- function(data) {
+  data %>%
+    mutate(
+      across(starts_with("date"), lubridate::ymd)
+    )
+}
+
+fix_edu <- function(data){
+  data %>%
+  mutate(
+            edu_category = case_when(
+                edu <= 2 ~ "illiterate",
+                between(edu, 3, 4) ~ "lower school",
+                between(edu, 5, 6) ~ "middle school",
+                between(edu, 7, 8) ~ "high school",
+                edu >= 9 ~ "higher education"
+            )
+  )
+}
+
+fix_wage <- function(data) {
+  data %>%
[…]

### 2021-01-04 descriptive/modules/turnover_cycles.R

@@ -1,0 +1,66 @@
+# ==============================================================================
+# input: sql database
+# output: descriptive statistics of turnover cycles in bureaucracy
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here::here("source/descriptive/modules/requirements.R")
+)
+
+rais <- tbl(rais_con, "rais")
+filiado <- tbl(rais_con, "filiado_mun")
+
+rais_tables <- c(
+    partisan = "rais_mun_partisan",
+    non_partisan = "rais_mun_non_partisan"
+)
+
+rais_filiado <- map_dfr(
+    rais_tables,
+    ~ dbGetQuery(sprintf("SELECT * FROM %s", .)),
+    .id = "is_filiado"
+)
+
+message("generate plots")
+vars_to_plot <- c("hired", "fired", "work_experience")
+plot_turnover <- map(
+    vars_to_plot,
+    ~ gg_summary(
+        x = "year",
+        y = .,
+        data = rais_filiado,
+        smooth = FALSE,
+        color = !!sym("is_filiado")
+    ) +
+        ggtitle(.)
+)
+
+rais_filiado %>%
+    ggplot(
+        aes(year, hired, color = is_filiado)
+    ) +
+    stat_summary_bin(
+        fun = "mean",
+        geom = "point"
+    )
+
+path_to_figs <- ifelse(
+    isTRUE(debug), "paper/figures/turnover/sample/", "paper/figures/turnover/"
+)
+
+file_names <- sprintf(
+    "plot_%s.pdf", vars_to_plot
+)
+
+message("save plots")
+path_out <- here(path_to_figs, file_names)
+pwalk(
+    list(
+        filename = path_out,
+        plot = plot_turnover
+    ),
+    save_plot
+)

### 2021-01-04 source/data/modules/database/create_rais_sql.R

@@ -18,7 +18,10 @@ rais_files <- list.files(
 )
 
 log_file <- here("log/log_create_sql.txt")
-debug <- FALSE
+
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
 
 source(
   here("source/data/modules/database/requirements.R")
@@ -115,6 +118,13 @@ dbExecute(
   "
   CREATE INDEX IF NOT EXISTS natureza_juridica_idx ON
   rais(nat_jur)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS id_employee_idx ON
+  rais(id_employee)
   "
 )
 

### 2021-01-04 utils/diagnostics_linkage.R

@@ -1,0 +1,65 @@
+summarise_duplicates <- function(data) {
+    # generates proportion of entries duplicated
+    # and the density of duplicated entries
+    # note it expects the var name to be n
+    data_summary <- data[
+        ,
+        .(
+            prop_duplicated = mean(n > 1),
+            density_duplicated = sum(n > 1) / sum(n)
+        )
+    ]
+
+    return(data_summary)
+}
+
+count_dt <- function(data, ...) {
+    # generates the count of values
+    var <- substitute(list(...))
+
+    data_count <- data[
+        ,
+        .(n = .N),
+        by = eval(var)
+    ]
+
+    return(data_count)
+}
+
+count_unique_by_group <- function(data, col, by) {
+    col <- substitute(unlist(col, recursive = FALSE))
+    by <- substitute(by)
+
+    data_count_unique <- data[
+        ,
+        .(n = uniqueN(eval(col))),
+        by = eval(by)
+    ]
+
+    return(data_count_unique)
+}
+
+generate_diagnostics_rais <- function(data) {
+    # generate a diagnostics table from rais
+    year <- unique(data[["year"]])
+
+    # sample 10 percent of the data
+    data_sample <- data[sample(1:nrow(data), nrow(data)/10, replace = FALSE)]
+
+    data_diagnostics_name <- data_sample %>%
+        setkey(name) %>%
+        count_dt(name) %>%
+        summarise_duplicates()
+
+    data_diagnostics_cpf <- data_sample %>%
+        setkey(cpf) %>%
+        count_unique_by_group(name, by = .(cpf)) %>%
+        summarise_duplicates()
+
+    data_diagnostics <- list(data_diagnostics_name, data_diagnostics_cpf) %>%
+        modify(~ mutate(., year = year)) %>%
+        set_names(c("name", "cpf")) %>%
+        bind_rows(.id = "type")
+
+    return(data_diagnostics)
+}

### 2021-01-04 utils/io.R

@@ -1,0 +1,40 @@
+reset_env <- function(init_env){
+  final_env <- ls(.GlobalEnv)
+
+  rm(
+    envir = .GlobalEnv,
+    list = setdiff(final_env, init_env)
+  )
+
+  gc()
+}
+
+build_repo <- function(module){
+  repo <- here("data", "clean", module)
+
+  if(!dir.exists(repo)){
+    dir.create(repo)
+  }else{
+    unlink(repo, recursive = T)
+    dir.create(repo)
+    }
+  }
+
+run_module <- function(module, domain) {
+  print(
+    paste0("running module ", module, "...")
+  )
+
+  build_repo(module)
+
+  source_file <- paste0("run_", module, ".R")
+  path <- here("source", "modules", domain, source_file)
+
+  init_env <- ls(.GlobalEnv)
+  source(path)
+  reset_env(init_env)
+
+  print(
+    paste("module", module, "complete!")
+  )
+}

### 2021-01-04 utils/matching.R

@@ -1,0 +1,71 @@
+# aux funs for matching
+
+blocked_fastLink <- function(year, state, df_rais, df_filiados) {
+    t <- year
+    s <- state
+
+    rais <- df_rais %>%
+        filter(
+            year == t,
+            state == s
+        )
+
+    filiados <- df_filiados %>%
+        filter(
+            (year_start < t) & (year_cancel > t) &
+                state == s
+        )
+
+    rais_filiados_link <- fastLink(
+        dfA = rais,
+        dfB = filiados,
+        varnames = c("first_name", "last_name", "middle_name"),
+        stringdist.match = c("first_name", "last_name", "middle_name"),
+        n.cores = 1
+    )
+
+    return(rais_filiados_link)
+}
+
+run_diagnostics_fastLink <- function(fastLink, rais, filiados) {
+    matched_fastLink <- getMatches(
+        dfA = rais,
+        dfB = filiados,
+        fl.out = fastLink,
+        combine.dfs = F
+    )
+
+    # match rate by first name
+    first_name_matches <- matched_fastLink %>%
+        reduce(
+            inner_join,
+            by = c("first_name", "last_name")
+        )
+
+    exact_matches <-
+
+    if (number_of_first_name_matches <= 40) {
+        stop("match rate for first names less than 80 percent.")
+    } else{
+        print(sprintf("match rate is %s percent.",
+        number_of_first_name_matches/50)
+        )
+    }
+
+    # how many duplicated id_employees are there for each electoral title?
+    proportion_duplicated_rais_filiados <- setDT(matched_fastLink)[
+        , count := uniqueN(elec_title),
+        by = id_employee
+    ] %>%
+        summarise(
+            proportion_duplicated = mean(count > 1)
+        ) %>%
+        pull()
+
+    if (proportion_duplicated_rais_filiados > 20) {
+        stop("duplication rate is higher than 20 percent.")
+    } else{
+        print(sprintf("duplication rate is %s"),
+        proportion_duplicated_rais_filiados)
+    }
+}

### 2021-01-04 utils/panel.R

@@ -1,0 +1,89 @@
+year_to_char <- function(data) {
+  data <- data %>%
+    mutate_at(
+      vars(year),
+      as.character
+    )
+
+  return(data)
+}
+
+fix_wage <- function(data) {
+  data %>%
+    mutate(
+      wage = wage * 0.573395 / (cpi / 100)
+    ) %>%
+    select(-cpi)
+}
+
+fix_contract <- function(data) {
+  data %>%
+    mutate(
+      contract_type = case_when(
+        contract <= 25 ~ "regular",
+        between(contract, 30, 35) ~ "career",
+        between(contract, 40, 50) ~ "temporary",
+        TRUE ~ "other"
+      )
+    )
+}
+
+fix_occupation <- function(data) {
+  data %>%
+    mutate(
+      cbo_group = recode(
+        occupation,
+        `1` = "bureaucrat_high",
+        `2` = "frontline_high",
+        `3` = "frontline_high",
+        `4` = "bureaucrat_low",
+        `5` = "frontline_low",
+        .default = "other"
+      ),
+      cbo_group_detail = recode(
+        occupation,
+        `0` = "army",
+        `1` = "executive",
+        `2` = "liberal arts",
+        `3` = "engineer",
+        `4` = "administration",
+        `5` = "services",
+        `6` = "agriculture",
+        `7` = "industry",
+        `8` = "industry",
+        `9` = "maintenance"
+      )
+    )
+}
+
+trim_rais <- function(data, ...) {
+  data %>%
+    transmute(
+      id_employee,
+      name = nome,
+      year,
+      cod_ibge_6 = municipio,
+      cbo_02 = cbo2002,
+      age,
+      edu,
+      gender = genero,
+      race = racacor,
+      work_experience = tempoempr,
+      wage = vlremmedia,
+      hired = if_else(tipoadmissao == 1 | tipoadmissao == 2, 1, 0),
+      fired = if_else(causadesli == 10 | causadesli == 11, 1, 0),
+      departure = if_else(causadesli == 20 | causadesli == 21, 1, 0),
+      date_admission,
+      date_separation,
+      contract = tipovinculo,
+      nat_jur = naturjur,
+      municipal = if_else(naturjur == 1031, 1, 0),
+      outcome = recode(fired, `0` = 0, `10` = 2, `11` = 2, .default = 1),
+      state = substr(cod_ibge_6, 1, 2),
[…]

### 2021-01-04 utils/preprocess.R

@@ -1,0 +1,116 @@
+extract_unique_records <- function(data, ...) {
+    data_unique <- data %>%
+        distinct(
+            id_employee,
+            name
+        ) %>%
+        create_split_name()
+
+    return(data_unique)
+}
+
+clean_name <- function(name) {
+    clean_name <- str_replace_all(name, "[^[:alpha:] ]", "") %>%
+        str_to_lower() %>%
+        iconv(to = "ASCII//TRANSLIT")
+
+    return(clean_name)
+}
+
+clean_filiados <- function(data, ...) {
+    clean_data <- data %>%
+        transmute(
+            cod_ibge_6,
+            electoral_title,
+            name = clean_name(member_name)
+            # party,
+            # date_start,
+            # date_end,
+            # date_cancel,
+            # ...
+        ) %>%
+        mutate_all(
+            as.character
+        )
+    # create_split_name() %>%
+    # extract_year_from_dates() %>%
+    # mutate(
+    #     state = str_sub(cod_ibge_6, 1, 2)
+    # )
+
+    return(clean_data)
+}
+
+dedupe_data <- function(data, vars) {
+    unique_data <- unique(data, by = vars)
+
+    return(unique_data)
+}
+
+create_split_name <- function(data, var = name) {
+    data_with_split_names <- data %>%
+        mutate(
+            {{ var }} := clean_name({{ var }}),
+            split_names = str_match(
+                {{ var }}, "(^\\w+)\\s*(.+)\\s*\\b(\\w+)$"
+            ),
+            first_name = split_names[, 2],
+            middle_name = split_names[, 3],
+            last_name = split_names[, 4]
+        ) %>%
+        select(-split_names)
+
+    return(data_with_split_names)
+}
+
+extract_first_name <- function(string){
+    first_name <- str_subset(string, "(^[a-z]+)")
+}
+
+extract_year_from_dates <- function(data) {
+    data_with_years <- data %>%
+        mutate(
+            across(
+                c(starts_with("date")),
+                extract_year,
+                .names = "year_{col}"
+            ),
+            year_termination = pmax(
+                year_date_end, year_date_cancel,
+                na.rm = T
+            ) %>%
+                replace_na(2019)
+        ) %>%
+        rename_with(
+            ~ str_replace(., "year_date", "year"),
+            starts_with("year_date")
+        )
+
+    return(data_with_years)
+}
+
[…]

### 2021-01-04 utils/preprocess_rais.R

@@ -1,0 +1,84 @@
+read_7z <- function(file_path, year, select = NULL, dest_dir = tempdir()) {
+    # extracts file from id folder
+    # into a dest_folder, silently
+    # returns data, cleans up the temp file
+    dest_dir_temp <- sprintf(
+        "%s/%s", dest_dir, "temp"
+    )
+
+    extraction_command <- sprintf(
+        "7za e -aoa -o%s %s", dest_dir_temp, file_path
+    )
+
+    system(extraction_command)
+
+    extracted_file_path <- list.files(
+        dest_dir_temp,
+        full.names = T
+    )
+
+    # separator <- if(year <= 2009) ";" else "auto"
+
+    data <- extracted_file_path %>%
+        map_dfr(
+            ~ fread(
+                .,
+                colClasses = "character",
+                select = select,
+                encoding = "Latin-1",
+                sep = ";"
+            )
+        )
+
+    unlink(dest_dir_temp, recursive = T)
+
+    return(data)
+}
+
+extract_id_file_names <- function(year, debug) {
+    print(
+        sprintf("start producing hash for year %s", year)
+    )
+
+    id_file_path <- list.files(
+        sprintf("%s/%s/", id_path, year),
+        pattern = "7z$",
+        full.names = T
+    )
+
+    if (isTRUE(debug)) {
+          id_file_path <- sample(id_file_path, 1)
+      } else {
+          id_file_path <- id_file_path
+      }
+
+    return(id_file_path)
+}
+
+clean_name <- function(name) {
+    clean_name <- str_replace_all(name, "[^[:alpha:] ]", "") %>%
+        str_to_lower() %>%
+        iconv(to = "ASCII//TRANSLIT")
+
+    return(clean_name)
+}
+
+extract_unique_cpf <- function(data, key = "cpf") {
+    unique_data <- data[!c("0", "99")] %>%
+        unique(
+            by = key
+        )
+
+    return(unique_data)
+}
+
+write_out_hash <- function(hash, filename) {
+    tibble(
+        cpf = keys(hash),
+        name = values(hash, USE.NAMES = F)
+    ) %>%
+        as.data.table() %>%
+        fwrite(
+            sprintf(here("data/clean/id/rais_hash/%s"), filename)
+        )
+}

### 2021-01-04 utils/record_linkage.R

@@ -1,0 +1,147 @@
+group_nest_dt <- function(dt, ..., .key = "data") {
+    stopifnot(is.data.table(dt))
+    by <- substitute(list(...))
+
+    dt <- dt[, list(list(.SD)), by = eval(by)]
+    setnames(dt, old = "V1", new = .key)
+    dt
+}
+
+unnest_dt <- function(dt, col, id) {
+    stopifnot(is.data.table(dt))
+
+    by <- substitute(id)
+    col <- substitute(unlist(col, recursive = FALSE))
+    dt[, eval(col), by = eval(by)]
+}
+
+filter_group_by_size <- function(data, ..., n = 1) {
+    by <- substitute(list(...))
+
+    data_filter <- data[
+        ,
+        if (.N == n) .SD,
+        by = eval(by)
+    ]
+
+    return(data_filter)
+}
+
+nested_join <- function(nested_data) {
+    data <- record_rais_filiados_list %>%
+        modify(
+            ~ mutate(
+                .,
+                kmer = substr(name, 1, 3),
+                first_name = str_extract(name, "^[a-z]+")
+            ) %>%
+                setkey(first_name)
+        )
+
+    cat("nest and join rais and filiados data")
+    record_rais_filiados_nested <- record_rais_filiados %>%
+        modify(
+            # ~ filter_group_by_size(., n = 1, name) %>%
+            ~ filter_group_by_size(., name) %>%
+                group_nest_dt(
+                    .,
+                    year, cod_ibge_6, kmer,
+                    .key = "nested_data"
+                ) %>%
+                mutate(
+                    nested_data = map(nested_data, ~ setkey(., name))
+                )
+        )
+}
+
+inner_merge <- function(x, y) {
+    merge(x, y, all = FALSE)
+}
+
+reset_env <- function(init_env) {
+    final_env <- ls(.GlobalEnv)
+
+    rm(
+        envir = .GlobalEnv,
+        list = setdiff(final_env, init_env)
+    )
+
+    gc()
+}
+
+diagnose_duplicates <- function(data, group) {
+    grouping <- substitute(group)
+
+    diagnostic <- data[
+        ,
+        .(n = .N),
+        by = eval(grouping)
+    ][
+        ,
+        .(
+            number_duplicated = sum(n > 1),
+            density_duplicated = sum(n[n > 1]) / sum(n),
+            total_records = sum(n)
+        )
+    ]
+
+    return(diagnostic)
+}
[…]

### 2021-01-05 descriptive/modules/requirements.R
modify write_data function

@@ -38,6 +38,13 @@ list_files <- function(path, pattern) {
 write_data <- function(dir, filename, data){
   file_extension <- str_extract(filename, "(?<=\\.)[a-z]+")
 
+    if(!dir.exists(dir)){
+    dir.create(dir, recursive = TRUE)
+  }else{
+    unlink(dir, recursive = T)
+    dir.create(dir, recursive = TRUE)
+    }
+
   switch(
     file_extension,
     csv = fwrite(

### 2021-01-05 source/data/modules/database/generate_partisanship.R

@@ -56,6 +56,6 @@ rais_filiado <- rais_filiado %>%
 
 rais_filiado %>%
     write_data(
-        dir = "summary",
+        dir = "data/clean/summary",
         file = "partisan_dominance.rds"
     )

### 2021-01-05 source/data/modules/database/globals.R

@@ -5,7 +5,7 @@ library(here)
 library(DBI)
 
 debug <- ifelse(
-    Sys.getenv("USER") == "gali" & dirname(getwd()) != "/media/gali/gali",
+    Sys.getenv("USER") == "gali" & dirname(getwd()) != "/hdd",
     TRUE,
     FALSE
 )

### 2021-01-05 descriptive/modules/globals.R
generate new scripts for index

@@ -5,7 +5,7 @@ library(here)
 library(DBI)
 
 debug <- ifelse(
-    Sys.getenv("USER") == "gali" & dirname(getwd()) != "/media/gali/gali",
+    Sys.getenv("USER") == "gali" & dirname(here()) != "/hdd",
     TRUE,
     FALSE
 )

### 2021-01-05 source/data/modules/database/create_rais_sql.R

@@ -112,30 +112,6 @@ for (file in rais_files) {
 }
 
 # ==============================================================================
-# generate indexes to spee dd up query
-# ==============================================================================
-dbExecute(
-  "
-  CREATE INDEX IF NOT EXISTS natureza_juridica_idx ON
-  rais(nat_jur)
-  "
-)
-
-dbExecute(
-  "
-  CREATE INDEX IF NOT EXISTS id_employee_idx ON
-  rais(id_employee)
-  "
-)
-
-dbExecute(
-  "
-  CREATE INDEX IF NOT EXISTS year_idx ON
-  rais(year)
-  "
-)
-
-# ==============================================================================
 # generate sample sql file
 # ==============================================================================
 rais_sample_con <- DBI::dbConnect(

### 2021-01-05 source/data/modules/database/generate_indexes.R

@@ -1,0 +1,60 @@
+
+# ==============================================================================
+# generate indexes to speed up query
+# ==============================================================================
+source(
+    here::here("source/data/modules/database/globals.R")
+)
+
+source(
+    here("source/data/modules/database/requirements.R")
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS natureza_juridica_idx ON
+  rais(nat_jur)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS id_employee_idx ON
+  rais(id_employee)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS mun_idx ON
+  rais(cod_ibge_6)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS year_idx ON
+  rais(year)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS entry_id_employee_idx ON
+  rais_bureaucrat_entry(id_employee)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS entry_year_idx ON
+  rais_bureaucrat_entry(id_employee)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS entry_mun_idx ON
+  rais_bureaucrat_entry(cod_ibge_6)
+  "
+)

### 2021-01-05 source/descriptive/modules/globals.R

@@ -5,7 +5,7 @@ library(here)
 library(DBI)
 
 debug <- ifelse(
-    Sys.getenv("USER") == "gali" & dirname(getwd()) != "/media/gali/gali",
+    Sys.getenv("USER") == "gali" & dirname(Sys.getenv("PWD")) != "/hdd",
     TRUE,
     FALSE
 )

### 2021-01-05 descriptive/modules/requirements.R
finalize new code for generating partisanship data

@@ -38,12 +38,9 @@ list_files <- function(path, pattern) {
 write_data <- function(dir, filename, data){
   file_extension <- str_extract(filename, "(?<=\\.)[a-z]+")
 
-    if(!dir.exists(dir)){
+  if(!dir.exists(dir)){
     dir.create(dir, recursive = TRUE)
-  }else{
-    unlink(dir, recursive = T)
-    dir.create(dir, recursive = TRUE)
-    }
+  }
 
   switch(
     file_extension,
@@ -588,7 +585,20 @@ gg_bar <- function(data, x, y){
       aes(reorder(!!sym(x), {{y}}), {{y}})
     ) +
     geom_col() +
-    labs(x = x)
+    labs(x = x) +
+    coord_flip()
+
+  return(plot)
+}
+
+gg_boxplot <- function(data, x, y, xlab){
+  plot <- data %>%
+    ggplot(
+      aes(reorder(!!sym(x), {{y}}), {{y}})
+    ) +
+    geom_boxplot() +
+    labs(x = xlab) +
+    coord_flip()
 
   return(plot)
 }
@@ -682,12 +692,9 @@ plot_mandate_year <- function(years = seq(2005, 2013, 4)) {
 save_plot <- function(plot, filename){
   dir <- dirname(filename)
 
-    if(!dir.exists(dir)){
+  if(!dir.exists(dir)){
     dir.create(dir, recursive = TRUE)
-  }else{
-    unlink(dir, recursive = T)
-    dir.create(dir, recursive = TRUE)
-    }
+  }
 
     ggsave(
       filename,

### 2021-01-05 source/data/modules/database/generate_indexes.R

@@ -38,6 +38,8 @@ dbExecute(
   "
 )
 
+# rais entry
+
 dbExecute(
   "
   CREATE INDEX IF NOT EXISTS entry_id_employee_idx ON
@@ -48,7 +50,7 @@ dbExecute(
 dbExecute(
   "
   CREATE INDEX IF NOT EXISTS entry_year_idx ON
-  rais_bureaucrat_entry(id_employee)
+  rais_bureaucrat_entry(year)
   "
 )
 
@@ -56,5 +58,22 @@ dbExecute(
   "
   CREATE INDEX IF NOT EXISTS entry_mun_idx ON
   rais_bureaucrat_entry(cod_ibge_6)
+  "
+)
+
+# filiado_mun
+
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS filiado_id_employee_idx ON
+  filiado_mun(id_employee)
+  "
+)
+
+dbExecute(
+  "
+  CREATE INDEX IF NOT EXISTS filiado_mun_idx ON
+  filiado_mun(cod_ibge_6)
   "
 )

### 2021-01-05 descriptive/build.R
code refactoring to eliminate redundancy

@@ -1,12 +1,0 @@
-# ==============================================================================
-# descriptive statistics of turnover of party affiliates
-# ==============================================================================
-build_from_source("descriptive", "requirements.R")
-
-module_selection <- c(
-    # "party_membership",
-    "bureau_partisanship",
-    "prior_to_bureaucy",
-    # "turnover_cycles",
-    "partisan_vs_non_partisan"
-)   

### 2021-01-05 descriptive/modules/bureau_partisanship.R

@@ -1,63 +1,0 @@
-# ==============================================================================
-# input: sql database of party affiliation
-# output: assess extent to which there is party dominance
-# ==============================================================================
-source(
-    here::here("source/descriptive/modules/globals.R")
-)
-
-source(
-    here("source/descriptive/modules/requirements.R")
-)
-
-partisan_summary <- read_rds(
-    here("data/clean/summary/partisan_dominance.rds")
-)
-
-# ==============================================================================
-# compute party dominance across different categories
-# ==============================================================================
-message("computing dominance...")
-
-group_vars <- c(
-    "cbo_group", "contract_type", "edu_category"
-)
-
-labels <- c(
-    "occupation category", "type of contract", "education"
-)
-
-message("printing out plots")
-
-plot_partisan <- pmap(
-    list(
-        data = partisan_summary,
-        x = group_vars,
-        xlab = labels
-    ),
-    gg_boxplot,
-    y = partisanship
-)
-
-# ==============================================================================
-# exporting plots and data
-# ==============================================================================
-message("exporting plots")
-path_to_figs <- ifelse(
-    isTRUE(debug),
-    "paper/figures/partisanship/sample/",
-    "paper/figures/partisanship/"
-)
-
-export_plots <- list(
-    filename = sprintf(
-        here(path_to_figs, "plot_partisanship_by_%s.pdf"),
-        group_vars
-        ),
-    plot = plot_partisan
-)
-
-pwalk(
-    export_plots,
-    save_plot
-)

### 2021-01-05 descriptive/modules/entry_into_bureaucracy.R

@@ -1,126 +1,0 @@
-# ==============================================================================
-# input: sql database of party affiliation
-# output: compare transition matrices for partisans and non-partisans
-# into the bureaucracy
-# ==============================================================================
-source(
-    here::here("source/descriptive/modules/globals.R")
-)
-
-source(
-    here::here("source/descriptive/modules/requirements.R")
-)
-
-rais <- tbl(rais_con, "rais")
-filiado <- tbl(rais_con, "filiado_mun")
-bureaucracy_entry <- tbl(rais_con, "rais_bureaucrat_entry")
-
-# ---------------------------------------------------------------------------- #
-message("generate transition matrix")
-
-# extract entry into bureaucracy
-join_cols <- c("id_employee", "cod_ibge_6", "year")
-outcome_cols <- c("cbo_02", "wage")
-rais_selected <- rais %>%
-    select(
-        all_of(
-            c(join_cols, outcome_cols)
-        )
-    )
-
-# generate entry level job for bureaucracy
-bureaucracy_entry_job <- rais_selected %>%
-    inner_join(
-        bureaucracy_entry,
-        on = c("id_employee", "cod_ibge_6", "year")
-    )
-
-# prepare data for join
-bureaucracy_entry_job <- bureaucracy_entry_job %>%
-    rename(
-        year_entry = year,
-        cbo_02_entry = cbo_02,
-        wage_entry = wage
-    ) %>%
-    collect() %>%
-    filter(
-        !is.na(cbo_02_entry)
-    )
-
-# last private sector job
-private_last_job <- dbGetQuery(
-    "
-    SELECT 
-    rais.id_employee,
-    rais.year,
-    rais.cod_ibge_6,
-    rais.cbo_02,
-    rais.wage
-    FROM rais
-    INNER JOIN rais_bureaucrat_entry
-    ON rais.id_employee = rais_bureaucrat_entry.id_employee
-    AND rais.year <= rais_bureaucrat_entry.year
-    WHERE nat_jur != 1031 AND cbo_02 IS NOT NULL
-    GROUP BY rais.id_employee
-    HAVING rais.year = MAX(rais.year)
-    "
-)
-
-private_last_job <- private_last_job %>%
-    group_by(id_employee) %>%
-    ungroup()
-
-# join at the individual level last private sector job and bureaucracy job
-transition_private_bureaucracy <- private_last_job %>%
-    left_join(
[…]

### 2021-01-05 descriptive/modules/globals.R

@@ -1,22 +1,0 @@
-library(tidyverse)
-library(dbplyr)
-library(data.table)
-library(here)
-library(DBI)
-
-debug <- ifelse(
-    Sys.getenv("USER") == "gali" & dirname(here()) != "/hdd",
-    TRUE,
-    FALSE
-)
-
-data_path <- "data/clean/"
-source_path <- "source/modules/database/"
-rais_sql <- here(
-  sprintf(
-    "data/clean/database/rais%s.sqlite3",
-    ifelse(isTRUE(debug), "_sample", "")
-  )
-)
-
-rais_con <- DBI::dbConnect(RSQLite::SQLite(), rais_sql)

### 2021-01-05 descriptive/modules/partisan_vs_non_partisan.R

@@ -1,77 +1,0 @@
-# ==============================================================================
-# input: rais municipal sql tables
-# output: descriptive statistics on partisan vs. non-partisan
-# ==============================================================================
-source(
-    here::here("source/descriptive/modules/globals.R")
-)
-
-source(
-    here("source/descriptive/modules/requirements.R")
-)
-
-# ==============================================================================
-# load-in data
-# ==============================================================================
-rais_mun_tables <- c(
-    partisan = "rais_mun_partisan",
-    non_partisan = "rais_mun_non_partisan"
-)
-
-rais_filiado <- rais_mun_tables %>%
-map_dfr(
-        ~ tbl(rais_con, .) %>%
-            mutate(year = as.integer(year)) %>%
-            collect(),
-        .id = "is_partisan"
-    )
-
-outcome_vars <- c(
-    Age = "age", 
-    Education = "edu",
-    Dismissals = "fired",
-    Hires = "hired",
-    Experience = "work_experience",
-    Wage = "wage"
-)
-
-# # compute statistics per year
-# rais_filiado_summary <- rais_filiado %>%
-#     group_by(is_partisan, year) %>%
-#     summarise(
-#         across(
-#             c("age", "edu", "fired", "hired", "work_experience"),
-#             mean, na.rm = TRUE
-#             ),
-#             wage = median(wage, na.rm = TRUE)
-#     )
-
-plot_descriptive <- map2(
-        outcome_vars,
-        names(outcome_vars),
-        ~ ggplot(
-            aes_string("year", .x, color = "is_partisan"),
-            data = rais_filiado
-        ) +
-        geom_smooth() +
-        guides(color = guide_legend("Partisanship")) +
-        labs(x = "Year", y = .y)
-    )
-
-message("exporting plots")
-plot_filenames <- generate_plot_filenames(
-    "paper/figures/active_employee/",
-    outcome_vars,
-    debug
-)
-
-pmap(
-    list(
-        plot = plot_descriptive,
-        filename = plot_filenames
-    ),
-    ggsave
-)
-
-message("module completed: generate descriptive statistics for active employees.")
-

### 2021-01-05 descriptive/modules/party_membership.R

@@ -1,80 +1,0 @@
-# ==============================================================================
-# input: sql database of party affiliation
-# output: compare membership cycles for workers and non-workers
-# ==============================================================================
-source(
-    here::here("source/descriptive/modules/globals.R")
-)
-
-source(
-    here("source/descriptive/modules/requirements.R")
-)
-
-# ==============================================================================
-# generate descriptive statistics of party members
-# ==============================================================================
-rais_sql <- tbl(rais_con, "rais") %>%
-    filter_municipal()
-    
-filiado_sql <- tbl(rais_con, "filiado_mun")
-
-rais_tables <- c(
-    partisan = "rais_mun_partisan",
-    non_partisan = "rais_mun_non_partisan"
-)
-
-rais_filiado <- map_dfr(
-    rais_tables,
-    ~ dbGetQuery(sprintf("SELECT * FROM %s", .)),
-    .id = "is_filiado"
-)
-
-# create entries and exits per year
-# segmented by employment
-filiado_unemployed <- filiado_sql %>%
-    anti_join(
-        rais_sql %>%
-            select(cod_ibge_6, id_employee),
-        by = c("cod_ibge_6", "id_employee")
-    ) %>%
-    collect()
-
-filiado_employed <- filiado_sql %>%
-    inner_join(
-        rais_sql %>%
-            select(cod_ibge_6, id_employee),
-        by = c("cod_ibge_6", "id_employee")
-    ) %>%
-    collect() %>%
-    distinct(filiado_id, .keep_all = TRUE)
-
-# compare duration of party membership between employed vs. unemployed
-filiado_turnover <- list(
-    filiado_employed,
-    filiado_unemployed
-) %>%
-    set_names(c("employed", "unemployed")) %>%
-    map_dfr(
-        . %>%
-            convert_to_date() %>%
-            pivot_longer(
-                c(date_start, date_end),
-                names_prefix = "date_",
-                names_to = "date"
-            ) %>%
-            count(date, value),
-        .id = "is_employed"
-    )
-
-# verify when data for membership termination begins
-filiado_turnover %>%
-    gg_point(
-        aes(x = value, y = n),
-        size = 2
-    ) +
-    plot_mandate_year() +
-    facet_wrap(. ~ date + is_employed) +
-    coord_cartesian(ylim = c(0, 50)) +
-    scale_x_date(limit = as.Date(c("2003-01-01", "2019-01-01")))
-
-# compare membership spells

### 2021-01-05 descriptive/modules/prior_to_bureaucracy.R

@@ -1,104 +1,0 @@
-# ==============================================================================
-# input: sql database of party affiliation
-# output: compare differential observables for partisans and non-partisans
-# before they enter the bureaucracy
-# ==============================================================================
-source(
-    here::here("source/descriptive/modules/globals.R")
-)
-
-source(
-    here("source/descriptive/modules/requirements.R")
-)
-
-filiado <- tbl(rais_con, "filiado_mun")
-career_filiado <- tbl(rais_con, "career_pre_bureaucracy_mun") %>%
-    collect()
-
-# # ---------------------------------------------------------------------------- #
-# # compute summary statistics for partisans and non-partisans
-# message("computing summary statistics...")
-
-# # note that the majority of partisans become party members after
-# # joining the bureau
-# plot_breakdown <- career_filiado %>%
-#     group_by(year, is_partisan) %>%
-#     summarise(n = sum(n)) %>%
-#     gg_point_line(
-#         aes(year, n/1e6, color = is_partisan)
-#     ) +
-#     scale_color_discrete(
-#         name = "Partisanship",
-#         labels = c(
-#             "pre_partisan" = "pre-partisan",
-#             "post_partisan" = "post-partisan",
-#             "non_partisan" = "non-partisan"
-#         )
-#     ) +
-#     labs(x = "Year", y = "Total (millions)")
-
-# group_vars <- c("year", "is_partisan")
-
-# career_filiado_mean <- career_filiado %>%
-#     compute_mean(
-#         all_of(group_vars),
-#         c(age, work_experience, edu)
-#     )
-
-# career_filiado_median <- career_filiado %>%
-#     group_by(
-#         across(all_of(group_vars))
-#     ) %>%
-#     compute_median(wage)
-
-# career_filiado_summary <- career_filiado_mean %>%
-#     left_join(
-#         career_filiado_median,
-#         by = group_vars
-#     ) %>%
-#     filter(!is.na(is_partisan))
-
-# message("complete!")
-
-# ---------------------------------------------------------------------------- #
-message("generating descriptive plots")
-
-vars_to_plot <- c("age", "work_experience", "edu") %>%
-    sprintf(fmt = "%s_mean", .) %>%
-    c("median_wage", "n")
-
-label_vars <- c("Age", "Work Experience", "Education", "Wage", "Total (million)")
-
-plot_summary <- map2(
-    vars_to_plot,
-    label_vars,
[…]

### 2021-01-05 descriptive/modules/requirements.R

@@ -1,703 +1,0 @@
-here_data <- partial(
-  here,
-  data_path
-)
-
-here_source <- partial(
-  here,
-  source_path
-)
-
-fread <- purrr::partial(
-  data.table::fread,
-  nThread = parallel::detectCores(),
-  nrows = nrows
-)
-
-dbGetQuery <- purrr::partial(
-  DBI::dbGetQuery,
-  conn = rais_con
-)
-
-nrows <- if (isTRUE(debug)) 1e4 else Inf
-
-# ==============================================================================
-# data io
-# ==============================================================================
-list_files <- function(path, pattern) {
-  files <- map2(
-    path,
-    pattern,
-    list.files
-  ) %>%
-    flatten_chr()
-
-  return(files)
-}
-
-write_data <- function(dir, filename, data){
-  file_extension <- str_extract(filename, "(?<=\\.)[a-z]+")
-
-  if(!dir.exists(dir)){
-    dir.create(dir, recursive = TRUE)
-  }
-
-  switch(
-    file_extension,
-    csv = fwrite(
-      data,
-      here(dir, filename)
-    ),
-    rds = write_rds(
-      data,
-      here(dir, filename)
-    )
-  )
-}
-
-# ==============================================================================
-# rais wrangling
-# ==============================================================================
-is_municipal <- rlang::quo(nat_jur == 1031)
-is_bureau <- rlang::quo(nat_jur == 1031 & occupation == 1)
-is_partisan <- rlang::quo(
-  year >= year_date_start & (year <= year_date_end | is.na(year_date_end))
-)
-
-year_to_char <- function(data) {
-  data <- data %>%
-    mutate_at(
-      vars(year),
-      as.character
-    )
-
-  return(data)
-}
-
-convert_to_date <- function(data) {
-  data %>%
-    mutate(
-      across(starts_with("date"), lubridate::ymd)
-    )
-}
-
-fix_edu <- function(data){
-  data %>%
-  mutate(
-            edu_category = case_when(
-                edu <= 2 ~ "illiterate",
-                between(edu, 3, 4) ~ "lower school",
-                between(edu, 5, 6) ~ "middle school",
-                between(edu, 7, 8) ~ "high school",
[…]

### 2021-01-05 descriptive/modules/turnover_cycles.R

@@ -1,66 +1,0 @@
-# ==============================================================================
-# input: sql database
-# output: descriptive statistics of turnover cycles in bureaucracy
-# ==============================================================================
-source(
-    here::here("source/descriptive/modules/globals.R")
-)
-
-source(
-    here::here("source/descriptive/modules/requirements.R")
-)
-
-rais <- tbl(rais_con, "rais")
-filiado <- tbl(rais_con, "filiado_mun")
-
-rais_tables <- c(
-    partisan = "rais_mun_partisan",
-    non_partisan = "rais_mun_non_partisan"
-)
-
-rais_filiado <- map_dfr(
-    rais_tables,
-    ~ dbGetQuery(sprintf("SELECT * FROM %s", .)),
-    .id = "is_filiado"
-)
-
-message("generate plots")
-vars_to_plot <- c("hired", "fired", "work_experience")
-plot_turnover <- map(
-    vars_to_plot,
-    ~ gg_summary(
-        x = "year",
-        y = .,
-        data = rais_filiado,
-        smooth = FALSE,
-        color = !!sym("is_filiado")
-    ) +
-        ggtitle(.)
-)
-
-rais_filiado %>%
-    ggplot(
-        aes(year, hired, color = is_filiado)
-    ) +
-    stat_summary_bin(
-        fun = "mean",
-        geom = "point"
-    )
-
-path_to_figs <- ifelse(
-    isTRUE(debug), "paper/figures/turnover/sample/", "paper/figures/turnover/"
-)
-
-file_names <- sprintf(
-    "plot_%s.pdf", vars_to_plot
-)
-
-message("save plots")
-path_out <- here(path_to_figs, file_names)
-pwalk(
-    list(
-        filename = path_out,
-        plot = plot_turnover
-    ),
-    save_plot
-)

### 2021-01-05 source/descriptive/modules/bureau_partisanship.R

@@ -10,74 +10,9 @@ source(
     here("source/descriptive/modules/requirements.R")
 )
 
-message("extract records and join with filiado data")
-
-rais_filiado <- dbGetQuery(
-    "
-    SELECT 
-        rais.cod_ibge_6,
-        rais.year,
-        rais.id_employee,
-        rais.cbo_02,
-        rais.cnae_95,
-        rais.edu,
-        rais.hours,
-        rais.wage,
-        rais.work_experience,
-        rais.contract_type,
-        filiado_mun.filiado_id,
-        filiado_mun.date_start,
-        filiado_mun.date_end,
-        filiado_mun.is_filiado
-    FROM rais
-    LEFT JOIN (SELECT *, 1.0 AS is_filiado FROM filiado_mun) AS filiado_mun
-    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
-    rais.id_employee = filiado_mun.id_employee
-    "
+partisan_summary <- read_rds(
+    here("data/clean/summary/partisan_dominance.rds")
 )
-
-rais_filiado <- tbl(rais_con, "rais") %>%
-    mutate(
-        occupation = str_sub(cbo_02, 1, 1)
-    ) %>%
-    mutate(
-        cbo_group_detail = case_when(
-        occupation == 0 ~ "army",
-        `1` = "executive",
-        `2` = "liberal arts",
-        `3` = "engineer",
-        `4` = "administration",
-        `5` = "services",
-        `6` = "agriculture",
-        `7` = "industry",
-        `8` = "industry",
-        `9` = "maintenance"
-      )
-    )
-
-dbGetQuery(
-    "
-    SELECT
-    edu,
-    AVG(COALESCE(is_filiado, 0)) AS party_dominance
-    FROM rais
-    LEFT JOIN (SELECT *, 1.0 AS is_filiado FROM filiado_mun) AS filiado_mun
-    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
-    rais.id_employee = filiado_mun.id_employee
-    GROUP BY edu
-    "   
-)
-
-message("generate party dummy and compute partisanship by group")
-rais_filiado_occupation <- rais_filiado %>%
-    mutate(
-        occupation = str_sub(cbo_02, 1, 1)
-    ) %>% 
-    fix_occupation() %>%
-    fix_edu() %>%
-    filter(
-        !is.na(cbo_group) & !is.na(cbo_group_detail)
-    )
 
[…]

### 2021-01-05 source/descriptive/modules/prior_to_bureaucracy.R

@@ -11,54 +11,9 @@ source(
     here("source/descriptive/modules/requirements.R")
 )
 
-filiado <- tbl(rais_con, "filiado_mun")
 career_filiado <- tbl(rais_con, "career_pre_bureaucracy_mun") %>%
-    collect()
-
-# # ---------------------------------------------------------------------------- #
-# # compute summary statistics for partisans and non-partisans
-# message("computing summary statistics...")
-
-# # note that the majority of partisans become party members after
-# # joining the bureau
-# plot_breakdown <- career_filiado %>%
-#     group_by(year, is_partisan) %>%
-#     summarise(n = sum(n)) %>%
-#     gg_point_line(
-#         aes(year, n/1e6, color = is_partisan)
-#     ) +
-#     scale_color_discrete(
-#         name = "Partisanship",
-#         labels = c(
-#             "pre_partisan" = "pre-partisan",
-#             "post_partisan" = "post-partisan",
-#             "non_partisan" = "non-partisan"
-#         )
-#     ) +
-#     labs(x = "Year", y = "Total (millions)")
-
-# group_vars <- c("year", "is_partisan")
-
-# career_filiado_mean <- career_filiado %>%
-#     compute_mean(
-#         all_of(group_vars),
-#         c(age, work_experience, edu)
-#     )
-
-# career_filiado_median <- career_filiado %>%
-#     group_by(
-#         across(all_of(group_vars))
-#     ) %>%
-#     compute_median(wage)
-
-# career_filiado_summary <- career_filiado_mean %>%
-#     left_join(
-#         career_filiado_median,
-#         by = group_vars
-#     ) %>%
-#     filter(!is.na(is_partisan))
-
-# message("complete!")
+    collect() %>%
+    mutate(year = as.numeric(year))
 
 # ---------------------------------------------------------------------------- #
 message("generating descriptive plots")
@@ -67,15 +22,18 @@ vars_to_plot <- c("age", "work_experience", "edu") %>%
     sprintf(fmt = "%s_mean", .) %>%
     c("median_wage", "n")
 
-label_vars <- c("Age", "Work Experience", "Education", "Wage", "Total (million)")
+label_vars <- c(
+    "Age", "Work Experience", "Education", "Wage", "Total (million)"
+)
 
 plot_summary <- map2(
     vars_to_plot,
     label_vars,
-    ~ gg_point_line(
+    ~ ggplot(
         data = career_filiado,
[…]

### 2021-01-05 source/descriptive/modules/requirements.R

@@ -38,6 +38,10 @@ list_files <- function(path, pattern) {
 write_data <- function(dir, filename, data){
   file_extension <- str_extract(filename, "(?<=\\.)[a-z]+")
 
+  if(!dir.exists(dir)){
+    dir.create(dir, recursive = TRUE)
+  }
+
   switch(
     file_extension,
     csv = fwrite(
@@ -581,7 +585,20 @@ gg_bar <- function(data, x, y){
       aes(reorder(!!sym(x), {{y}}), {{y}})
     ) +
     geom_col() +
-    labs(x = x)
+    labs(x = x) +
+    coord_flip()
+
+  return(plot)
+}
+
+gg_boxplot <- function(data, x, y, xlab){
+  plot <- data %>%
+    ggplot(
+      aes(reorder(!!sym(x), {{y}}), {{y}})
+    ) +
+    geom_boxplot() +
+    labs(x = xlab) +
+    coord_flip()
 
   return(plot)
 }
@@ -675,12 +692,9 @@ plot_mandate_year <- function(years = seq(2005, 2013, 4)) {
 save_plot <- function(plot, filename){
   dir <- dirname(filename)
 
-    if(!dir.exists(dir)){
+  if(!dir.exists(dir)){
     dir.create(dir, recursive = TRUE)
-  }else{
-    unlink(dir, recursive = T)
-    dir.create(dir, recursive = TRUE)
-    }
+  }
 
     ggsave(
       filename,

### 2021-01-05 source/descriptive/build.R
create transition amtrix

@@ -7,6 +7,6 @@ module_selection <- c(
     # "party_membership",
     "bureau_partisanship",
     "prior_to_bureaucy",
-    # "turnover_cycles",
+    "entry_into_bureaucracy",
     "partisan_vs_non_partisan"
-)   
+)

### 2021-01-05 paper/draft.tex
fix bibliography
edit intro
analyze further the differential compensation

@@
 
+\section{Introduction}
+\label{sec:intro}
+
+How does
+
+Literature on patronage and selection: \citet{robinson2013political}.
+Extant literature on selection of public officials have primarily focused on who becomes a politician \citet{dal2017becomes}.
+There remains unanswered questions regarding who enters the bureaucracy, what types of employment they held, and what are the differential compensations among public officials.
+
+In this paper, I leverage micro-level partisan information with employment data to assess the degree to which there are differential compensation structures and differences in observables among partisan and non-partisan.
+What I find is that the benefits of public secotr employment do not accrue to poorer voters.
+Rather
+
+\section{Context and Data}
+\label{sec:context}
+
+\section{Empirical strategy}
+\label{sec:empirical}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+\newpage
+
+\bibliography{patronage}
+
 \end{document}

### 2021-01-07 source/data/modules/database/generate_municipal_tables.R
fix filiado data

@@ -13,14 +13,13 @@ source(
 
 RSQLite::initExtension(rais_con)
 
-sprintf(
-    "DROP TABLE IF EXISTS %s", c("rais_mun_partisan", "rais_mun_non_partisan")
-) %>%
-    walk(dbExecute)
+# sprintf(
+#     "DROP TABLE IF EXISTS %s", c("rais_mun_partisan", "rais_mun_non_partisan")
+# ) %>%
+#     walk(dbExecute)
 
-dbExecute(
+rais_mun_non_partisan <- dbGetQuery(
     "
-    CREATE TABLE IF NOT EXISTS rais_mun_non_partisan AS
     SELECT
         rais.cod_ibge_6,
         rais.year,
@@ -33,15 +32,15 @@ dbExecute(
         COUNT(*) AS n
     FROM rais
     LEFT JOIN filiado_mun
-    ON rais.id_employee = filiado_mun.id_employee
+    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6, 
+    rais.id_employee = filiado_mun.id_employee
     WHERE rais.nat_jur = 1031 AND filiado_mun.id_employee IS NULL
     GROUP BY rais.cod_ibge_6, rais.year
     "
 )
 
-dbExecute(
+rais_mun_partisan <- dbGetQuery(
     "
-    CREATE TABLE IF NOT EXISTS rais_mun_partisan AS
     SELECT
         rais.cod_ibge_6,
         rais.year,
@@ -54,9 +53,39 @@ dbExecute(
         COUNT(*) AS n
     FROM rais
     INNER JOIN filiado_mun
-    ON rais.id_employee = filiado_mun.id_employee
+    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
+    rais.id_employee = filiado_mun.id_employee
     WHERE rais.nat_jur = 1031
     GROUP BY rais.cod_ibge_6, rais.year
+    "
+)
+
+message("write out tables")
+rais_mun_non_partisan %>%
+    write_data(
+        dir = "summary",
+        "rais_mun_non_partisan.csv"
+    )
+
+rais_mun_partisan <- dbGetQuery(
+    "
+    SELECT
+        rais.cod_ibge_6,
+        rais.year,
+        rais.hired,
+        rais.fired,
+        rais.age,
+        rais.edu,
+        rais.work_experience,
+        rais.wage,
+        filiado_mun.party,
[…]

### 2021-01-12 source/data/modules/database/generate_municipal_tables.R
export maps

@@ -18,6 +18,7 @@ RSQLite::initExtension(rais_con)
 # ) %>%
 #     walk(dbExecute)
 
+message("non partisan")
 rais_mun_non_partisan <- dbGetQuery(
     "
     SELECT
@@ -32,13 +33,15 @@ rais_mun_non_partisan <- dbGetQuery(
         COUNT(*) AS n
     FROM rais
     LEFT JOIN filiado_mun
-    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6, 
+    ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
     rais.id_employee = filiado_mun.id_employee
-    WHERE rais.nat_jur = 1031 AND filiado_mun.id_employee IS NULL
+    WHERE rais.nat_jur = 1031 AND
+    filiado_mun.id_employee IS NULL
     GROUP BY rais.cod_ibge_6, rais.year
     "
 )
 
+message("partisan")
 rais_mun_partisan <- dbGetQuery(
     "
     SELECT
@@ -55,38 +58,20 @@ rais_mun_partisan <- dbGetQuery(
     INNER JOIN filiado_mun
     ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND
     rais.id_employee = filiado_mun.id_employee
-    WHERE rais.nat_jur = 1031
+    WHERE rais.nat_jur = 1031 AND
+    CAST(rais.year AS INT) >= filiado_mun.year_start AND
+    CAST(rais.year AS INT) <= filiado_mun.year_termination
     GROUP BY rais.cod_ibge_6, rais.year
     "
 )
 
 message("write out tables")
-rais_mun_non_partisan %>%
-    write_data(
-        dir = "summary",
-        "rais_mun_non_partisan.csv"
-    )
-
-rais_mun_partisan <- dbGetQuery(
-    "
-    SELECT
-        rais.cod_ibge_6,
-        rais.year,
-        rais.hired,
-        rais.fired,
-        rais.age,
-        rais.edu,
-        rais.work_experience,
-        rais.wage,
-        filiado_mun.party,
-        filiado_mun.date_start,
-        filiado_mun.date_end,
-        filiado_mun.date_cancel
-    FROM rais
-    INNER JOIN filiado_mun
-    ON rais.id_employee = filiado_mun.id_employee
-    WHERE rais.nat_jur = 1031
-    "
+export <- list(
[…]

### 2021-01-12 source/data/modules/database/generate_partisanship.R

@@ -41,6 +41,7 @@ rais_filiado <- map(
         c("occupation", "contract", "education")
     )
 
+gg_ma
 # fix grouping columns
 rais_filiado <- rais_filiado %>%
     modify_at(

### 2021-01-12 source/data/modules/database/upload_filiado_sql.R

@@ -38,9 +38,9 @@ create_table <- sprintf(
         id_employee INTEGER,
         cod_ibge_6 TEXT,
         party TEXT,
-        date_start DATE,
-        date_end DATE,
-        date_cancel DATE,
+        date_start INTEGER,
+        date_end INTEGER,
+        date_cancel INTEGER,
         year_start INTEGER,
         year_end INTEGER,
         year_cancel INTEGER,
@@ -95,7 +95,8 @@ filiado <- filiado %>%
         ~ mutate(
             .,
             across(
-                starts_with("date"), ~ str_remove_all(., "-")
+                starts_with("date"), ~ str_remove_all(., "-") %>% 
+                as.integer
             ),
             across(
                 starts_with("year"), as.integer

### 2021-01-12 source/descriptive/modules/globals.R

@@ -3,6 +3,7 @@ library(dbplyr)
 library(data.table)
 library(here)
 library(DBI)
+library(sf)
 
 debug <- ifelse(
     Sys.getenv("USER") == "gali" & dirname(Sys.getenv("PWD")) != "/hdd",

### 2021-01-12 source/descriptive/modules/mapping_partisanship.R

@@ -1,0 +1,120 @@
+# ==============================================================================
+# input: rais municipal sql tables
+# output: descriptive statistics on partisan vs. non-partisan
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here("source/descriptive/modules/requirements.R")
+)
+
+plot_map_partisanship <- 
+# ---------------------------------------------------------------------------- #
+message("import map")
+
+map_br <- st_read(
+    here("data/clean/maps"),
+    layer = "municipio"
+) %>%
+    mutate(
+        cod_ibge_6 = MUNICÍPI0 %>%
+            as.character() %>%
+            as.integer()
+    )
+
+# ---------------------------------------------------------------------------- #
+message("importing data")
+party_dominance <- read_data(
+    "clean",
+    "summary",
+    "partisan_dominance.rds"
+)
+
+party_dominance_mun <- party_dominance %>%
+    pluck("occupation") %>%
+    filter(year == max(year)) %>%
+    group_by(cod_ibge_6) %>%
+    summarise(
+        partisanship = weighted.mean(partisanship, n),
+        .groups = "drop"
+    )
+
+party_dominance_disaggregated <- party_dominance %>%
+    map(
+        ~ filter(., year == max(year))
+    )
+
+# fix education categories
+party_dominance_disaggregated <- party_dominance_disaggregated %>%
+    modify_at(
+        "education",
+    ~mutate(
+        ., 
+        edu_category = if_else(
+            str_detect(edu_category, "lower|middle"),
+            "lower-middle school",
+            edu_category
+        )
+    ) %>%
+    group_by(cod_ibge_6, edu_category) %>%
+    summarise(
+        partisanship = weighted.mean(partisanship, n),
+        .groups = "drop"
+    )
+    )
+
+# ---------------------------------------------------------------------------- #
+message("join data")
+map_br_partisan <- map_br %>%
+    left_join(
+        party_dominance_mun,
+        by = "cod_ibge_6"
+    )
+
+map_br_disaggregated <- party_dominance_disaggregated %>%
+    map(
+        ~ left_join(
+            map_br,
+            .,
+            by = "cod_ibge_6"
+        )
+    )
+
+
+map_partisanship <- map_br_partisan %>%
+    gg_map(partisanship)
+
[…]

### 2021-01-12 source/descriptive/modules/partisan_vs_non_partisan.R

@@ -14,15 +14,13 @@ source(
 # load-in data
 # ==============================================================================
 rais_mun_tables <- c(
-    partisan = "rais_mun_partisan",
-    non_partisan = "rais_mun_non_partisan"
+    partisan = "rais_mun_partisan.csv.gz",
+    non_partisan = "rais_mun_non_partisan.csv.gz"
 )
 
 rais_filiado <- rais_mun_tables %>%
-map_dfr(
-        ~ tbl(rais_con, .) %>%
-            mutate(year = as.integer(year)) %>%
-            collect(),
+    map_dfr(
+        ~read_data(., type = "clean", dir = "summary"),
         .id = "is_partisan"
     )
 

### 2021-01-12 source/descriptive/modules/requirements.R

@@ -35,6 +35,24 @@ list_files <- function(path, pattern) {
   return(files)
 }
 
+here_data <- function(type, dir, file) {
+  path <- here("data", type, dir, file)
+  
+  return(path)
+}
+
+read_data <- function(type, dir, file) {
+  file_path <- here_data(type, dir, file)
+
+  if (str_detect(file, ".rds$")) {
+    data <- read_rds(file_path)
+  } else {
+    data <- fread(file_path)
+  }
+  
+  return(data)
+}
+
 write_data <- function(dir, filename, data){
   file_extension <- str_extract(filename, "(?<=\\.)[a-z]+")
 
@@ -496,6 +514,22 @@ gg_point_line <- function(data, mapping = aes(), ...) {
     geom_line(
       ...
     )
+}
+
+gg_map <- function(data, var){
+  ggplot(
+    data = data,
+        aes(
+            fill = {{var}},
+            color = NA
+        )
+    ) +
+    geom_sf() +
+    scale_fill_distiller(
+        palette = "YlGnBu",
+        direction = 1
+    ) +
+    theme_void()
 }
 
 geom_errorbar_tidy <- geom_errorbarh(

### 2021-01-12 paper/draft.tex
expand intro

@@
 In this paper, I leverage micro-level partisan information with employment data to assess the degree to which there are differential compensation structures and differences in observables among partisan and non-partisan.
-What I find is that the benefits of public secotr employment do not accrue to poorer voters.
-Rather
+What I find is that the benefits of public sector employment do not accrue to poorer voters.
+Rather, what we see in Brazil is the accumulation of the benefits of public employment to a wealthier class of party members.
+Positions at higher levels of government such as executive leadership and administration are mostly allocated to party members.
+In contrast, less profitable offices in frontline services are less partisan.
+
+Additionally, party members are concentrated in tenured positions.
+Over X\% of career bureaucrats are affiliated to a party.
+Party members tend to stay in the bureaucracy for longer, and receive a higher compensation than their non-partisan counterparts.
 

### 2021-01-12 source/descriptive/modules/mapping_partisanship.R

@@ -110,7 +110,7 @@ map_education <- map_br_disaggregated %>%
 export_maps <- list(
     plot = list(map_occupation, map_contract, map_education),
     filename = sprintf(
-        here("paper/figures/maps/%s.pdf"), names(party_dominance)
+        here("paper/figures/maps/%s.png"), names(party_dominance)
         )
 )
 

### 2021-01-13 source/descriptive/modules/prior_to_bureaucracy.R
alter prior to bureaucracy plots

@@ -15,6 +15,18 @@ career_filiado <- tbl(rais_con, "career_pre_bureaucracy_mun") %>%
     collect() %>%
     mutate(year = as.numeric(year))
 
+career_filiado <- career_filiado %>%
+    filter(
+        is_partisan != "post_partisan"
+    ) %>%
+    mutate(
+        is_partisan = recode(
+            is_partisan,
+            "non_partisan" = "non_partisan",
+            "pre_partisan" = "partisan"
+        )
+    )
+
 # ---------------------------------------------------------------------------- #
 message("generating descriptive plots")
 

### 2021-01-13 source/descriptive/modules/prior_to_bureaucracy.R
finalize new outline

@@ -45,7 +45,9 @@ plot_summary <- map2(
         data = career_filiado,
         aes_string(x = "year", y = .x, color = "is_partisan")
     ) +
-    geom_smooth() +
+    geom_smooth(
+        span = 3
+    ) +
     scale_colour_discrete(
         name = "Partisanship",
         labels = c(

### 2021-01-13 paper/draft.tex
finalize new outline
finalize intro

--- 2021-01-05 paper/draft.tex
@@
 Leveraging a novel dataset of partisan affiliation and employment data on every municipal bureaucrat in Brazil, I find that party members are more likely to be overcompensated than their peers, concentrating in areas of executive leadership, while being less educated than their peers.
-In addition, their employment spells tend to be more durable over time, leading to negative consequences.
-These findings provide raise important questions regarding the duration of partisan ties and its consequences for local bureaucracies.}
+In addition, their employment spells tend to be more durable over time, leading to the build-up of party members in the bureaucracy over time.
+These findings provide raise important questions regarding the nature of patronage, party building and its consequences for local bureaucracies.}
 
@@
 
-How does
+Who benefits from patronage? And what are its consequences for local administration?
 
@@
 
-In this paper, I leverage micro-level partisan information with employment data to assess the degree to which there are differential compensation structures and differences in observables among partisan and non-partisan.
-What I find is that the benefits of public secotr employment do not accrue to poorer voters.
-Rather
+Empirical research on allocation of public sector jobs has consistently found that politically motivated allocation of public sector jobs is an important determinant of public employment in the developed and developing world \citep{finan2017personnel}.
+In the United States, appointments to the federal bureaucracy involve considerations of party loyalty and ideological alignment with the president \citep{lewis2010politics, hollibaugh2014presidents}.
+In Brazil, whether it be at the presidential \citep{pracca2011political} or at the municipal level \citep{colonnelli2018patronage,brollo2017victor}, politicians have discretion over bureaucratic appointments and frequently use these to further their political goals.
+
+A set of explanations have been proposed to explain the determinants of this allocation, whether it be to reduce frictions in policy implementation due to ideological divergence \citep{krause2016experiential}, exploit the benefits of strong ties \citep{toral2019benefits} or to reward followers for campaign contributions \citep{colonnelli2018patronage}.
+These findings have provided important insights into the drivers of patronage appointments into employment in the public sector.
+Yet as important as the drivers of public sector appointment are the characteristics of those entering the bureaucracy \citep{finan2015personnel}.
+Who becomes a bureaucrat?
+Are political appointees systematically different from their non-partisan counterparts?
+And where, within the bureaucracy, do these patronage appointees go?
+
+In this paper, I focus on party membership as a determinant of not only whether or not individuals enter into public sector employment, but the kind of employment you receive.
+In particular, using a rich panel dataset of all public sector workers, I identify the party membership of all municipal bureaucrats in Brazil to provide a novel set of empirical findings: 1) the pre-bureaucracy characteristics of public sector workers, 2)
+
+In this paper, I leverage micro-level partisan information with employment data to assess the degree to which there are differential compensation structures and differences in observables among partisan and non-partisan members of the bureaucracy.
+The main finding is that patronage is an elite game, benefitting the wealthiest formal sector workers and allocating them into the best-paying jobs in the public sector.
+The benefits of public sector employment accrue to a wealthier class of workers, in particular those already in profitable positions in the private sector.
+Additionally, wealthy party members accumulate jobs at higher levels of government such as executive leadership and administration, which are better compensated.
+In contrast, less profitable offices in frontline services experience less exposure to partisan members.
+
+Additionally, party members are concentrated in tenured positions.
+Over X\% of career bureaucrats are affiliated to a party.
+Party members tend to stay in the bureaucracy for longer, and receive a higher compensation than their non-partisan counterparts.
+
+Patronage in the allocation of public sector jobs
 

### 2021-01-14 paper/__latexindent_temp.tex
generate intro

--- 2021-01-13 paper/draft.tex
@@
 
-In this paper, I focus on party membership as a determinant of not only whether or not individuals enter into public sector employment, but the kind of employment you receive.
-In particular, using a rich panel dataset of all public sector workers, I identify the party membership of all municipal bureaucrats in Brazil to provide a novel set of empirical findings: 1) the pre-bureaucracy characteristics of public sector workers, 2)
+In this paper, I focus on party membership as a determinant of not only whether or not individuals enter into public sector employment, but the type of employment they receive.
+In particular, using a rich panel data set of all public sector workers, I identify the party membership of all municipal bureaucrats in Brazil to provide a novel set of empirical findings: 1) the pre-bureaucracy characteristics of public sector workers, 2) employment trajectories within the bureaucracy and 3) post-public sector employment of party members.
+This complete revolving-door of party and non-party members provide a unique frame-by-frame evolution of the employment trajectories of pre- and post-bureaucrats in a developing world context.
 
-In this paper, I leverage micro-level partisan information with employment data to assess the degree to which there are differential compensation structures and differences in observables among partisan and non-partisan members of the bureaucracy.
-The main finding is that patronage is an elite game, benefitting the wealthiest formal sector workers and allocating them into the best-paying jobs in the public sector.
-The benefits of public sector employment accrue to a wealthier class of workers, in particular those already in profitable positions in the private sector.
-Additionally, wealthy party members accumulate jobs at higher levels of government such as executive leadership and administration, which are better compensated.
-In contrast, less profitable offices in frontline services experience less exposure to partisan members.
+The main finding is that patronage primarily benefits a local economic elite, accruing to the richest formal sector workers and allocating them into the best-paying jobs in the municipal bureaucracy.
+These higher compensation structures are not commensurate to skills on a set of observable qualifications, such as education level and work experience, suggesting that these benefits accrue from channels other than individual skills.
+Additionally, party members accumulate jobs at higher levels of government such as executive leadership and administration, which are better compensated, as well as benefiting from longer income streams due to privileged access to tenured contracts.
 
-Additionally, party members are concentrated in tenured positions.
-Over X\% of career bureaucrats are affiliated to a party.
-Party members tend to stay in the bureaucracy for longer, and receive a higher compensation than their non-partisan counterparts.
-
-Patronage in the allocation of public sector jobs
+These findings suggest that, overall, patronage does not necessarily accrue to poor voters, contrary to theoretical and empirical findings in studies of clientelism in the developing world \citep{stokes2013brokers}.
+A different logic seems to be at play.
+Patronage to party loyalists can be used strategically as a mean of securing control over the bureaucracy -- through tenured contracts -- as well as securing buy-in from wealthy patrons or notables in the local economy, consistent with empirical findings by \citet{colonnelli2018patronage}.
+Patronage therefore can serve a crucial role in securing access to economic resources through wealthy patrons, which in turn can be allocated to finance campaigns and other exercises in party building.
 

### 2021-01-14 paper/draft.tex
finish contribution

@@
 
-In this paper, I focus on party membership as a determinant of not only whether or not individuals enter into public sector employment, but the kind of employment you receive.
-In particular, using a rich panel dataset of all public sector workers, I identify the party membership of all municipal bureaucrats in Brazil to provide a novel set of empirical findings: 1) the pre-bureaucracy characteristics of public sector workers, 2)
+In this paper, I focus on party membership as a determinant of not only whether or not individuals enter into public sector employment, but the type of employment they receive.
+In particular, using a rich panel data set of all public sector workers, I identify the party membership of all municipal bureaucrats in Brazil to provide a novel set of empirical findings: 1) the pre-bureaucracy characteristics of public sector workers, 2) employment trajectories within the bureaucracy and 3) post-public sector employment of party members.
+This complete revolving-door of party and non-party members provide a unique frame-by-frame evolution of the employment trajectories of pre- and post-bureaucrats in a developing world context.
 
-In this paper, I leverage micro-level partisan information with employment data to assess the degree to which there are differential compensation structures and differences in observables among partisan and non-partisan members of the bureaucracy.
-The main finding is that patronage is an elite game, benefitting the wealthiest formal sector workers and allocating them into the best-paying jobs in the public sector.
-The benefits of public sector employment accrue to a wealthier class of workers, in particular those already in profitable positions in the private sector.
-Additionally, wealthy party members accumulate jobs at higher levels of government such as executive leadership and administration, which are better compensated.
-In contrast, less profitable offices in frontline services experience less exposure to partisan members.
+The main finding is that patronage primarily benefits a local economic elite, accruing to the richest formal sector workers and allocating them into the best-paying jobs in the municipal bureaucracy.
+These higher compensation structures are not commensurate to skills on a set of observable qualifications, such as education level and work experience, suggesting that these benefits accrue from channels other than individual skills.
+Additionally, party members accumulate jobs at higher levels of government such as executive leadership and administration, which are better compensated, as well as benefiting from longer income streams due to privileged access to tenured contracts.
 
-Additionally, party members are concentrated in tenured positions.
-Over X\% of career bureaucrats are affiliated to a party.
-Party members tend to stay in the bureaucracy for longer, and receive a higher compensation than their non-partisan counterparts.
+These findings suggest that, overall, patronage does not necessarily accrue to poor voters, contrary to theoretical and empirical findings in studies of clientelism in the developing world \citep{stokes2013brokers}.
+A different logic seems to be at play.
+Patronage to party loyalists can be used strategically as a mean of securing control over the bureaucracy -- through tenured contracts -- as well as securing buy-in from wealthy patrons or notables in the local economy, consistent with empirical findings by \citet{colonnelli2018patronage}.
+Patronage therefore can serve a crucial role in securing access to economic resources through wealthy patrons, which in turn can be allocated to finance campaigns and other exercises in party building.
 
-Patronage in the allocation of public sector jobs
+This paper contributes to literature on clientelism that outlines the political logic of patronage allocations.
+The contribution is twofold: 1) first, I find that patronage is an elite-elite game, generating a form of patronage that is distinct from the politician-voter nexus that has been traditionally the focus of extant literature on clientelism \citet{diaz2016political} and 2) I find that the benefits of patronage go beyond simply an electoral pay-off.
+Instead, what patronage accomplishes is securing access to economic resources that can be tapped into once employment is offered to wealthy patrons, similar in vein to the theoretical findings by \citet{robinson2013political}.
+Patronage ties the hands of parties and patrons, in effect capturing the benefits of public sector employment to finance efforts towards party-building.
 

### 2021-01-15 paper/draft.tex
expand context and data

@@
 This paper contributes to literature on clientelism that outlines the political logic of patronage allocations.
-The contribution is twofold: 1) first, I find that patronage is an elite-elite game, generating a form of patronage that is distinct from the politician-voter nexus that has been traditionally the focus of extant literature on clientelism \citet{diaz2016political} and 2) I find that the benefits of patronage go beyond simply an electoral pay-off.
-Instead, what patronage accomplishes is securing access to economic resources that can be tapped into once employment is offered to wealthy patrons, similar in vein to the theoretical findings by \citet{robinson2013political}.
-Patronage ties the hands of parties and patrons, in effect capturing the benefits of public sector employment to finance efforts towards party-building.
+The contribution is twofold: 1) first, I find that patronage is a patron-elite game, generating a form of patronage that is distinct from the politician-voter nexus that has been traditionally the focus of extant literature on clientelism \citet{stokes2013brokers, diaz2016political} and 2) I find that the benefits of patronage go beyond simply an electoral pay-off.
+Instead, what patronage accomplishes is securing access to economic resources that can be tapped into once employment is offered to wealthy patrons, similar in spirit to the theoretical findings by \citet{robinson2013political}.
+Patronage binds parties and patrons, in effect capturing the benefits of public sector employment to finance efforts towards party-building.
+
+The paper is structured as follows.
+Section \ref{sec:context} provides context for the data and the hiring process for local bureaucracies, while section \ref{sec:descriptive} provides some descirptive statistics outlining the differences between partisan and non-partisan members.
+Section \ref{sec:empirical} outlines the empirical strategy.
+Section \ref{sec:conclusion} concludes.
 
@@
 \label{sec:context}
+
+\subsection{Partisan affiliation in Brazil}
+
+Partisanship in Brazil is voluntary, with over 11 percent of registered voters affiliated to a party \citep{}
+
+There is an ongoing debate on the strength of partisan ties in Brazil.
+On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citet{desposato2006parties,ames2002deadlock}.
+On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}.
+Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- which disproportionately
+
+\section{Descriptive Statistics}
+\label{sec:descriptive}
 

### 2021-01-15 paper/draft.tex
additional info regarding party registration

@@
 
-Partisanship in Brazil is voluntary, with over 11 percent of registered voters affiliated to a party \citep{}
+Partisanship in Brazil is voluntary and widespread, with over 11 percent of registered voters affiliated to a party \citep{speck2015estudo}\footnote{For context, in most OECD countries party registration does not exceed 5 percent of the electorate.
+See \citet{biezen2014decline}.}.
+Registration occurs in the following sequence: a voter reports to a local party office, and the party officials then register the voter officially through the \emph{Tribunal Regional Eleitoral} (Regional Electoral Office).
+This registration is then centralized by the \emph{Tribunal Superior Eleitoral} (Supreme Electoral Office) and updated accordingly.
+If there are overlapping registrations, former ones become annulled and are reported as irregular to local party officials.
+Each voter is therefore only allowed to register for a single party, without any ceilings or floors regarding the duration of this affiliation.
 

### 2021-01-15 paper/draft.tex
begin outlining patronage section

@@
 On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}.
-Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- which disproportionately
+Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- with each one of these measures painting an opposite picture of the relative strength of party ties.
+
+One thing is clear: party ties at the electoral level are durable, with most voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}.
+Noting that party affiliations are registered at the municipal level, this empirical fact aligns with qualitative evidence provided by \citet{palmeira1995comicios}, who notes that parties at the local level constitute \emph{grupos pol\'{i}ticos} (political factions) with well-defined boundaries and power disputes.
+This relative stability of party ties at the electorate level for the minority of voters who are registered with a parties suggests a distinct dynamic tying an elite group of party members to the city hall.
+
+\subsection{Municipal employment and patronage in Brazil}
 

### 2021-01-15 paper/draft.tex
add municipal employment description

@@
 There is an ongoing debate on the strength of partisan ties in Brazil.
-On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citet{desposato2006parties,ames2002deadlock}.
+On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citep{desposato2006parties,ames2002deadlock}.
 On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}.
@@
 
-\section{Descriptive Statistics}
+Municipal employment in Brazil is under local jurisdiction, with personnel appointments under the exclusive authority of the executive branch.
+Salaries, contract modalities and terminations are also under municipal jurisdiction.
+There are multiple forms of contract available, a permanent contract (\emph{estatut\'{a}rio}), a regular contract (\emph{CLT}) and temporary hires.
+Personnel expenditures cannot exceed a ceiling of 60 percent of the local budget, but as long as the ceiling is respected, other levels of government cannot interfere with local personnel decisions.
+
+Municipal discretion leaves ample room for patronage, whether it be to support mayoral coalitions as outlined in the first paper of this dissertation or to reward contributors and party loyalists \citep{colonnelli2018patronage,brollo2017victor}.
+These politically motivated appointments are particularly prominent in high-level positions, the so called \emph{cargos de confian\,{c}a}, local ministerial positions that are both high in compensation and provide access to decisions over key public services such as transportation and health.
+
+\section{Descriptive statistics}
 \label{sec:descriptive}

### 2021-01-18 paper/draft.tex
continue drafting municipal employment and patronage

@@
 There are multiple forms of contract available, a permanent contract (\emph{estatut\'{a}rio}), a regular contract (\emph{CLT}) and temporary hires.
-Personnel expenditures cannot exceed a ceiling of 60 percent of the local budget, but as long as the ceiling is respected, other levels of government cannot interfere with local personnel decisions.
+Personnel expenditures cannot exceed a ceiling of 60 percent of the local budget, but as long as the ceiling is not exceeded, other levels of government cannot interfere with local personnel decisions.
+
+There are no civil service laws regulating municipal employment.
+With the exception of permanent contracts, municipal employees are subject to the same labor laws as private sector workers.
+The lack of a civil service system at the local level has been the subject of extensive research \citep{souza2004governos}, and the high turnover that are associated with municipal employment have been widely documented \citep{akhtari2017political}.
+Labor unions do exist, in particular in the educational sector, but these are regional in focus and concentrated in metropolitan areas.
+According to the latest education census (\emph{Censo do Magist\'{e}rio}), approximately 11 percent of educational staff was unionized.
 
@@
 These politically motivated appointments are particularly prominent in high-level positions, the so called \emph{cargos de confian\,{c}a}, local ministerial positions that are both high in compensation and provide access to decisions over key public services such as transportation and health.
+Regarding the nexus between party membership and public sector employment, \citet{brollo2017victor} finds that party members who are politically aligned with the winning mayor are 30 percent more likely to receive public sector employment than their runner-up counterparts.
+
+These empirical findings suggest a clear political nexus between employment into local bureaucracies and political motives in Brazilian municipalities.
+This paper aims to provide an extensive treatment of the differences in the observable characteristics of party members - as opposed to their non-party members.
 

### 2021-01-18 paper/draft.tex
finalize characterization of municipalities

@@
 
-\subsection{Partisan affiliation in Brazil}
+\subsection{Brazilian municipal government}
 
-Partisanship in Brazil is voluntary and widespread, with over 11 percent of registered voters affiliated to a party \citep{speck2015estudo}\footnote{For context, in most OECD countries party registration does not exceed 5 percent of the electorate.
-See \citet{biezen2014decline}.}.
-Registration occurs in the following sequence: a voter reports to a local party office, and the party officials then register the voter officially through the \emph{Tribunal Regional Eleitoral} (Regional Electoral Office).
-This registration is then centralized by the \emph{Tribunal Superior Eleitoral} (Supreme Electoral Office) and updated accordingly.
-If there are overlapping registrations, former ones become annulled and are reported as irregular to local party officials.
-Each voter is therefore only allowed to register for a single party, without any ceilings or floors regarding the duration of this affiliation.
+Brazil is a federal republic comprised of 26 states and over 5500 municipalities.
+Each municipality is headed by a mayor and city councilors (\emph{vereadores}).
+Local officials are elected for a four-year term, with reelection.
+Each local election takes place at the same time, in October, with the new administration taking office in January of the following year.
+With decentralization embedded in the enactment of a new constitution in 1988, local municipalities were given large autonomy with respect to public services such as education and healthcare, as well as building an administrative infrastructure to oversee its daily operations \citep{souza2017modernizaccao}.
 
-There is an ongoing debate on the strength of partisan ties in Brazil.
-On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citep{desposato2006parties,ames2002deadlock}.
-On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}.
-Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- with each one of these measures painting an opposite picture of the relative strength of party ties.
+Municipal budgets are financed with a mix of federal transfers and local revenues.
+Smaller municipalities tend to rely on federal transfers, which are subject to oversight by higher levels of government, either federal or state, but in practice are under large discretion by municipal governments.
+The past two decades have seen a growing share of expenditures social service concentrated by municipalities, which has led to expansion in access to and the quality of local services \citep{arretche2015trajetorias}.
+Additionally, laws and regulations concerning local economies and society are largely autonomous and instituted by the mayor's office, subject to revision and approval by the local city council \citep{brelaz2013processo}.
 
-One thing is clear: party ties at the electoral level are durable, with most voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}.
-Noting that party affiliations are registered at the municipal level, this empirical fact aligns with qualitative evidence provided by \citet{palmeira1995comicios}, who notes that parties at the local level constitute \emph{grupos pol\'{i}ticos} (political factions) with well-defined boundaries and power disputes.
-This relative stability of party ties at the electorate level for the minority of voters who are registered with a parties suggests a distinct dynamic tying an elite group of party members to the city hall.
+Finally, it municipalities are in their vast majority quite small, with approximately 90 percent of municipalities having a population of less than 50 thousand people.
+In contrast, the 27 state capitals (including Bras\'{i}lia) concentrate over 23 percent of the Brazilian population.\footnote{See coverage \href{https://www.google.com/search?q=tamanho+municipio+brasil+poulacao&source=lmns&hl=en&sa=X&ved=2ahUKEwjyiLjLh6buAhXFn-AKHbzEBxAQ_AUoAHoECAEQAA}{here}.} This largely uneven concentration of the Brazilian population across its municipal governments, as well as the prevalence of small, poorer municipalities, has led observants to conclude that municipal politics is often characterized by clientelism, with local political elites concentrating power through the strategic use of public resources \citep{leal2012coronelismo}.
 
@@
 
+\subsection{Partisan affiliation in Brazil}
+
+Partisanship in Brazil is voluntary and widespread, with over 11 percent of registered voters affiliated to a party \citep{speck2015estudo}\footnote{For context, in most OECD countries party registration does not exceed 5 percent of the electorate.
+See \citet{biezen2014decline}.}.
+Registration occurs in the following sequence: a voter reports to a municipal party office, and the party officials then register the voter officially through the \emph{Tribunal Regional Eleitoral} (Regional Electoral Office).
+This registration is then collected and centralized by the \emph{Tribunal Superior Eleitoral} (Supreme Electoral Office) and updated accordingly.
+If there are overlapping registrations, former ones become annulled and are reported as irregular to local party officials.
+Each voter is therefore only allowed to register for a single party, without any ceilings or floors regarding the duration of this affiliation.
+
+There is an ongoing debate on the strength of partisan ties in Brazil.
+On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citep{desposato2006parties,ames2002deadlock}.
+On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}.
+Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- with each one of these measures painting an opposite picture of the relative strength of party ties.
+
+One thing is clear: party ties at the electoral level are durable, with most voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}.
+Noting that party affiliations are registered at the municipal level, this empirical fact aligns with qualitative evidence provided by \citet{palmeira1995comicios}, who notes that parties at the local level constitute political factions (\emph{grupos pol\'{i}ticos} ) with well-defined boundaries and power disputes.
+This relative stability of party ties at the electorate level for the minority of voters who are registered with a parties suggests a distinct dynamic tying an elite group of party members to the city hall.
+
+\section{Data}
+
 \section{Descriptive statistics}

### 2021-01-20 paper/draft.tex
finalize context and data for party membership and employment

@@
 \section{Data}
+\label{sec:data}
+
+\subsection{Municipal and Private Sector Employment}
+
+Data on formal sector employment is collected by the Department of Labor in Brazil through a census instrument, \emph{RAIS}.
+This instrument is completed electronically by all formally registered companies regardless of workforce size or capital, subject to sanctions if evidence of misreporting is found.\footnote{See coverage and description \href{https://agenciabrasil.ebc.com.br/economia/noticia/2020-03/prazo-para-entrega-da-rais-comeca-hoje-e-vai-ate-17-de-abril}{here}.} In total, over 97 percent of companies in the formal sector are included in the \emph{RAIS}, as well as all public sector employment, including federal, state and municipal level bureaucrats.
+The quality of the data is subject to constant review by the Department of Labor, which relies on accurate information to calculate taxes and retirements benefits for the Brazilian workforce.
+
+The \emph{RAIS} data is structured as follows.
+Each row corresponds to a job associated to a worker, which may or may not appear more than once in the dataset -- since workers can hold multiple jobs at the same time.
+Each worker is tagged by a unique national id, the \emph{CPF}, which allows for following workers across sectors and over time.
+Additionally, the identified \emph{RAIS} contains the names and dates of birth of workers.\footnote{Access to the identified \emph{RAIS} was generously provided by the Department of Economics at Princeton University, hosted at the Industrial Relations office.} These are the unique primary keys that make possible join this dataset to other sources of data.
+Other studies have leveraged similar join approaches to infer employment benefits of campaign contributors \citep{colonnelli2018patronage}, as well as whether party members receive public sector employment if they belong to the same party as the winning mayor \citep{brollo2017victor}.
+
+In contrast to these previous studies, I leverage the rich dataset of covariates that includes, \emph{inter alia}, wage compensation, education, work experience and type of contract, to estimate differential returns to compensation, outlined in greater detail in Section \ref{sec:empirical}.
+The richness of \emph{RAIS} allows for a complete radiography of the public sector, revealing how compensation structures are altered by their intersection with patronage dynamics.
+In particular, because of the panel nature of its structure, researchers can follow bureaucrats before and after their public sector employment spell, providing an unprecedently accurate depiction of the municipal revolving door.
+
+\subsection{Partisan affiliation}
+
+Data on partisan affiliation is detailed as well.
+The \emph{TSE}, the national electoral authority, collects data on all party members in the national territory, including commencement and termination dates, affiliated party, municipality and name of the registered party member, as well as their electoral title.\footnote{The data is available \href{https://filia-consulta.tse.jus.br/}{here}.} Parties are mandated by electoral law to report all party members as well as providing updated registry information throughout the year.
+Parties have municipal offices in which citizens may update their registry, and these party offices in turn report to the state electoral authority, which relays this information to national electoral authorities.
+If irregularities are found parties may be sanctioned by the \emph{TSE}.
+
+The party affiliation data is structured as a rolling registry (\emph{fita espelho}), in which existing and expired party memberships are preserved.
+Each row therefore corresponds to a particular party member entry, containing the state, municipality, electoral zone and date of the registration.
+Additionally, the registry indicates the status of the membership, whether it is active (\emph{regular}) or canceled (\emph{cancelado/desfiliado}), as well as the date of the change in the membership status.
+Individual identifiers are provided, including the name and a nationally unique electoral title (\emph{t\'{i}tulo eleitoral}) that allows researchers to join names contained in the \emph{RAIS} and party affiliation data.
+
+Note that party affiliation in Brazil is high, with over 10 percent of nationally registered voters affiliated to a political party \citep{speck2015estudo}.
+Partisan ties at the electorate level are stable, in contrast with more strategic party switching occurs at the political candidate level \citep{desposato2006parties}.
+As of 2019, the year for which party affiliation data was extracted, there were over 11 million active party affiliates, with around 5 million registries either canceled or terminated by party members.
 

### 2021-01-25 paper/draft.tex
include maps into paper

@@
 
+In this section, I provide descriptive statistics that illustrate broader patterns in the distribution of party affiliated bureaucrats at the municipal level in Brazil.
+Figure \ref{fig:map_pooled} illustrates the subnational distribution of party affiliation in Brazil for the year of 2015.
+There are a few patterns that emerge.
+
+First, that the distribution of partisan affiliation varies significantly over the national territory, with its predominance in the Northeast, as well as in the Northern regions of the country.
+The Midwest and Southeast appear to have the lowest prevalence of partisan affiliation.
+Second, that there is wide variation with respect to the prevalence of partisanship.
+In particular, some municipalities are almost entirely occupied by party members, with over 60 percent of its workforce affiliated to some party.
+
+\begin{figure}
+    \includegraphics{figures/maps/pooled.png}
+    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}.
+Darker colors indicate a larger degree of partisan affiliation.
+This visualization includes all bureaucrats, from high-level managers to service staff.}
+    \label{fig:map_pooled}
+\end{figure}
+
 \section{Empirical strategy}

### 2021-01-25 source/data/modules/database/generate_municipal_tables.R

@@ -37,7 +37,7 @@ rais_mun_non_partisan <- dbGetQuery(
     rais.id_employee = filiado_mun.id_employee
     WHERE rais.nat_jur = 1031 AND
     filiado_mun.id_employee IS NULL
-    GROUP BY rais.cod_ibge_6, rais.year
+    GROUP BY rais.cod_ibge_6, rais.year, SUBSTR(rais.cbo_02, 1, 1)
     "
 )
 
@@ -61,13 +61,17 @@ rais_mun_partisan <- dbGetQuery(
     WHERE rais.nat_jur = 1031 AND
     CAST(rais.year AS INT) >= filiado_mun.year_start AND
     CAST(rais.year AS INT) <= filiado_mun.year_termination
-    GROUP BY rais.cod_ibge_6, rais.year
+    GROUP BY rais.cod_ibge_6, rais.year, SUBSTR(rais.cbo_02, 1, 1)
     "
 )
 
 message("write out tables")
 export <- list(
-    object = list(rais_mun_partisan, rais_mun_non_partisan),
+    object = list(rais_mun_partisan, rais_mun_non_partisan) %>%
+        map(
+            ~ rename(., occupation = `SUBSTR(cbo_02, 1, 1)`) %>%
+                fix_occupation()
+        ),
     file = c("rais_mun_partisan.csv.gz", "rais_mun_non_partisan.csv.gz")
 )
 

### 2021-01-25 source/data/modules/database/generate_partisanship.R

@@ -41,7 +41,6 @@ rais_filiado <- map(
         c("occupation", "contract", "education")
     )
 
-gg_ma
 # fix grouping columns
 rais_filiado <- rais_filiado %>%
     modify_at(

### 2021-01-25 source/descriptive/modules/mapping_partisanship.R

@@ -82,7 +82,6 @@ map_br_disaggregated <- party_dominance_disaggregated %>%
         )
     )
 
-
 map_partisanship <- map_br_partisan %>%
     gg_map(partisanship)
 
@@ -108,9 +107,9 @@ map_education <- map_br_disaggregated %>%
     facet_wrap(. ~ edu_category)
 
 export_maps <- list(
-    plot = list(map_occupation, map_contract, map_education),
+    plot = list(map_partisanship, map_occupation, map_contract, map_education),
     filename = sprintf(
-        here("paper/figures/maps/%s.png"), names(party_dominance)
+        here("paper/figures/maps/%s.png"), c("pooled", names(party_dominance))
         )
 )
 

### 2021-01-25 paper/draft.tex
generate pooled map of partisanship brazil

@@
 Second, that there is wide variation with respect to the prevalence of partisanship.
-In particular, some municipalities are almost entirely occupied by party members, with over 60 percent of its workforce affiliated to some party.
+In particular, some municipalities are almost entirely occupied by party members, with over 60 percent of its workforce affiliated to a party.
+Others are largely autonomous from party members, with less than 10 percent of municipal workers affiliated to a party.
 
-\begin{figure}
+\begin{figure}[H]
+    \centering
     \includegraphics{figures/maps/pooled.png}

### 2021-01-25 source/descriptive/modules/mapping_partisanship.R

@@ -115,5 +115,7 @@ export_maps <- list(
 
 pwalk(
     export_maps,
-    ggsave
+    ggsave,
+    height = 5,
+    width = 5
 )

### 2021-01-26 source/data/modules/database/generate_municipal_tables.R
export data for partisanship

@@ -24,6 +24,7 @@ rais_mun_non_partisan <- dbGetQuery(
     SELECT
         rais.cod_ibge_6,
         rais.year,
+        SUBSTR(rais.cbo_02, 1, 1),
         AVG(hired) AS hired,
         AVG(fired) AS fired,
         AVG(age) AS age,
@@ -47,6 +48,7 @@ rais_mun_partisan <- dbGetQuery(
     SELECT
         rais.cod_ibge_6,
         rais.year,
+        SUBSTR(rais.cbo_02, 1, 1),
         AVG(hired) AS hired,
         AVG(fired) AS fired,
         AVG(age) AS age,
@@ -69,7 +71,7 @@ message("write out tables")
 export <- list(
     object = list(rais_mun_partisan, rais_mun_non_partisan) %>%
         map(
-            ~ rename(., occupation = `SUBSTR(cbo_02, 1, 1)`) %>%
+            ~ rename(., occupation = `SUBSTR(rais.cbo_02, 1, 1)`) %>%
                 fix_occupation()
         ),
     file = c("rais_mun_partisan.csv.gz", "rais_mun_non_partisan.csv.gz")

### 2021-02-28 source/data/modules/database/create_rais_sql.R
fix party membership data

@@ -29,7 +29,7 @@ source(
 
 # data --------------------------------------------------------------------
 cpi <- fread(
-  here("data/raw/cpi.csv"),
+  here("data/raw/cpi/cpi.csv"),
   colClasses = c("year" = "numeric")
 )
 

### 2021-02-28 source/descriptive/modules/party_membership.R

@@ -18,15 +18,8 @@ rais_sql <- tbl(rais_con, "rais") %>%
     
 filiado_sql <- tbl(rais_con, "filiado_mun")
 
-rais_tables <- c(
-    partisan = "rais_mun_partisan",
-    non_partisan = "rais_mun_non_partisan"
-)
-
-rais_filiado <- map_dfr(
-    rais_tables,
-    ~ dbGetQuery(sprintf("SELECT * FROM %s", .)),
-    .id = "is_filiado"
+filiado <- fread(
+    here("data/raw/tse/filiado.csv.gz")
 )
 
 # create entries and exits per year

### 2021-03-03 source/descriptive/build.R
generate new visualizations of dollar amounts

@@ -4,7 +4,8 @@
 build_from_source("descriptive", "requirements.R")
 
 module_selection <- c(
-    # "party_membership",
+    "party_membership",
+    "party_contribution",
     "bureau_partisanship",
     "prior_to_bureaucy",
     "entry_into_bureaucracy",

### 2021-03-03 source/descriptive/modules/party_donation.R

@@ -1,0 +1,53 @@
+# ==============================================================================
+# input: party affiliation and campaign donation
+# output: compare donations for partisans and non-partisans
+# ==============================================================================
+source(
+    here::here("source/descriptive/modules/globals.R")
+)
+
+source(
+    here("source/descriptive/modules/requirements.R")
+)
+
+# ==============================================================================
+# set up
+# ==============================================================================
+# only electoral title available for year 2008
+campaign <- fread(
+    here("data/raw/tse/receita_candidatos_2008.csv.gz"),
+    encoding = "Latin-1"
+) %>%
+ janitor::clean_names()
+
+filiado <- fread(
+    here("data/raw/tse/filiado.csv.gz")
+)
+
+# ==============================================================================
+# visualize breakdown of donations by partisan vs. non-partisan
+# ==============================================================================
+filiado_active <- filiado %>% 
+    fix_year_filiado() %>%
+    filter_active_filiado(2008)
+
+campaign_filiado <- campaign %>% 
+    filter(
+        ds_titulo == "Recursos de pessoas físicas"
+    ) %>%
+    transmute(
+        ds_nr_titulo_eleitor = as.character(ds_nr_titulo_eleitor) %>%
+            str_pad(., 12, "left", "0"),
+        ds_titulo,
+        vr_receita
+    ) %>%
+    left_join(
+        filiado_active %>% mutate(is_filiado = 1),
+        by = c("ds_nr_titulo_eleitor" = "electoral_title")
+    )
+    
+campaign_filiado %>%
+    group_by(is_filiado) %>% 
+    summarise(
+        sum(as.numeric(vr_receita), na.rm = TRUE)/n()
+    )

### 2021-03-03 source/descriptive/modules/party_membership.R

@@ -24,50 +24,77 @@ filiado <- fread(
 
 # create entries and exits per year
 # segmented by employment
-filiado_unemployed <- filiado_sql %>%
-    anti_join(
-        rais_sql %>%
-            select(cod_ibge_6, id_employee),
-        by = c("cod_ibge_6", "id_employee")
-    ) %>%
-    collect()
+# filiado_unemployed <- filiado_sql %>%
+#     anti_join(
+#         rais_sql %>%
+#             select(cod_ibge_6, id_employee),
+#         by = c("cod_ibge_6", "id_employee")
+#     ) %>%
+#     collect()
 
-filiado_employed <- filiado_sql %>%
-    inner_join(
-        rais_sql %>%
-            select(cod_ibge_6, id_employee),
-        by = c("cod_ibge_6", "id_employee")
-    ) %>%
-    collect() %>%
-    distinct(filiado_id, .keep_all = TRUE)
+# filiado_employed <- filiado_sql %>%
+#     inner_join(
+#         rais_sql %>%
+#             select(cod_ibge_6, id_employee),
+#         by = c("cod_ibge_6", "id_employee")
+#     ) %>%
+#     collect() %>%
+#     distinct(filiado_id, .keep_all = TRUE)
 
 # compare duration of party membership between employed vs. unemployed
-filiado_turnover <- list(
-    filiado_employed,
-    filiado_unemployed
-) %>%
-    set_names(c("employed", "unemployed")) %>%
-    map_dfr(
-        . %>%
-            convert_to_date() %>%
-            pivot_longer(
-                c(date_start, date_end),
-                names_prefix = "date_",
-                names_to = "date"
-            ) %>%
-            count(date, value),
-        .id = "is_employed"
+# filiado_turnover <- list(
+#     filiado_employed,
+#     filiado_unemployed
+# ) %>%
+#     set_names(c("employed", "unemployed")) %>%
+#     map_dfr(
+#         . %>%
+#             convert_to_date() %>%
+#             pivot_longer(
+#                 c(date_start, date_end),
+#                 names_prefix = "date_",
+#                 names_to = "date"
+#             ) %>%
+#             count(date, value),
+#         .id = "is_employed"
+#     )
+
+# first determine how many party members there are per annum
+filiado_date <- filiado %>% 
+    transmute(
+        date_start, 
+        date_end = coalesce(date_end, date_cancel)
+     ) %>%
[…]

### 2021-03-03 source/descriptive/modules/requirements.R

@@ -99,6 +99,17 @@ convert_to_date <- function(data) {
     )
 }
 
+fix_year_filiado <- function(data){
+  data %>%
+  mutate(
+        across(
+            starts_with("date"), 
+            ~ str_extract(., "^\\d{4}") %>%
+                as.integer
+        )
+    )
+}
+
 fix_edu <- function(data){
   data %>%
   mutate(
@@ -162,12 +173,11 @@ fix_occupation <- function(data) {
     select(-occupation)
 }
 
-filter_active_filiado <- function(data) {
+filter_active_filiado <- function(data, year) {
   data %>%
-    generate_year_filiado() %>%
     filter(
-      year >= year_date_start &
-        year <= year_date_end
+      year >= date_start &
+        (year <= date_end|is.na(date_end))
     )
 }
 

### 2021-03-05 source/descriptive/modules/party_donation.R
fix partisanship data

@@ -49,5 +49,5 @@ campaign_filiado <- campaign %>%
 campaign_filiado %>%
     group_by(is_filiado) %>% 
     summarise(
-        sum(as.numeric(vr_receita), na.rm = TRUE)/n()
+        sum(as.numeric(vr_receita), na.rm = TRUE)
     )

### 2021-03-08 source/descriptive/modules/party_donation.R
visualize campaign contribution

@@ -47,6 +47,9 @@ campaign_filiado <- campaign %>%
     )
     
 campaign_filiado %>%
+    replace_na(
+        list(is_filiado = 0)
+    ) %>%
     group_by(is_filiado) %>% 
     summarise(
         sum(as.numeric(vr_receita), na.rm = TRUE)

### 2021-03-09 paper/draft.tex
fix party membership data
generate new visualizations of dollar amounts
incorporate info on party affiliation

@@
 
+Additionally, it is worthwhile noting that the affiliation of party members has been increasing over time.
+Importantly, this evolution has been rather rapid over time.
+While in the year 1997 there were around 200 thousand active party members, by the year 2018 this has increased to 10 million.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = 10cm, height = 10cm]{figures/partisanship/plot_partisan_by_year.pdf}
+    \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
+    \label{fig:map_pooled}
+\end{figure}
+
+Another interesting feature to note in the data is that there are discontinuities in the number of party members.
+The dashed lines indicate the year of a local election in Brazil, such as 2002 or 2018.
+The jumps in party affiliation are expressive, in particular for earlier electoral cycles.
+In particular, for the electoral year of 2002, there were around 2 million new party members registered as a result of that particular local election.
+While these jumps have decreased in more recent elections, it is an empirical regularity that highlights the electoral motivation driving party affiliation in Brazil.
+
 \section{Descriptive statistics}
@@
 
+\emph{perform an Oaxaca decomposition to understand the variant compensations}
+
 \section{Conclusion}

### 2021-03-09 source/descriptive/modules/party_donation.R

@@ -46,11 +46,28 @@ campaign_filiado <- campaign %>%
         by = c("ds_nr_titulo_eleitor" = "electoral_title")
     )
     
-campaign_filiado %>%
+campaign_filiado_total <- campaign_filiado %>%
     replace_na(
         list(is_filiado = 0)
     ) %>%
     group_by(is_filiado) %>% 
     summarise(
-        sum(as.numeric(vr_receita), na.rm = TRUE)
+        total_contribution = sum(as.numeric(vr_receita)/1e6, na.rm = TRUE)
+    )
+
+campaign_filiado_total %>%
+    mutate(
+        is_filiado = recode(is_filiado, `0` = "non-partisan", `1` = "partisan")
+    ) %>%
+    ggplot(
+        aes(is_filiado, total_contribution, fill = is_filiado)
+    ) +
+    geom_col(
+        width = 0.5
+    ) +
+    scale_y_continuous(breaks = scales::pretty_breaks()) +
+    xlab("") +
+    ylab("Total contribution (million reais)") +
+    ggsave(
+        here("paper/figures/plot_contribution.pdf")
     )

### 2021-03-09 source/descriptive/modules/party_membership.R

@@ -92,7 +92,15 @@ filiado_active_year %>%
     ) +
     labs(
         x = "Year",
-        y = "Total active party members"
+        y = "Total active party members (millions)"
+    ) +
+    geom_vline(
+        xintercept = seq(2002, 2018, 4),
+        linetype = "dashed"
+    ) +
+    scale_y_continuous(
+        breaks = c(0, 5e6, 1e7),
+        labels = c(0, 5, 10)
     ) +
     ggsave(
         here("paper/figures/partisanship/plot_partisan_by_year.pdf")

### 2021-03-10 paper/draft.tex
ad who benefits

@@
 Each worker is tagged by a unique national id, the \emph{CPF}, which allows for following workers across sectors and over time.
-Additionally, the identified \emph{RAIS} contains the names and dates of birth of workers.\footnote{Access to the identified \emph{RAIS} was generously provided by the Department of Economics at Princeton University, hosted at the Industrial Relations office.} These are the unique primary keys that make possible join this dataset to other sources of data.
+Additionally, the identified \emph{RAIS} contains the names and dates of birth of workers.\footnote{Access to the identified \emph{RAIS} was generously provided by the Department of Economics at Princeton University, hosted at the Industrial Relations Office.} These are the unique primary keys that make possible join this dataset to other sources of data.
 Other studies have leveraged similar join approaches to infer employment benefits of campaign contributors \citep{colonnelli2018patronage}, as well as whether party members receive public sector employment if they belong to the same party as the winning mayor \citep{brollo2017victor}.
@@
 
+\subsection{Who benefits from patronage?}
+
+Who should politicians buy off: rich or poor voters?
+There are competing views on who benefits from patronage.
+On the one hand, from a politician's perspective it would be cheaper to buy voters who are poor, insofar as the return on investment would be greater for poorer voters, who need and value more monetary incentives \citet{stokes2013brokers}.
+Other scholars argue that in order to maximize electoral returns, politicians target swing voters, those who are indifferent to either party and therefore, again, require the least resources to buy off \citet{dixit1996determinants}.
+
+On the other, there are good reasons to expect politicians to target rich voters as well. There is a well-established literature on money and politics,
+
 \section{Descriptive statistics}

### 2021-03-10 source/descriptive/modules/party_donation.R

@@ -24,10 +24,24 @@ filiado <- fread(
     here("data/raw/tse/filiado.csv.gz")
 )
 
+bureaucrat_entry <- tbl(rais_con, "rais_bureaucrat_entry")
+
+filiado_bureaucrat <- tbl(rais_con, "filiado_mun") %>%
+    select(filiado_id, id_employee, year_start, year_termination) %>%
+    filter(year_start <= 2008 & year_termination > 2008) %>%
+    inner_join(
+        bureaucrat_entry,
+        by = "id_employee"
+    ) %>%
+    collect()
+
 # ==============================================================================
 # visualize breakdown of donations by partisan vs. non-partisan
 # ==============================================================================
 filiado_active <- filiado %>% 
+    mutate(
+        filiado_id = row_number()
+    ) %>%
     fix_year_filiado() %>%
     filter_active_filiado(2008)
 

### 2021-03-10 paper/draft.tex
add image
add section on why and who benefits from patronage

@@
     \centering
-    \includegraphics[width = 10cm, height = 10cm]{figures/partisanship/plot_partisan_by_year.pdf}
+    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_by_year.pdf}
     \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
@@
 
-\subsection{Who benefits from patronage?}
+\subsection{Why engage in patronage?}
 
-Who should politicians buy off: rich or poor voters?
-There are competing views on who benefits from patronage.
-On the one hand, from a politician's perspective it would be cheaper to buy voters who are poor, insofar as the return on investment would be greater for poorer voters, who need and value more monetary incentives \citet{stokes2013brokers}.
+Why do politicians engage in patronage?
+And who should politicians buy off?
+There are competing views on who benefits from patronage, and what is motivating it in the first place.
+On the one hand, from a politician's perspective it can be cheaper to buy voters who are poor, insofar as these voters need and value more monetary incentives \citet{stokes2013brokers}.
 Other scholars argue that in order to maximize electoral returns, politicians target swing voters, those who are indifferent to either party and therefore, again, require the least resources to buy off \citet{dixit1996determinants}.
 
-On the other, there are good reasons to expect politicians to target rich voters as well. There is a well-established literature on money and politics,
+The socioeconomic status of voters is not, however, the only consideration that politicians take into place.
+Patronage can be used by politicians to reward political loyalists for resources they have contributed for an electoral campaign, a practice generally denoted as \emph{prebendism}\citep{van2007meet}.
+In Brazil, studies have found that campaign contributors are more likely to be employed in the local bureaucracy, in the magnitude of a 30 percent increase in the probability of being hired \citep{colonnelli2018patronage}.
+Additionally, there is strong evidence that employment in the bureaucracy is used as a spoil to reward party loyalists, again with evidence from Brazilian local bureaucracies \citep{brollo2017victor}.
+
+Why should patronage occur post-election?
+There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office.
+In particular, a politician cannot offer job to a voter prior to holding office.
+Moreover, public sector jobs are coarse goods in the sense that they cannot be infinitely divided among voters.
+There are tight budget and administrative constraints that distinguishes a public sector jobs from other forms of resources available to politicians.
+
+If public sector jobs are scarce, and there are only so many spots that can be filled post-election, a different form of consideration takes place from the standpoint of the politician.
+Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, constituencies can vary widely with respect to the resources that they bring in to the table.
+A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}.
+What is overseen is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.
+
+This is in fact what the data shows.
 

### 2021-03-10 paper/draft.tex
add additional info on wages

@@
 
-\subsection{Why engage in patronage?}
+\subsection{Locking-In Wealthy Patrons}
 
@@
 A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}.
-What is overseen is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.
+What is underemphasized is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.
 
-This is in fact what the data shows.
+This is in fact what the data suggests.
+First, on average party members who enter the bureaucracy are wealthier than their non-partisan counterparts, in particular at the moment that the mayor assumes office.
+With the RAIS data, it is possible to identify for each and every employee the last wage in the non-municipal sector prior to public sector employment.
+Figure \ref{fig:wage_differential} plots the median of the last non-municipal sector wage by partisanship status.
+The gap between partisan and non-partisan wages is the highest for the beginning of the mayoral term, which is often associated with more intense patronage activity.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_median_wage.pdf}
+    \caption{\textbf{Wage Gap between Partisans and Non-Partisans}. The graph tracks the time trend in the median wage }
+    \label{fig:wage_differential}
+\end{figure}
+
+Additionally
 
@@
     \centering
-    \includegraphics{figures/maps/pooled.png}
+    \includegraphics[width = 10cm, height = 8cm]{figures/maps/pooled.png}
     \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}.

### 2021-03-10 source/descriptive/modules/prior_to_bureaucracy.R

@@ -27,6 +27,15 @@ career_filiado <- career_filiado %>%
         )
     )
 
+# aggregate
+career_filiado_year <- career_filiado %>%
+    group_by(year, is_partisan) %>%
+    summarise(
+        across(
+            c(ends_with("mean"), median_wage),
+        ~weighted.mean(., n)
+        )
+    )
 # ---------------------------------------------------------------------------- #
 message("generating descriptive plots")
 
@@ -42,12 +51,11 @@ plot_summary <- map2(
     vars_to_plot,
     label_vars,
     ~ ggplot(
-        data = career_filiado,
+        data = career_filiado_year,
         aes_string(x = "year", y = .x, color = "is_partisan")
     ) +
-    geom_smooth(
-        span = 3
-    ) +
+    geom_point() +
+    geom_line() +
     scale_colour_discrete(
         name = "Partisanship",
         labels = c(
@@ -56,6 +64,7 @@ plot_summary <- map2(
             "non_partisan" = "non-partisan"
         )
     ) +
+    geom_vline(xintercept = seq(2005, 2013, 4), lty = "dotted") +
     labs(x = "Year", y = .y)
 )
 

### 2021-03-10 paper/draft.tex
add employment spells

@@
     \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_median_wage.pdf}
-    \caption{\textbf{Wage Gap between Partisans and Non-Partisans}. The graph tracks the time trend in the median wage }
+    \caption{\textbf{Wage Gap between Partisans and Non-Partisans}. The graph tracks the time trend in the median wage of employees prior to entering the municipal bureaucracy.}
     \label{fig:wage_differential}
@@
 
-Additionally
+Additionally, these same workers tend to have a longer experience working in the formal sector prior to employment.
+Combined with the higher compensations, the employment evidence suggests that that party members who first enter the bureaucracy in the initial years of the mandate belong to a wealthier strata than their non-partisan counterparts.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_work_experience_mean.pdf}
+    \caption{\textbf{Work Experience Differential between Partisans and Non-Partisans}. Partisan bureaucrats tend to arrive from longer work tenures in the non-municipal sectors.}
+    \label{fig:work_experience}
+\end{figure}
 

### 2021-03-15 paper/draft.tex
test run data donations

@@
 
-Why should patronage occur post-election?
+The timing of clientelistic transfers ie key.
 There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office.
@@
 If public sector jobs are scarce, and there are only so many spots that can be filled post-election, a different form of consideration takes place from the standpoint of the politician.
-Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, constituencies can vary widely with respect to the resources that they bring in to the table.
+Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, members of the constituency can vary widely with respect to the resources that they bring in to the table.
 A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}.
-What is underemphasized is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.
+What is underemphasized is the fact that politicians can use patronage to lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.
 
@@
 Combined with the higher compensations, the employment evidence suggests that that party members who first enter the bureaucracy in the initial years of the mandate belong to a wealthier strata than their non-partisan counterparts.
+They have experienced both longer bouts of employment in the formal sector, as well as higher compensations in their tenure.
+While it is clear that these hires are a reward for previous contributions, it is not clear whether a party member who is hired into the bureaucracy is more likely to donate to their party.
+\emph{How can I show that party members who are hired into the bureaucracy are more likely to donate?}
 
@@
 
+In sum, employment evidence seems to suggest that party members who enter the bureaucracy in the first years of mandate come predominantly from
+
 \section{Descriptive statistics}
@@
 There are a few patterns that emerge.
+The first one is that there is
 

### 2021-03-15 source/descriptive/modules/party_donation.R

@@ -35,6 +35,10 @@ filiado_bureaucrat <- tbl(rais_con, "filiado_mun") %>%
     ) %>%
     collect()
 
+filiado_mun <- fread(
+    here("data/clean/id/filiado_with_id_employee_mun.csv.gz")
+)
+
 # ==============================================================================
 # visualize breakdown of donations by partisan vs. non-partisan
 # ==============================================================================
@@ -50,14 +54,14 @@ campaign_filiado <- campaign %>%
         ds_titulo == "Recursos de pessoas físicas"
     ) %>%
     transmute(
-        ds_nr_titulo_eleitor = as.character(ds_nr_titulo_eleitor) %>%
+        electoral_title = as.character(ds_nr_titulo_eleitor) %>%
             str_pad(., 12, "left", "0"),
         ds_titulo,
         vr_receita
     ) %>%
     left_join(
         filiado_active %>% mutate(is_filiado = 1),
-        by = c("ds_nr_titulo_eleitor" = "electoral_title")
+        by = c("electoral_title")
     )
     
 campaign_filiado_total <- campaign_filiado %>%

### 2021-03-16 source/descriptive/modules/party_donation.R
no way of knowing data from party members

@@ -89,3 +89,34 @@ campaign_filiado_total %>%
     ggsave(
         here("paper/figures/plot_contribution.pdf")
     )
+
+# ==============================================================================
+# test whether we can identify workers
+# ==============================================================================
+# can we make a one-to-one join?
+filiado_mun_first <- filiado_mun[order(date_start)] %>%
+    .[, .SD[1], id_employee]
+
+# eliminate defective entries
+filiado_mun_first <- filiado_mun_first %>%
+    filter(
+        str_detect(date_start, "^199|^20")
+    )
+
+# extract only if there is a unique record per date key
+filiado_unique <- filiado %>%
+    select(electoral_title, cod_ibge_6, party, starts_with("date")) %>%
+    filter(
+        str_detect(date_start, "^199|^20")
+    ) %>%
+    group_by(cod_ibge_6, party, date_start, date_end, date_cancel) %>%
+    mutate(n = n()) %>%
+    ungroup() %>%
+    filter(n == 1) %>%
+    select(-n)
+
+# able to recover around 21% of filiados
+filiado_with_id_employee <- filiado_mun_first %>%
+    merge(filiado_unique, all.x = TRUE)
+
+# look within the ones who are party members

### 2021-03-17 paper/draft.tex
no way of knowing data from party members
plot summary partisanship

@@
 
-\subsection{Locking-In Wealthy Patrons}
+\subsection{Wealthy Patrons}
 
@@
 
-The socioeconomic status of voters is not, however, the only consideration that politicians take into place.
+The relative cost of voters is not, however, the only consideration that politicians take into place.
 Patronage can be used by politicians to reward political loyalists for resources they have contributed for an electoral campaign, a practice generally denoted as \emph{prebendism}\citep{van2007meet}.
@@
 
-The timing of clientelistic transfers ie key.
+The timing of clientelistic transfers is key.
 There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office.
@@
 
-This is in fact what the data suggests.
+While proving the second point is beyond the scope of this paper, there is strong evidence that the first dynamic, that patronage disproportionately targets wealthy patrons, is true.
 First, on average party members who enter the bureaucracy are wealthier than their non-partisan counterparts, in particular at the moment that the mayor assumes office.
@@
 While it is clear that these hires are a reward for previous contributions, it is not clear whether a party member who is hired into the bureaucracy is more likely to donate to their party.
-\emph{How can I show that party members who are hired into the bureaucracy are more likely to donate?}
+In order to determine this, it would be necessary to compare party members just before and after joining the local public sector, and see whether or not they are more likely to give campaign contributions.
 
@@
 
-In sum, employment evidence seems to suggest that party members who enter the bureaucracy in the first years of mandate come predominantly from
+In sum, employment evidence seems to suggest that party members who enter the bureaucracy in the first years of mandate come predominantly from wealthier strata of the population.
+Additionally, they have been working for longer time periods in the formal sector.
+However, on average, party members who enter the bureaucracy are less educated and older than their non-partisan counterparts, as evident in figure \ref{fig:partisan_edu_et_age}.
+Finally, while it is not clear what
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = 7cm, height = 8cm]{figures/partisanship/plot_edu_mean.pdf}
+    \includegraphics[width = 7cm, height = 8cm]{figures/partisanship/plot_age_mean.pdf}
+    \caption{\textbf{Heterogeneity in Education and Age.}}
+    \label{fig:partisan_edu_et_age}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width = 10cm, height = 8cm]{figures/maps/pooled.png}
+    \includegraphics[width = 8cm, height = 8cm]{figures/maps/pooled.png}
     \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}.

### 2021-03-17 source/descriptive/modules/prior_to_bureaucracy.R

@@ -34,7 +34,8 @@ career_filiado_year <- career_filiado %>%
         across(
             c(ends_with("mean"), median_wage),
         ~weighted.mean(., n)
-        )
+        ),
+        n = n()
     )
 # ---------------------------------------------------------------------------- #
 message("generating descriptive plots")
@@ -64,6 +65,7 @@ plot_summary <- map2(
             "non_partisan" = "non-partisan"
         )
     ) +
+    ggtitle(.y) +
     geom_vline(xintercept = seq(2005, 2013, 4), lty = "dotted") +
     labs(x = "Year", y = .y)
 )

### 2021-03-18 paper/draft.tex
generate partisannship spells

@@
 
-Note that party affiliation in Brazil is high, with over 10 percent of nationally registered voters affiliated to a political party \citep{speck2015estudo}.
-Partisan ties at the electorate level are stable, in contrast with more strategic party switching occurs at the political candidate level \citep{desposato2006parties}.
+Note that party affiliation in Brazil is relatively high, with over 10 percent of nationally registered voters affiliated to a political party, as compared with other advanced democracies that hover around 5 percent \citep{speck2015estudo}.
+Partisan ties at the voter level are stable, in contrast with more strategic party switching occurs at the political candidate level \citep{desposato2006parties}.
 As of 2019, the year for which party affiliation data was extracted, there were over 11 million active party affiliates, with around 5 million registries either canceled or terminated by party members.
 
-Additionally, it is worthwhile noting that the affiliation of party members has been increasing over time.
+It is important to note that the affiliation of party members has been increasing over time.
 Importantly, this evolution has been rather rapid over time.
 While in the year 1997 there were around 200 thousand active party members, by the year 2018 this has increased to 10 million.
+While the causes of this rapid increase are not clear, it is most likely due to both increased registration and better quality data on partisan affiliation as the collection of information by the national electoral authority.
 
@@
 
-\section{Descriptive statistics}
-\label{sec:descriptive}
+\subsection{Prevalence of Patronage}
+\label{sec:prevalence_patronage}
 
-In this section, I provide descriptive statistics that illustrate broader patterns in the distribution of party affiliated bureaucrats at the municipal level in Brazil.
-Figure \ref{fig:map_pooled} illustrates the subnational distribution of party affiliation in Brazil for the year of 2015.
-There are a few patterns that emerge.
-The first one is that there is
-
-First, that the distribution of partisan affiliation varies significantly over the national territory, with its predominance in the Northeast, as well as in the Northern regions of the country.
+The distribution of partisan affiliation varies significantly over the national territory, with its predominance in the Northeast, as well as in the Northern regions of the country.
 The Midwest and Southeast appear to have the lowest prevalence of partisan affiliation.
@@
     \centering
-    \includegraphics[width = 8cm, height = 8cm]{figures/maps/pooled.png}
+    \includegraphics[width = 9cm, height = 9cm]{figures/maps/pooled.png}
     \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}.
 Darker colors indicate a larger degree of partisan affiliation.
-This visualization includes all bureaucrats, from high-level managers to service staff.}
+Sample includes all bureaucrats, from high-level managers to service staff.}
     \label{fig:map_pooled}

### 2021-03-18 source/descriptive/modules/party_membership.R

@@ -10,6 +10,8 @@ source(
     here("source/descriptive/modules/requirements.R")
 )
 
+library(lubridate)
+
 # ==============================================================================
 # generate descriptive statistics of party members
 # ==============================================================================
@@ -62,6 +64,7 @@ filiado <- fread(
 # first determine how many party members there are per annum
 filiado_date <- filiado %>% 
     transmute(
+        party,
         date_start, 
         date_end = coalesce(date_end, date_cancel)
      ) %>%
@@ -105,4 +108,27 @@ filiado_active_year %>%
     ggsave(
         here("paper/figures/partisanship/plot_partisan_by_year.pdf")
     )
-# compare membership spells
+
+# duration spells
+filiado_date_spell <- filiado %>%
+    # sample_n(100) %>%
+    transmute(
+        date_end = coalesce(date_end, date_cancel) %>%
+            if_else(
+                . == "",
+                date_record_extraction, .
+            ),
+        duration = interval(
+            as_date(date_start), as_date(date_end)
+        ) / years(1)
+    )
+
+# what we ultimately want is number of party members by year
+filiado_date_spell %>%
+    mutate(
+        duration = round(duration)
+    ) %>%
+    ggplot() +
+    geom_histogram(
+        aes(duration), binwidth = 1
+    )

### 2021-03-19 source/descriptive/modules/party_membership.R
create histogram of partisan spells

@@ -10,8 +10,6 @@ source(
     here("source/descriptive/modules/requirements.R")
 )
 
-library(lubridate)
-
 # ==============================================================================
 # generate descriptive statistics of party members
 # ==============================================================================
@@ -24,43 +22,6 @@ filiado <- fread(
     here("data/raw/tse/filiado.csv.gz")
 )
 
-# create entries and exits per year
-# segmented by employment
-# filiado_unemployed <- filiado_sql %>%
-#     anti_join(
-#         rais_sql %>%
-#             select(cod_ibge_6, id_employee),
-#         by = c("cod_ibge_6", "id_employee")
-#     ) %>%
-#     collect()
-
-# filiado_employed <- filiado_sql %>%
-#     inner_join(
-#         rais_sql %>%
-#             select(cod_ibge_6, id_employee),
-#         by = c("cod_ibge_6", "id_employee")
-#     ) %>%
-#     collect() %>%
-#     distinct(filiado_id, .keep_all = TRUE)
-
-# compare duration of party membership between employed vs. unemployed
-# filiado_turnover <- list(
-#     filiado_employed,
-#     filiado_unemployed
-# ) %>%
-#     set_names(c("employed", "unemployed")) %>%
-#     map_dfr(
-#         . %>%
-#             convert_to_date() %>%
-#             pivot_longer(
-#                 c(date_start, date_end),
-#                 names_prefix = "date_",
-#                 names_to = "date"
-#             ) %>%
-#             count(date, value),
-#         .id = "is_employed"
-#     )
-
 # first determine how many party members there are per annum
 filiado_date <- filiado %>% 
     transmute(
@@ -70,7 +31,8 @@ filiado_date <- filiado %>%
      ) %>%
     fix_year_filiado() %>%
     filter(
-        between(date_start, 1997, 2018)
+        between(date_start, 1997, 2019) &
+        between(date_end, 1997, 2019)
     )
 
 range_record <- range(filiado_date$date_start)
@@ -111,24 +73,47 @@ filiado_active_year %>%
 
 # duration spells
 filiado_date_spell <- filiado %>%
-    # sample_n(100) %>%
     transmute(
+        date_start,
         date_end = coalesce(date_end, date_cancel) %>%
             if_else(
                 . == "",
                 date_record_extraction, .
             ),
-        duration = interval(
-            as_date(date_start), as_date(date_end)
[…]

### 2021-03-19 paper/draft.tex
create histogram of partisan spells
fix code for filtering active partisans

@@
 It is important to note that the affiliation of party members has been increasing over time.
-Importantly, this evolution has been rather rapid over time.
+Figure \ref{fig:number_partisanship} illustrates this pattern.
+The growth in the number of partisan affiliates has been rather rapid over time.
 While in the year 1997 there were around 200 thousand active party members, by the year 2018 this has increased to 10 million.
@@
     \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
-    \label{fig:map_pooled}
+    \label{fig:number_partisan}
 \end{figure}
@@
 The jumps in party affiliation are expressive, in particular for earlier electoral cycles.
-In particular, for the electoral year of 2002, there were around 2 million new party members registered as a result of that particular local election.
+For instance, in the electoral year of 2002, there were around 2 million new party members registered as a result of that particular local election.
 While these jumps have decreased in more recent elections, it is an empirical regularity that highlights the electoral motivation driving party affiliation in Brazil.
+
+Regarding partisan duration spells, a few patterns emerge as in the . The spikes every four years mirror the jumps in partisanship in Figure \ref{fig:number_partisan} are mirrored
 

### 2021-03-19 source/descriptive/modules/party_membership.R

@@ -23,11 +23,15 @@ filiado <- fread(
 )
 
 # first determine how many party members there are per annum
-filiado_date <- filiado %>% 
+filiado_date <- filiado %>%
     transmute(
         party,
+        region = str_sub(cod_ibge_6, 1, 1),
         date_start, 
-        date_end = coalesce(date_end, date_cancel)
+        date_end = coalesce(date_end, date_cancel) %>%
+            if_else(
+                . == "", "2019", .
+            )
      ) %>%
     fix_year_filiado() %>%
     filter(
@@ -45,6 +49,13 @@ filiado_active_year <- map_dfr(
         tibble(
             n = .
         ) %>%
+        mutate(year = .x)
+)
+
+filiado_active_year_by_region <- map_dfr(
+    range_record,
+    ~ filter_active_filiado(year = .x, data = filiado_date) %>%
+        count(region) %>%
         mutate(year = .x)
 )
 
@@ -69,6 +80,40 @@ filiado_active_year %>%
     ) +
     ggsave(
         here("paper/figures/partisanship/plot_partisan_by_year.pdf")
+    )
+
+# growth in membership by state
+filiado_active_year_by_region %>% 
+    filter(
+        between(year, 1997, 2018)
+    ) %>%
+    mutate(
+        region = recode(
+            region,
+            `1` = "North",
+            `2` = "Northeast",
+            `3` = "Southeast",
+            `4` = "South",
+            `5` = "Midwest"
+        )
+    ) %>%
+    gg_point_line(
+        aes(year, n, color = region)
+    ) +
+    labs(
+        x = "Year",
+        y = "Total active party members (millions)"
+    ) +
+    geom_vline(
+        xintercept = seq(2002, 2018, 4),
+        linetype = "dashed"
+    ) +
+    scale_y_continuous(
+        breaks = c(0, 5e6, 1e7),
+        labels = c(0, 5, 10)
+    ) +
+    ggsave(
+        here("paper/figures/partisanship/plot_partisan_by_region.pdf")
     )
 
 # duration spells
@@ -99,12 +144,10 @@ filiado_date_spell <- filiado_date_spell %>%
         year_date_start >= 1985 &
         year_date_start <= 2019 & 
         year_date_end <= 2019
-    ) %>%
-    mutate(
-        duration = round(duration)
     )
 
 filiado_date_spell %>%
+    filter(year_date_start >= 1997) %>%
     filter(duration > 0) %>%
     ggplot() +
     geom_histogram(

### 2021-03-19 paper/draft.tex
keep editing main text

@@
 
-Regarding partisan duration spells, a few patterns emerge as in the . The spikes every four years mirror the jumps in partisanship in Figure \ref{fig:number_partisan} are mirrored
+Disaggregating the data further into regions, we find significant heterogeneities with respect to regional trends (Figure \ref{fig:number_partisan_region}).The region that concentrates the most rapid growth and the largest amount of party members is the Southeast, a densely populated and most economically developed region in Brazil.
+It is closely followed by the Northeast, a region that although relatively poor, has a been characterized by deep partisan roots that underlie patronage exchanges \citep{ansell2010auctioning}.
+The South, an affluent but predominantly rural region has similar levels of partisanship to the Northeast.
+In contrast, party membership in the Northern and Midwestern regions are low.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_by_year.pdf}
+    \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
+    \label{fig:number_partisan_region}
+\end{figure}
+
+Regarding partisan duration spells, a few patterns emerge as in the. The spikes every four years mirror the jumps in partisanship in Figure \ref{fig:number_partisan} are mirrored
 

### 2021-03-19 paper/draft.tex
include partisanship by occupation

@@
 It is important to note that the affiliation of party members has been increasing over time.
-Figure \ref{fig:number_partisanship} illustrates this pattern.
+Figure \ref{fig:number_partisan} illustrates this pattern.
 The growth in the number of partisan affiliates has been rather rapid over time.
@@
     \centering
-    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_by_year.pdf}
-    \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
+    \includegraphics[width = 7cm, height = 7cm]{figures/partisanship/plot_partisan_by_year.pdf}
+    \includegraphics[width = 7cm, height = 7cm]{figures/partisanship/plot_partisan_by_region.pdf}
+    \caption{\textbf{Number of active party members by year}.
+Pooled (left) and disaggregated by region (right).
+There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
     \label{fig:number_partisan}
@@
 
-Disaggregating the data further into regions, we find significant heterogeneities with respect to regional trends (Figure \ref{fig:number_partisan_region}).The region that concentrates the most rapid growth and the largest amount of party members is the Southeast, a densely populated and most economically developed region in Brazil.
+Disaggregating the data further into regions, we find significant heterogeneities with respect to regional trends.The region that concentrates the most rapid growth and the largest amount of party members is the Southeast, a densely populated and most economically developed region in Brazil.
 It is closely followed by the Northeast, a region that although relatively poor, has a been characterized by deep partisan roots that underlie patronage exchanges \citep{ansell2010auctioning}.
@@
 
-\begin{figure}[H]
+Finally, we present data on party membership spells in Figure \ref{fig:party_spell}.
+A few patterns emerge.
+First, the distribution of spells is rather uniform, with spikes every four years that mirror the jumps in partisanship in Figure \ref{fig:number_partisan}.
+
+\begin{figure}
     \centering
-    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_by_year.pdf}
-    \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
-    \label{fig:number_partisan_region}
+    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_spell.pdf}
+    \caption{\textbf{Party memebership spells}. Data includes all party memberships that start in 1985 - the Brazilian democratic transition - and 2019.}
+    \label{fig:party_spell}
 \end{figure}
-
-Regarding partisan duration spells, a few patterns emerge as in the. The spikes every four years mirror the jumps in partisanship in Figure \ref{fig:number_partisan} are mirrored
 
@@
     \centering
-    \includegraphics[width = 9cm, height = 9cm]{figures/maps/pooled.png}
+    \includegraphics[width = 8cm, height = 8cm]{figures/maps/pooled.png}
     \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}.
+Darker colors indicate a larger degree of partisan affiliation.
+Sample includes all bureaucrats, from high-level managers to service staff.}
+    \label{fig:map_pooled}
+\end{figure}
+
+Alongside these spatial differences, there are
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = 8cm, height = 8cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
+    \caption{\textbf{Proportion of party-affiliated members by occupation}.
 Darker colors indicate a larger degree of partisan affiliation.

### 2021-03-19 source/descriptive/modules/bureau_partisanship.R

@@ -51,7 +51,7 @@ path_to_figs <- ifelse(
 
 export_plots <- list(
     filename = sprintf(
-        here(path_to_figs, "plot_partisanship_by_%s.pdf"),
+        here(path_to_figs, "%s.pdf"),
         group_vars
         ),
     plot = plot_partisan

### 2021-03-24 paper/draft.tex
set up new pipeline for data

@@
 
-\subsection{Prevalence of Patronage}
+\subsection{Islands of Patronage}
 \label{sec:prevalence_patronage}
@@
 
-Alongside these spatial differences, there are
+Alongside these spatial differences, there are significant differences with respect to the types of positions occupied by party members.
+Figure \ref{fig:partisan_by_cbo} breaks down the distribution of party members across different levels of the bureaucracy, using CBO (occupation) classifications provided by the RAIS data.
+Bureaucrat high corresponds to high-level management positions, such as directors and managers, bureaucrat low are low-level management such as supervisors and clerks.
+Frontline providers includes all types of service providers, where frontline high includes doctors and teachers, while low are other service providers such as janitors or security guards.
+
+Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors. These are also the best
 
@@
 Sample includes all bureaucrats, from high-level managers to service staff.}
-    \label{fig:map_pooled}
+    \label{fig:partisan_by_cbo}
 \end{figure}

### 2021-03-24 source/data/modules/database/generate_municipal_tables.R

@@ -78,6 +78,6 @@ export <- list(
 )
 
 export %>%
-    pwalk(write_data, dir = "summary")
+    pwalk(write_data, dir = "data/clean/summary")
 
 message("create municipal tables complete!")

### 2021-03-24 source/data/modules/database/generate_partisanship.R

@@ -16,6 +16,8 @@ group_vars <- c(
     "SUBSTR(cbo_02, 1, 1)", "contract_type", "edu"
 )
 
+RSQLite::initExtension(rais_con)
+
 rais_query <- sprintf(
     "
     SELECT
@@ -23,7 +25,7 @@ rais_query <- sprintf(
     rais.year,
     %1$s,
     SUM(1.0) AS n,
-    AVG(COALESCE(is_filiado, 0)) AS partisanship
+    AVG(COALESCE(is_filiado, 0)) AS partisanship,
     FROM rais
     LEFT JOIN (SELECT *, 1.0 AS is_filiado FROM filiado_mun) AS filiado_mun
     ON rais.cod_ibge_6 = filiado_mun.cod_ibge_6 AND

### 2021-03-24 source/descriptive/modules/turnover_cycles.R

@@ -13,14 +13,15 @@ source(
 rais <- tbl(rais_con, "rais")
 filiado <- tbl(rais_con, "filiado_mun")
 
-rais_tables <- c(
-    partisan = "rais_mun_partisan",
-    non_partisan = "rais_mun_non_partisan"
+rais_tables <- list.files(
+    here("data/clean/summary/"),
+    pattern = "^rais",
+    full.names = TRUE
 )
 
 rais_filiado <- map_dfr(
     rais_tables,
-    ~ dbGetQuery(sprintf("SELECT * FROM %s", .)),
+    fread,
     .id = "is_filiado"
 )
 

### 2021-03-24 source/descriptive/modules/turnover_cycles.R
generate new plots distinguishing partisans from non-partisans

@@ -10,8 +10,23 @@ source(
     here::here("source/descriptive/modules/requirements.R")
 )
 
+plot_partisan_stat <- function(data, var){
+     ggplot(
+        aes_string("year", var, color = "is_filiado"),
+        data = data
+    ) +
+    stat_summary_bin(
+        fun = "mean",
+        geom = "point"
+    ) +
+    stat_summary_bin(
+        fun = "mean",
+        geom = "line"
+    ) +
+    ggtitle(var)
+}
+
 rais <- tbl(rais_con, "rais")
-filiado <- tbl(rais_con, "filiado_mun")
 
 rais_tables <- list.files(
     here("data/clean/summary/"),
@@ -23,30 +38,17 @@ rais_filiado <- map_dfr(
     rais_tables,
     fread,
     .id = "is_filiado"
-)
+) %>%
+    mutate(
+        is_filiado = recode(is_filiado, `1` = "non-partisan", `2` = "partisan")
+    )
 
 message("generate plots")
-vars_to_plot <- c("hired", "fired", "work_experience")
+vars_to_plot <- c("hired", "fired", "age", "edu", "wage", "work_experience")
 plot_turnover <- map(
     vars_to_plot,
-    ~ gg_summary(
-        x = "year",
-        y = .,
-        data = rais_filiado,
-        smooth = FALSE,
-        color = !!sym("is_filiado")
-    ) +
-        ggtitle(.)
+    ~ plot_partisan_stat(var = ., data = rais_filiado)
 )
-
-rais_filiado %>%
-    ggplot(
-        aes(year, hired, color = is_filiado)
-    ) +
-    stat_summary_bin(
-        fun = "mean",
-        geom = "point"
-    )
 
 path_to_figs <- ifelse(
     isTRUE(debug), "paper/figures/turnover/sample/", "paper/figures/turnover/"

### 2021-03-24 paper/draft.tex
partisanship

@@
 
-\subsection{Wealthy Patrons}
+\subsection{Who enters the bureaucracy?}
 
@@
     \includegraphics[width = 7cm, height = 8cm]{figures/partisanship/plot_age_mean.pdf}
-    \caption{\textbf{Heterogeneity in Education and Age.}}
+    \caption{\textbf{Heterogeneity in Education and Age.} Party members tend to be less educated and older than their non-partisan counterparts.}
     \label{fig:partisan_edu_et_age}
 \end{figure}
+
+This empirical record suggests important dynamics occurring in the selection phase of the bureaucracy: who enters the public sector.
+Another important question is once these party members enter the bureaucracy, where do they go to.
+The next section explores empirical regularities with respect to the concentration of party members once they are in the local bureaucracies, or islands of patronage.
 
@@
 
-Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors. These are also the best
+Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors.
+These are also the best paying positions and the ones with the highest proportion of career protection in the form of contract tenure (\emph{estatutário}).
+Moving down the hierarchy, it is clear that party members predominantly occupy high level positions, in particular staying at the
 

### 2021-03-24 paper/draft.tex
frontline providers

@@
 Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors.
-These are also the best paying positions and the ones with the highest proportion of career protection in the form of contract tenure (\emph{estatutário}).
-Moving down the hierarchy, it is clear that party members predominantly occupy high level positions, in particular staying at the
+These are also the best paying positions and the ones with the highest proportion of career protection in the form of contract tenure (\emph{estatut'{a}rio}).
+Moving down the hierarchy, it is clear that party members predominantly occupy high level positions, with a much larger prevalence of party members in bureaucrat low, frontline high and then frontline low, as compared to other positions in the bureaucracy.
 

### 2021-03-25 paper/draft.tex
include additional info on proportion of party members

@@
     \centering
-    \includegraphics[width = 8cm, height = 8cm]{figures/maps/pooled.png}
+    \includegraphics[width = 9cm, height = 9cm]{figures/maps/pooled.png}
     \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}.
@@
 
-Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors.
-These are also the best paying positions and the ones with the highest proportion of career protection in the form of contract tenure (\emph{estatut'{a}rio}).
+Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors, as illustrated by figure \ref{fig:partisan_breakdown}.
+While there is wide variation with respect to the proportion of partisans, the median proportion of party members in top-level bureaucrat positions is 28 percent in the time period between  These are also the best paying positions and the ones with the highest proportion of career protection in the form of contract tenure (\emph{estatut'{a}rio}).
 Moving down the hierarchy, it is clear that party members predominantly occupy high level positions, with a much larger prevalence of party members in bureaucrat low, frontline high and then frontline low, as compared to other positions in the bureaucracy.
@@
     \centering
-    \includegraphics[width = 8cm, height = 8cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
-    \caption{\textbf{Proportion of party-affiliated members by occupation}.
-Darker colors indicate a larger degree of partisan affiliation.
-Sample includes all bureaucrats, from high-level managers to service staff.}
-    \label{fig:partisan_by_cbo}
+    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
+    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_contract_type.pdf}
+    \caption{\textbf{Proportion of party-affiliated members}. On the left, breakdown of party membership by types of occupation. On the right, by type of contract.}
+    \label{fig:partisan_breakdown}
 \end{figure}
+
+Additionally, party members are predominantly concentrated in career positions, which are protected from dismissals and departures.
+In contrast, temporary and other regular types of contract subject to turnover are relatively more protected from partisan influence.
+This empirical finding has substantive implications with respect to the nature of patronage in Brazil and developing countries more broadly.
+A large body of literature around civil service systems has proposed that establishing a career track with autonomy from political processes is a key development in protecting bureaucracies from political influence \citep{grindle2012jobs,carpenter2020forging}.
+
+However, just because there are well established forms of career service does not mean that partisan affiliation is necessarily filtered out in the selection process.
+In particular, career services, with the exception of a formal requirement to do some form of public examination, does not necessarily prevent bureaucrats from joining a political party - and seldom do so.
+In fact, career services only structure the progression or retention of bureaucrats once they are in civil service, meaning that even if these positions are in theory autonomous from politically motivated turnover, these mechanisms do not prevent the selection of individuals who are already affiliated to a party from joining the ranks of career bureaucrats.
 

### 2021-03-25 paper/draft.tex
the recruitment of local elites

@@
     \centering
-    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
-    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_contract_type.pdf}
+    \includegraphics[width = 7cm, height = 7cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
+    \includegraphics[width = 7cm, height = 7cm]{figures/partisanship/plot_partisanship_by_contract_type.pdf}
     \caption{\textbf{Proportion of party-affiliated members}. On the left, breakdown of party membership by types of occupation. On the right, by type of contract.}
@@
 In fact, career services only structure the progression or retention of bureaucrats once they are in civil service, meaning that even if these positions are in theory autonomous from politically motivated turnover, these mechanisms do not prevent the selection of individuals who are already affiliated to a party from joining the ranks of career bureaucrats.
+
+Finally, party members once in the bureaucracy both earn more and stay for longer than their non-partisan counterparts.
+Figure \ref{fig:partisan_trajectory} illustrates these trends.
+Between 2003 and 2017, party members earned an average premium of around 22 percent over their non-partisan counterparts.
+Additionally, they have worked for longer in the bureaucracy as well, with an average 16 percent higher tenure in the municipal government.
+Note that while we aggregate these different figures across the years, there is widespread variation across municipalities in these different premiums.
+Unpacking these average values is an important avenue for future research.
+
+\begin{figure}
+    \centering
+    \includegraphics[width = 7cm, height = 7cm]{figures/turnover/plot_wage.pdf}
+    \includegraphics[width = 7cm, height = 7cm]{figures/turnover/plot_work_experience.pdf}
+    \caption{\textbf{Partisan trajectories in the bureaucracy.} Party members earn more while working in the bureaucracy (left) and also tend to work for longer (right).}
+    \label{fig:partisan_trajectory}
+\end{figure}
+
+In sum, the evidence suggests that party members concentrate in the summits of power.
+This reinforces priors regarding the control of bureaucratic management by politically appointed bureaucrats in a quantifiable way and in the context of a large, developing country with decentralized structure of local governance \citep{lewis2010politics}.
+Connecting this empirical evidence with the record that the majority of party members are coming from the upper strata of the formal economy, it becomes evident that party members are in fact occupying the upper echelons of the private and public economies, accruing important financial gains in the process and over the long run.
+
+This raises the question of what are the broader implications of local party organizations on the compensation and incentives for its members to join.
+Is this simply a club good that provides access to public sector employment?
+And why are only members of the wealthier classes being recruited?
+In the next section, I explore the incentives for parties to recruit the local elites for running campaign elections.
+Brazil contains detailed information on party contributions that I leverage to explore the motivation for recruitment of local elites.
 

### 2021-03-29 paper/draft.tex
redraft intro

@@
 
-Who benefits from patronage? And what are its consequences for local administration?
-
-Literature on patronage and selection: \citet{robinson2013political}.
-Extant literature on selection of public officials have primarily focused on who becomes a politician \citet{dal2017becomes}.
-There remains unanswered questions regarding who enters the bureaucracy, what types of employment they held, and what are the differential compensations among public officials.
+Who benefits from patronage? What motivates allocation of public sector jobs in the first place? A
 
@@
 
-One thing is clear: party ties at the electoral level are durable, with most voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}.
+One thing is clear: party ties at the electoral level are durable, with many voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}.
 Noting that party affiliations are registered at the municipal level, this empirical fact aligns with qualitative evidence provided by \citet{palmeira1995comicios}, who notes that parties at the local level constitute political factions (\emph{grupos pol\'{i}ticos} ) with well-defined boundaries and power disputes.
@@
 
-Data on partisan affiliation is detailed as well.
+Data on partisan affiliation in Brazil is rich.
 The \emph{TSE}, the national electoral authority, collects data on all party members in the national territory, including commencement and termination dates, affiliated party, municipality and name of the registered party member, as well as their electoral title.\footnote{The data is available \href{https://filia-consulta.tse.jus.br/}{here}.} Parties are mandated by electoral law to report all party members as well as providing updated registry information throughout the year.
@@
 A few patterns emerge.
-First, the distribution of spells is rather uniform, with spikes every four years that mirror the jumps in partisanship in Figure \ref{fig:number_partisan}.
+First, the distribution of spells is uniform, with spikes every four years that mirror the jumps in partisanship in Figure \ref{fig:number_partisan}.
+There are no clear patterns with respect to skewedness, suggesting that partisan affiliation duration is distributed in a wave-like pattern, with discontinuities that reflect the occurrence of municipal elections rather than other, long-term dynamics.
+Overall, the evidence on party membership spells suggests that a constant stream of party members is interspersed with local elections.
+However, once these membership ties are formed, they are stable and long-lasting.
 
@@
     \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_spell.pdf}
-    \caption{\textbf{Party memebership spells}. Data includes all party memberships that start in 1985 - the Brazilian democratic transition - and 2019.}
+    \caption{\textbf{Party membership spells}.
+Data includes all party memberships that start in 1985 - the Brazilian democratic transition - and 2019.
+Note that there are no significant skewedness in the data either to the left or to the right, indicating that there is a constant stream of party members with the exception of local election spikes.}
     \label{fig:party_spell}
@@
 
-In sum, the evidence suggests that party members concentrate in the summits of power.
+In sum, the evidence suggests that party members concentrate in the upper echelons of the local bureaucracy.
 This reinforces priors regarding the control of bureaucratic management by politically appointed bureaucrats in a quantifiable way and in the context of a large, developing country with decentralized structure of local governance \citep{lewis2010politics}.
-Connecting this empirical evidence with the record that the majority of party members are coming from the upper strata of the formal economy, it becomes evident that party members are in fact occupying the upper echelons of the private and public economies, accruing important financial gains in the process and over the long run.
+Connecting this empirical evidence with the record that the majority of party members are coming from the upper strata of the formal economy, it is clear that party members are in fact transiting between the upper echelons of the private and public economies, accruing important financial gains in the process and over the long run.
 
@@
 
+\section{Campaign Finance and the Need for Wealthy Patrons}
+Municipal elections are expensive.
+In 2016, municipal races commanded an average amount of \emph{X} amounts of dollars.
+In order to finance these races, politicians rely on a variety of resources, and at the local level, these are primarily party-member donations.
+In particular, while in the larger municipalities there is a greater reliance on corporate donations, for the majority of smaller municipalities, the contribution are in large part individual-level contributions.\
+
+Figure \emph{x} illustrates this pattern.
+
 \section{Empirical strategy}
 \label{sec:empirical}
+The previous sections present suggestive evidence of systematic differences acros partisan and non-partisan members.
+In particular, party members seem to be wealthier, less educated and concentrated in the upper strata of the local bureaucracy.
+However, it is not clear from these aggregate statistics whether the higher compensation is due to
+
+In this section, I perform a simple Oaxaca decomposition to pin down heterogeneous compensation across party and non-partisan members with respect to observables.
+In particular, I model wages as a function of age, work experience, education level and gender.
+To allow for greater flexibility with respect to compensation across states, I perform the decomposition broken down by state, and allow coefficients to vary by state, year, and
 

### 2021-03-29 paper/draft.tex
final paper

@@
 
-Who benefits from patronage? What motivates allocation of public sector jobs in the first place? A
+Who benefits from patronage in local bureaucracies?
+Are there systematic differences between patronage beneficiaries and their non-partisan counterparts?
+And what are its implications on local public sector employment?
+Patronage in developing countries is a wide-spread and well-documented phenomenon \citep{robinson2013political, grindle2012jobs}.
+Previous research on the topic has shown that the prevalence of patronage appointments in the bureaucracy can lead to lower economic development \citep{evans1999bureaucracy,kohli2004state}, as well as instability in the execution of policy and the primacy of short-term political gains over long-term provision of public goodss \citep{remmer2007political}.
 

### 2021-03-29 paper/draft.tex
appointment

@@
 
-Empirical research on allocation of public sector jobs has consistently found that politically motivated allocation of public sector jobs is an important determinant of public employment in the developed and developing world \citep{finan2017personnel}.
+Empirical research on allocation of public sector jobs has consistently found that political motivations are an important determinant of public employment in the developed and developing world \citep{finan2017personnel}.
 In the United States, appointments to the federal bureaucracy involve considerations of party loyalty and ideological alignment with the president \citep{lewis2010politics, hollibaugh2014presidents}.
@@
 These findings have provided important insights into the drivers of patronage appointments into employment in the public sector.
-Yet as important as the drivers of public sector appointment are the characteristics of those entering the bureaucracy \citep{finan2015personnel}.
+Yet the literature has been relatively silent about the characteristics of those entering the bureaucracy \citep{finan2015personnel}.
 Who becomes a bureaucrat?

### 2021-03-29 paper/draft.tex
add data pipeline

@@
 
+\subsection{Joining Public Sector Employment to Partisanship Data}
+One of the main contributions of this paper is performing a join of public sector employment data to partisanship data, to allow us to identify and follow partisan affiliates in their entire career.
+The key challenge in joining is the low coverage of keys that allow for a unique match across these databases.
+The problem is as follows.
+Only a subset of party members have the unique national id, \emph{CPF},associated with their registration, which is uniquely identified by the national electoral title (\emph{t'{i}tulo eleitoral}).
+
+To maximize coverage, the party membership data is first joined to the candidacy data generated by the \emph{TSE} for every electoral cycle using the electoral title as the key.
+This allows us to recover over 300 thousand national id's.
+Additionally, we use the complete names of both the party members and the \emph{RAIS} employees to perform an exact name match for every municipality in the dataset for active employees and party members, by year.
+To prevent many to many merges, complete names are deduplicated by municipality, our smallest geographical unit.
+Deduplication leads only to minimal data loss: for the \emph{RAIS} data we exclude around 6 percent of the names database per year, while for the party membership data we exclude around 5 percent.
+
+In total, we achieve the match of 18.5 million unique matches between the the \emph{RAIS} dataset and the party membership data.
+To give a sense of the scale of the data ETL (extraction, transformation and loading): in the year of 2015 there were in total 54 million employees in the \emph{RAIS} data, while in the party membership there were 9.2 million active membership ties.
+Partitioning the data into manageable chunks requires the development of both a semantic layer and data pipelines that are often beyond the scope of political science and enter the domain of data science and functional programming that fall under the domain of computer science.\footnote{For great resources on the topic, see the following \href{https://r4ds.had.co.nz/index.html}{\emph{R for Data Science}} resource.}
+
 \subsection{Who enters the bureaucracy?}

### 2021-03-29 paper/draft.tex
add final section regarding patronage

@@
 To give a sense of the scale of the data ETL (extraction, transformation and loading): in the year of 2015 there were in total 54 million employees in the \emph{RAIS} data, while in the party membership there were 9.2 million active membership ties.
-Partitioning the data into manageable chunks requires the development of both a semantic layer and data pipelines that are often beyond the scope of political science and enter the domain of data science and functional programming that fall under the domain of computer science.\footnote{For great resources on the topic, see the following \href{https://r4ds.had.co.nz/index.html}{\emph{R for Data Science}} resource.}
+Partitioning the data into manageable chunks requires the development of both a semantic layer and data pipelines that are often beyond the scope of political science and enter the domain of data science and functional programming that fall under the domain of computer science.\footnote{For great resources on the topic, see the following \href{https://r4ds.had.co.nz/index.html}{\emph{R for Data Science}} resource.} The entire data pipeline is documented \href{https://github.com/galileukim/ego_patronage}{here}.
 
-\subsection{Who enters the bureaucracy?}
+\subsection{Who benefits from patronage?}
 
@@
 There are competing views on who benefits from patronage, and what is motivating it in the first place.
-On the one hand, from a politician's perspective it can be cheaper to buy voters who are poor, insofar as these voters need and value more monetary incentives \citet{stokes2013brokers}.
+On the one hand, it can be cheaper to buy voters who are poor, insofar as these voters need and value more monetary incentives \citet{stokes2013brokers}.
 Other scholars argue that in order to maximize electoral returns, politicians target swing voters, those who are indifferent to either party and therefore, again, require the least resources to buy off \citet{dixit1996determinants}.
@@
 
-\section{Campaign Finance and the Need for Wealthy Patrons}
-Municipal elections are expensive.
+\section{Campaign Finance and Wealthy Patrons}
+Municipal elections are expensive to finance.
 In 2016, municipal races commanded an average amount of \emph{X} amounts of dollars.
-In order to finance these races, politicians rely on a variety of resources, and at the local level, these are primarily party-member donations.
-In particular, while in the larger municipalities there is a greater reliance on corporate donations, for the majority of smaller municipalities, the contribution are in large part individual-level contributions.\
+In order to pay for these electoral races, politicians rely on a variety of resources, and at the local level, these are primarily party-member donations.
+In particular, while in the larger municipalities there is a greater reliance on corporate donations, for the majority of smaller municipalities, the contribution are in large part individual-level contributions.
+Securing access to wealthy patrons willing to donate is key
 
-Figure \emph{x} illustrates this pattern.
+Figure \emph{x} illustrates this pattern. It shows for the year 2008, the only year in which there is data on the electoral title available, who donates to politicians.
 
-\section{Empirical strategy}
-\label{sec:empirical}
-The previous sections present suggestive evidence of systematic differences acros partisan and non-partisan members.
+\begin{figure}
+    \centering
+    \includegraphics[width = 8cm, height = 8cm]{figures/partisanship/plot_contribution.pdf}
+    \caption{\textbf{Breakdown of Individual Donation by Party Membership.} We }
+\end{figure}
+
+% \section{Empirical strategy}
+% \label{sec:empirical}
+% The previous sections present suggestive evidence of systematic differences acros partisan and non-partisan members.
 In particular, party members seem to be wealthier, less educated and concentrated in the upper strata of the local bureaucracy.
@@
 
-In this section, I perform a simple Oaxaca decomposition to pin down heterogeneous compensation across party and non-partisan members with respect to observables.
+% In this section, I perform a simple Oaxaca decomposition to pin down heterogeneous compensation across party and non-partisan members with respect to observables.
 In particular, I model wages as a function of age, work experience, education level and gender.
@@
 
-\emph{perform an Oaxaca decomposition to understand the variant compensations}
+% \emph{perform an Oaxaca decomposition to understand the variant compensations}
 
